### 2013-06-27 inst/doc/vingette.Rmd
README updated to include basic example

reml
====

R package for reading, writing, integrating and publishing data using the Ecological Metadata Language (EML) format.

Quick Start
===========

Install the R package:

<!-- r highlighting on github is broken, so we'll use {r} codeblocks -->

```{r}
#install_github("reml", "ropensci")
library("reml")
```

Writing R data into EML
-----------------------

Configure general metadata you may want to reuse again and again through settings:

```{r}
 eml$set(givenName = "Carl", surName = "Boettiger", email = "cboettig@ropensci.org")
```

Consider some dataset as an R `data.frame`.

```{r}
  dat = data.frame(river=c("SAC", "SAC", "AM"),
                        spp = c("king", "king", "ccho"),
                        stg = c("smolt", "parr", "smolt"),
                        ct =  c(293L, 410L, 210L))
```

Provide definitions for the columns.  These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful. See "Advanced Use" for details on adding richer information, such as the method used to collect the data or set the geographic, taxanomic, or temporal coverage of an individual column.

```{r}
      col_metadata = c(river = "http://dbpedia.org/ontology/River",
                       spp = "http://dbpedia.org/ontology/Species",
                       stg = "Life history stage",
                       ct = "count")
```

Define the units used in each column.  For factors, this is a definition of the levels involved.  For numeric data, specify the units from [this list](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-unitTypeDefinitions.html#StandardUnitDictionary).  For dates, specify the format, (e.g. YYYY or MM-DD-YY). For character strings, a definition of the kind of string can be given, (e.g. species scientific name), otherwise the column description will be used.

```{r}
      unit_metadata =
       list(river = c(SAC = "The Sacramento River", AM = "The American River"),
            spp = c(king = "King Salmon", ccho = "Coho Salmon"),
            stg = c(parr = "third life stage", smolt = "fourth life stage"),
            ct = "number")

```

The hard work is over, time to generate some EML.

```{r}
eml_write(dat, col_metadata, unit_metadata, file="my_eml_data.xml")
```

See the [EML generated](https://github.com/ropensci/eml/tree/master/inst/examples/my_eml_data.xml) by this example.

Publish EML
-----------

Reading EML
-----------

Integrating Multiple EML files into a single data frame
-------------------------------------------------------

### 2013-06-27 R/eml_write.R

@@ -4,6 +4,7 @@
 #' @import XML
 #' @export
 #' @examples
+#'  eml$set(givenName = "Carl", surName = "Boettiger", email = "cboettig@@gmail.com")
 #'  dat = data.frame(river=c("SAC", "SAC", "AM"), 
 #'                    spp = c("king", "king", "ccho"), 
 #'                    stg = c("smolt", "parr", "smolt"),
@@ -21,11 +22,13 @@
 #'  node = eml_write(dat, col_metadata, unit_metadata) 
 #
 eml_write = function(..., file = "eml_data.xml"){
-  template = xmlParse("inst/examples/hf250.xml")
+  eml_namespaces = c(eml = "eml://ecoinformatics.org/eml-2.1.0", 
+                     ds = "eml://ecoinformatics.org/dataset-2.1.0",
+                     xs = "http://www.w3.org/2001/XMLSchema",
+                     xsi = "http://www.w3.org/2001/XMLSchema-instance",
+                     stmml = "http://www.xml-cml.org/schema/stmml-1.1")
 
-  emlroot = newXMLNode("eml:eml", 
-                        namespaceDefinitions =
-                        xmlNamespaceDefinitions(template, simplify=TRUE))
+  emlroot = newXMLNode("eml:eml", namespaceDefinitions = eml_namespaces)
   emldoc = newXMLDoc(node = emlroot)
   dataset = eml_dataset(...)
   addChildren(emlroot, dataset)

### 2013-06-28 inst/doc/reml.Rmd
draft manuscript

* The Ecological Metadata Language (EML) has been around for a while now, but has so far seen limited use.

* R is a widely adopted among the community, many researchers already familiar with reading and writing data from R.

* R provides a rich set of data types and constructs that can be mapped into EML (e.g. integers, floats, ordered/nominal factors, character strings)

### Current state of EML

* Generation of EML is limited.  e.g. to LTER centers with informatics support. 
* Applications of existing EML more limited still (?)

* Software to work with EML is limited to tools with low adoption rates across the community (e.g. Kepler).

 Researchers are isolated from the metadata encoding, instead providing plain-text descriptions in MS Word documents.  Allows for low barrier of entry but also makes a barrier to richer encoding.

### Promise of REML

### Authoring tools

### Parsing tools

### Publishing tools

### 2013-06-28 R/eml_dataTable.R
documentation updated

@@ -44,7 +44,7 @@ eml_dataTable = function(dataframe, col_metadata, unit_metadata,
 
   # Write the data to a csv file
   if(is.null(csvfile))
-    csvfile = gsub("(.{16}).+", "\\1", paste(paste(names(dataframe), collapse="_"), ".csv", sep = ""))
+    csvfile = paste(gsub("(.{16}).+", "\\1", paste(names(dataframe), collapse="_")), ".csv", sep = "")
   write.csv(dataframe, file=csvfile) 
 
   if(is.null(file_description))

### 2013-06-29 inst/doc/reml.Rmd
documentation

@@
+## Why EML?
+
+Archival, open data is great ... blah blah.  Good reuse is all about good Metadata blah blah...
+
+But specifying metadata is typically onerous, time-consuming and error prone.
+Your metadata doesn't define your date format/missing value character/etc etc?
+What, that's metadata too?
+Wouldn't it be nice if someone thought of all the things we should record and just made a template we could use?
+
+Oh, wait, they did.  It's called EML, and it's been around a long time.  Like carved in petroglyphs long time, in the days before a public internet existed.
+
+Oh, you haven't heard of it?  That's probably because what should be a time-saving template appears more like a huge technological barrier best left to tech geeks.   But it isn't.
+
+> Terence Parr once said "being an expert in XML is like being an expert in CSV".
+
+Or at least, it doesn't have to be that way.  We just need better tools.
+
+## Why REML?
+
 * The Ecological Metadata Language (EML) has been around for a while now, but has so far seen limited use.

### 2013-06-29 inst/doc/vingette.Rmd
updated documentation
documentation updated
Updated readme generated from inst/doc/vignette.Rmd

@@
 
-R package for reading, writing, integrating and publishing data using the Ecological Metadata Language (EML) format.
+An R package for reading, writing, integrating and publishing data using the Ecological Metadata Language (EML) format.
+
+* Maintainer: Carl Boettiger
+* Authors: Carl Boettiger, Matt Jones, Karthik Ram, Duncan Temple Lang
+* License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
+* This is a dynamic document, generated by [knitr](http://yihui.name/knitr) using [vignette.Rmd](https://github.com/ropensci/reml/blob/master/inst/doc/vingette.Rmd) as the source.
+
+```{r include=FALSE}
+source("~/.knitr_defaults.R")
+```
 
@@
 
-<!-- r highlighting on github is broken, so we'll use {r} codeblocks -->
-
 ```{r}
-#install_github("reml", "ropensci")
+library("devtools")
+install_github("XMLSchema", "omegahat")
+install_github("reml", "ropensci")
 library("reml")
@@
 ```{r}
- eml$set(givenName = "Carl", surName = "Boettiger", email = "cboettig@ropensci.org")
+ eml$set(contact_givenName = "Carl", contact_surName = "Boettiger", contact_email = "cboettig@ropensci.org")
 ```
@@
 
-See the [EML generated](https://github.com/ropensci/eml/tree/master/inst/examples/my_eml_data.xml) by this example.
+See the [EML generated](https://github.com/ropensci/eml/tree/master/inst/doc/my_eml_data.xml) by this example.
+
+#### In development.
+
+* See [Open issues in writing EML](https://github.com/ropensci/reml/issues?labels=write+eml&page=1&state=open)
 
@@
 
+Once we have generated a basic EML file, we can publish it to a variety
+of academic repositories where it will be indexed, permenently archived,
+more easily citable, and publicly discoverable.
+
+In this example, we will publish the above
+EML to [figshare](http://figshare.com) using the
+[rfigshare](https://github.com/ropensci/rfigshare) package. Like most
+academic repostories, figshare requires some basic metadata for any entry.
+Much of this we can automatically extract from our existing EML file. 
+Additional metadata unique to figshare (e.g. matching it's allowed "categories"
+can also be provided.
+
+```{r publish_figshare, eval=FALSE} <!-- We don't want to generate a DOI every time we run the vignette -->
+id = eml_publish("my_eml_data.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML")
+```
+
+In return, figshare provides the object with a DOI, which is added to the EML.
+
+```{r figshare_doi, eval=FALSE}
+doc <- eml_read("my_eml_data.xml")
+citation(doc)
+```
+
+For the sake of completeness, the explicitly figshare metadata is also written into a 
+dedicated `<additionalMetadata>` node with the attribute `id = figshare`.  This allows
+us to distinguish between metadata that is an is not available to figshare's indexing,
+e.g. if we use the figshare search tool.
+
+#### In development
+
+See [Open issues in publishing EML](https://github.com/ropensci/reml/issues?labels=publish+eml&page=1&state=open)
+
 Reading EML
@@
 
+#### In development
+
+See [Open issues in reading EML](https://github.com/ropensci/reml/issues?labels=read+eml&page=1&state=open)
+
 Integrating Multiple EML files into a single data frame
 -------------------------------------------------------
+
+#### In development
+
+See [Open issues in integrating EML](https://github.com/ropensci/reml/issues?labels=integrate+eml&page=1&state=open)

### 2013-06-30 inst/doc/vingette.Rmd
Updated example XML file and vingette to show new eml_write command option for title.
Fixed typo in variable name that caused bug whereby the entityDescription field was blank, and updated the vignette and example doument to show the description field.
update import dependencies

@@
 ```{r}
+install.packages("uuid",,'http://rforge.net/',type='source')
 library("devtools")
@@
 ```{r}
-eml_write(dat, col_metadata, unit_metadata, file="my_eml_data.xml")
+eml_write(dat, col_metadata, unit_metadata, .title="A test EML file", file="my_eml_data.xml", file_description="Test data, only intended for testing")
 ```

### 2013-06-30 R/eml_figshare.R

@@ -21,7 +21,7 @@
 #' can take the figshare_id to change any fields of an existing figshare object, 
 #' and in doing so will also update the EML file.  
 #' 
-#' @import rfigshare
+#' @import XML
 #' @export
 #' @examples 
 #' \dontrun{

### 2013-07-07 inst/doc/vingette.Rmd
updated README based on inst/doc/vignette.Rmd

@@
 
-```{r}
+```{r, results="hide", message=FALSE, warning=FALSE, eval=FALSE}
 install.packages("uuid",,'http://rforge.net/',type='source')
@@
 install_github("reml", "ropensci")
+```
+
+Load the package:
+
+```{r, results="hide", message=FALSE, warning=FALSE}
 library("reml")
@@
 ```{r publish_figshare, eval=FALSE} <!-- We don't want to generate a DOI every time we run the vignette -->
-id = eml_publish("my_eml_data.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML")
+id = eml_publish("my_eml_data.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML", destination="figshare")
 ```
 
+`eml_publish` 
 In return, figshare provides the object with a DOI, which is added to the EML.

### 2013-07-07 inst/doc/reml-blog-post.Rmd
documentation updated

--- 2013-06-29 inst/doc/reml.Rmd
@@
-## Why EML?
+You know the spiel.  Archivally preserved open data is great [1]. For you [2], for your citations [3], for your grants, [4] and of course, for scientific progress more generally [5].
 
-Archival, open data is great ... blah blah.  Good reuse is all about good Metadata blah blah...
+And you know that making data reusable is all really all about having good metadata [6-398].
 
@@
 
-> Terence Parr once said "being an expert in XML is like being an expert in CSV".
+Or at least, it doesn't have to be that way.  We just need better tools.
 
-Or at least, it doesn't have to be that way.  We just need better tools.
+You already use XML every day.
+Microsoft office?
+Those `.docx`, `xlsx`, etc., are just a bunch of XML files zipped together (just try `unzip file.docx`)  The tools are just good enough to hide the XML from you.
+XML is so general that it's essentially meaningless to say a file is XML.
+Could be anything.
+Terence Parr once said "being an expert in XML is like being an expert in CSV".
+
+That's where schema come in.
 
@@
 * R provides a rich set of data types and constructs that can be mapped into EML (e.g. integers, floats, ordered/nominal factors, character strings)
-
-### Current state of EML
-
-* Generation of EML is limited.  e.g. to LTER centers with informatics support. 
-* Applications of existing EML more limited still (?)
-
-* Software to work with EML is limited to tools with low adoption rates across the community (e.g. Kepler).
-
- Researchers are isolated from the metadata encoding, instead providing plain-text descriptions in MS Word documents.  Allows for low barrier of entry but also makes a barrier to richer encoding.
-
-### Promise of REML
-
-### Authoring tools
-
-### Parsing tools
-
-### Publishing tools

### 2013-07-07 inst/doc/reml.Rmd
documentation updated

--- 2013-06-28 inst/doc/reml.Rmd
@@
+Abstract
+
+1. Recording good metadata is at the heart of effective data synthesis and reuse
+2.
+
+Vertically integrated repositories vs metadata driven repositories.
+
+Data synthesis
+
+## Why EML?
+
+But specifying metadata is typically onerous, time-consuming and error prone.
+
+Such definitions are also not machine readable.
+
+Or at least, it doesn't have to be that way.  We just need better tools.
+
+## Why REML?
+
 * The Ecological Metadata Language (EML) has been around for a while now, but has so far seen limited use.
@@
 
+### Getting Started: A minimal example.
+
+Imagine we are working with a `data.frame` in R analgous to Table 1.
+
+<!--Writing a data.frame into EML -->
+
+Writing this table into EML
+
 ### Authoring tools

### 2013-07-09 inst/doc/reml.Rmd
writing draft

@@
+The future belongs to those who can leverage existing data and provide reusable data.
+
+1. Reuse is essential
+2. Reuse requires good metadata standards
+3. reml provides an easy way to utilize the Ecological Metadata Language to generate, access, and publish data.
+
+Data in ecology and evolutionary biology research collected for a
+particular purpose can often prove valuable to an entirely different
+line of research. One study may collect morphometric measurements of
+coral reef fish taken to understand their evolution, while another
+reuses only the species names and geographic locations of the samples,
+along with other data, to aid the design of marine reserves.  Such reuse
+is fundamental to much of our research, and yet it is often difficult
+or laborious to discover and obtain. The publications are the only
+documentation of what data is available.  The distribution and discovery
+of publications is geared towards reaching audience interested in common
+set of questions, not a common set of data. Even within a subdispline,
+re-use of data is often laborious or impossible due to lack of data or
+sufficient documentation of the data.
+
+Vertically intergrated data repositories such as Genbank, GBIF [Canhos 2004],
+or TreeBase [Morell 1996] provide one route to addressing the challenges 
+of data integration.  These data warehouses provide all there data in a common 
+structure that can usually be queried through a web interface. This makes it
+relatively easy to discover available data and access it in a consistent and
+predictable format.  However, such utility comes at a cost.  This approach
+discards the original data model, preserving only the data that can be shoehorned
+into a common data standard.
+
+<!-- is "data model" more confusing than "data format"? -->
+
+These challenges have long been recognized and broadly accepted that
+they can be best addressed through establishing and using appropriate
+metadata [Michener 1997]. In this approach it is the metadata, rather than
+the data itself, that conforms to a standard.  The Knowledge Network for
+Biocomplexity (KNB) is one example of a metadata-based ecological data 
+repository [Berkley 2001], [KNB].  The metadata approach preserves the original data
+model from the collecting study. Rather than requiring that the raw data
+be preserved in a standard format -- standard file types, standard column headings, 
+standard date formats, missing value formats, etc etc, -- the metadata standard
+approach provides a consisent way to document whatever data model the researchers
+use.  In principle, this approach is both more powerful than vertical integration, 
+(as the full context of the data is preserved), and easier to implement for both
+researcher and repository (as the researcher does not to to convert data to 
+the format required by the repository, and very heterogenous data models can be 
+stored and searched for in a single system).   [Jones 2006] provides
+an excellent explanation and comparison of these approaches.
+
+The Ecological Metadata Language
+
+Consequently ecological
+data reuse remains difficult and haphazard. __@mbjones: is that unfair?__
+
+Widely adopted computing environments such as the R statistical sofware
+provide a key opportunity to increase adoption of metadata standards.
+
+Methods in Ecology and Evolution has published XXX application notes or articles presenting new software packages.  XXX of them are R packages.
+
+[Jones 2006]: http://dx.doi.org/10.1146/annurev.ecolsys.37.091305.110031
+[KNB]: http://knb.ecoinformatics.org
+[Berkley 2001]: http://dx.doi.org/10.1109/4236.957896
+[Michener 1997]: http://dx.doi.org/10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2
+[Morell 1996]: http://dx.doi.org/10.1126/science.273.5275.569
+[Canhos 2004]: https://journals.ku.edu/index.php/jbi/article/viewArticle/3
+
+1. Commonly used data structures in R provide basic metadata automatically
+2. We can more easily avoid redundant metadata entry 
+3. Automate quality control checks
+
+4. Publish
+5. Search 
+6. Build a synthesis platform
+
+### Challenges
+
+Completely automated data integration will require rich semantic information. Consider the simple example of combining the two data tables shown in Figure 2.
+
+The EML schema already provides some
+
+Semantics.
+
+It's all metadata.
+
+Share 
+Metadata
+raw data
+formats: file formats, table formats, cell formats 
+good null values
+easy to combine (context data)
+quality control
+repository
+license
+
 Abstract

### 2013-07-10 inst/doc/reml.Rmd
writing outline

@@
 
-The Ecological Metadata Language
+Unfortunately, adoption of these metadata standards for annotating research
+data has been limited, owing to both technological and social challenges 
+involved [Reichman 2011].
+
+Things to discuss:
+
+*  It's not just a problem of sharing culture 
+* Sharing is not effective: close competitors can scoop you, but broader reach can't be realized without good metadata 
+* [Strasser survey showing lagging data management practices]
+* [Stodden survey showing effort in curating data is a barrier]
 
@@
 
+[Reichman 2011]: http://dx.doi.org/10.1126/science.1197962
 [Jones 2006]: http://dx.doi.org/10.1146/annurev.ecolsys.37.091305.110031
@@
 [Canhos 2004]: https://journals.ku.edu/index.php/jbi/article/viewArticle/3
+
+[Michener 2006]: http://dx.doi.org/10.1016/j.ecoinf.2005.08.004
+[Michener 2012]: http://dx.doi.org/10.1016/j.tree.2011.11.016
 

### 2013-07-10 inst/doc/reml.Rmd
updated draft outline

@@
 Consequently ecological
-data reuse remains difficult and haphazard. __@mbjones: is that unfair?__
+data reuse remains difficult and haphazard.
 
@@
 [Canhos 2004]: https://journals.ku.edu/index.php/jbi/article/viewArticle/3 
-
 [Michener 2006]: http://dx.doi.org/10.1016/j.ecoinf.2005.08.004
@@
 3. Automate quality control checks
+4. Publish
+5. Arbitrarily complex search 
+6. Building a synthesis platform
 
-4. Publish
-5. Search 
-6. Build a synthesis platform
+### Automatic metadata detection
 
-### Challenges
+Not sure this counts as a section - but we can automatically detect at least the difference between numeric and character data, factors, ordered factors, etc.
+
+### Reusing fields
+
+- Examples of filling in contact information, etc using `eml$set`, using an existing EML file as a template, and using an external yaml file.
+
+- Similar example for `column_metadata` and `unit_metadata` based on an existing EML file (just using `eml_read`).
+
+### Automatic Quality Control checks
+
+<!-- Not sure if we'll actually have/need to have all this in place for this paper?  -->
+Examples of validation against schema.  Examples checking md5sums on files or other corruption checks?  Examples checking species names (taxize?)
+
+### Publish data
+
+Examples of commands for publishing EML files to external repositories.
+
+### Complex search queries
+
+Illustrate how to query for data
+
+### A synthesis platform
+
+Illustrate a basic example of synthesis.
+While arbitrary synthesis will require semantics, a lab could store their data in EML format and build up there own library of common scripts to extract the data they need.
+As a lab database grows, the same queries can continue to deliver more data.
+By scripting the data access, analyses can be automatically updated as the data set grows.
+
+Consider simple example in regression between different morphological characters of species, or such.
+
+## Challenges
 
@@
 
-The EML schema already provides some
+<!-- Example illustrating the need for semantics to resolve differences -->
 
-Semantics.
+### Implementation and architecture
+
+The `reml` package is built for the Ecological Markup Language schema, [version 2.1.1], using the R software environment and the XML package for parsing and generating XML [Temple Lang 2013, XML].
+
+<!-- Possibly cite all dependencies and suggests from the DESCRIPTION --> 
+<!-- Reference any installation and configuration instructions necessary (e.g. API credentials for figshare) -->
+
+[version 2.1.1]: http://knb.ecoinformatics.org/software/eml/eml-2.1.1
+[Temple Lang 2013, XML]: http://cran.r-project.org/web/packages/XML/
+
+### Quality control
+
+The `reml` package can perform three levels of validation on EML files written from or read into R: validating the XML, validating against the EML 2.1.1 Schema, and performing
+
+Several mechanisms are in place to evaluate the quality of the `reml` package functions themselves.
+In addition to the examples from the documentation and the automated package checks provided for the R system, the `reml` package includes a complete test suite using the testthat package [Wickham 2011].
+
+[Wickham 2011]: http://vita.had.co.nz/papers/testthat.html
+
+### Software reuse and support
+
+The `reml` package is available through the CRAN repository under a CC0 license.
+At the time of writing, the current version is `r packageDescription("reml", fields="Version")`.
+The current development branch can be found on Github at [https://github.com/ropensci/reml].
+See the NEWS file for changes in recent versions.
+Users are encouraged to submit bug reports or feature requests in the issues log.
+Contact ropensci-discuss@googlegroups.com or the package maintainer for help.
+
+There are many promising directions for further development of the `reml` package.
+The project's Milestones page provides a list of steps currently planned or under consideration.
+In particular, these milestones include increased support for interactive mode of entering data (wizards), and support for integration of semantics defined through the Web Ontology language (OWL), as discussed in the Challenges section.
+
+### Acknowledgements
+
+This project was supported in part by a grant from the Alfred P Sloan Foundation (CB and KR), NSF Grant DBI-1306697 (CB), ... `reml` is part of the rOpenSci project, [http://ropensci.org].
+
+OLD OUTLINE
+----------------------------------------------------------------------------------------------
 

### 2013-07-21 inst/doc/one-simple-way.Rmd
notes

# One simple way to share your data

```{r, include=FALSE}
source("~/.knitr_defaults.R")
options(xtable.type = 'html')
```

Use `reml`.

You have your data as a `data.frame` in R, convient for your analyses and graphs.  How best to archive share that data in a way that will maximize its utility to your lab group, your future self?  What will most impress your NSF data management review, or improve the chances that others will use and cite your data set?  Good metadata.

White et al. outline "Nine simple ways to make it easier to (re)use your data".  Here we present one simple way to meet these criteria.  Before we worry about what the criteria are, let's just jump in and make our archival data shared:

For sake of argument, imagine a data.frame looking something like this:

```{r, results="asis"}
dat = data.frame(river=c("SAC", "SAC", "AM"),
                 spp = c("king", "king", "ccho"),
                 stg = c("smolt", "parr", "smolt"),
                 ct =  c(293L, 410L, 210L))

xtable::xtable(dat)
```

In case our column header abbreviations are not obvious to others (or our future selves), we take a moment to define them:

```{r}
col_metadata = c(river = "http://dbpedia.org/ontology/River",
                 spp = "http://dbpedia.org/ontology/Species",
                 stg = "Life history stage",
                 ct = "count")
```

Define the units used in each column, or the abbreviations we use in factors.

```{r}
unit_metadata = 
  list(river = c(SAC = "The Sacramento River", AM = "The American River"),
       spp = c(king = "King Salmon", ccho = "Coho Salmon"),
       stg = c(parr = "third life stage", smolt = "fourth life stage"),
       ct = "number")
```

Specify the basic file metadata of author/contact, title, and a description, and choose where to publish:

```{r}
eml_write(dat, col_metadata, unit_metadata, 
          .contact="Carl Boettiger <cboettig@gmail.com>",  
          .title="A test EML file", 
          file_description="Test data, only intended for testing",
          destination="figshare")
```

We can provide much richer annotation than this example includes.  We can also store information such as the contact name and address in a configuration option to avoid having to provide it repeatedly.  Let's see how this example does on the Nine simple criteria.

## 1. Share your data

The data is publically available at the DOI returned. At this DOI we find two files: the raw data in `.csv` format, and the EML data file providing machine readable metadata.  We'll unpack all the information provided in this metadata as we go along.  Many other options for sharing are possible, including private sharing and sharing through a variety of data publication platforms.

## 2. Provide metadata

We provide `n` lines of metadata in the EML file, one of the "metadata standards" mentioned by White et al.  This includes basic information like the title, author(s), and contact address.

> 1) the what, when, where, and how of data collection,

Methods sections in the file provide answers to these questions.  Though our minimal example does not describe the collection itself, note that we nonetheless have generated a `methods` node, which describes the software that created the metadata file (after all, this is a step in the analysis).

> 2) how to find and access the data,

Our example returns a DOI and access URLs for the files in the data set:

Rich access options are possible depending on the publication outlet, including specifying what users or groups have access to the data.

> 3) suggestions on the suitability of the data for answering specific questions,

We can see at a glance what taxa are covered by the data.  More detailed description can be provided in an abstract.

4) warnings about known problems or inconsistencies in the data, and 
5) information to check that the data are properly imported, such as the number of rows and columns in the dataset and the total sum of numerical columns

Checks have automatically been performed in generating the data, and are performed in reading in the data into R.

## 3. Provide unprocsessed data

The second file provides the unprocessed data as a .csv file

## 4. Use standard formats

Of the "nine ways", this one places the greatest burden on the creator and user.  Rather than coercing our data into prescribed formats, EML lets us simply describe whatever format we are using.

- Standard file formats

Our example generate standard .csv file format, and declares this explicitly so that users can be sure what character is used to seperate values, rows, etc:

```xml
      <physical>
        <objectName>river_spp_stg_ct.csv</objectName>
        <dataFormat>
          <textFormat>
            <numHeaderLines>1</numHeaderLines>
            <recordDelimiter>\r\n</recordDelimiter>
            <attributeOrientation>column</attributeOrientation>
            <simpleDelimited>
              <fieldDelimiter>,</fieldDelimiter>
            </simpleDelimited>
          </textFormat>
        </dataFormat>
      </physical>
```

- Standard table formats
- Standard cell formats

Rather than insist everyone use the ISO8601 format for dates, the user can declare whatever date format they prefer.

## 5. Good null values

R has good null values in its use of `data.frames`, so we already have these.  Better yet, we write out our choice of null value in the metadata.  Reading in the data, we can explicitly pass this to `read.csv` rather than just assume the null value is appropriate.  This is handled automatically by `eml_read`.

## 6. Make it easy to combine

Context nodes serve the pupose described in this section.   
Machine-readable metadata,

### 7. Perform basic quality control

We have quality control on the metadata as well as the data.  While user mistakes are still possible, automated validation of the metadata provided and

## 8. Use an established repository

`reml` can automatically push our data to a variety of established repositories, as illustrated in this example.

## 9. Licenses

We see by default `reml` has provided us with the liberal CC0 license for the data, consistent with major repositories such as Dryad.  Other options can be specified if needed.

Ethan White et al have recently published an important paper on "Nine simple ways to share your data", which has garnered considerable attention even as a preprint [^1]

[^1]: Disclaimer: I reviewed this paper, Karthik Ram served as handling editor.

### 2013-07-26 inst/doc/vingette.Rmd
readme recompiled (fix parsing error)

@@
 
-See the [EML generated](https://github.com/ropensci/eml/tree/master/inst/doc/my_eml_data.xml) by this example.
+See the [EML generated](https://github.com/ropensci/reml/tree/master/inst/doc/my_eml_data.xml) by this example.
 
@@
 
-```{r publish_figshare, eval=FALSE} <!-- We don't want to generate a DOI every time we run the vignette -->
+<!-- We don't want to generate a DOI every time we run the vignette -->
+```{r publish_figshare, eval=FALSE} 
 id = eml_publish("my_eml_data.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML", destination="figshare")
@@
 
-`eml_publish` 
 In return, figshare provides the object with a DOI, which is added to the EML.

### 2013-08-04 inst/doc/one-simple-way.Rmd
writing continues, various edits

@@
 
+Sharing is as easy as `publish_reml("my_eml_data.xml")`
+
 The data is publically available at the DOI returned.
@@
 
-4) warnings about known problems or inconsistencies in the data, and 
-5) information to check that the data are properly imported, such as the number of rows and columns in the dataset and the total sum of numerical columns
+> 4) warnings about known problems or inconsistencies in the data, and 
+> 5) information to check that the data are properly imported, such as the number of rows and columns in the dataset and the total sum of numerical columns
 
@@
 
-The second file provides the unprocessed data as a .csv file
+The second file provides the unprocessed data as a `.csv` file.  Because `csv` file formats vary
 

### 2013-08-04 inst/examples/S4_draft_methods.Rmd
examples of what S4 coercion methods might look like; still working out questions

## What example functions would look like
```{r}
eml_dataset <- function(x)
  as(x, "eml:dataset")

eml_write <- function(x, file = NULL){
  eml <- as(x, "eml:eml")
  doc <- newXMLDoc(node=eml)
  saveXML(doc, file=file)
}
```

some testing data

```{r}
dat = data.frame(river = c("SAC",  "SAC",   "AM"),
                 spp   = c("king",  "king", "ccho"),
                 stg   = c("smolt", "parr", "smolt"),
                 ct    = c(293L,    410L,    210L))
```

New metadata format

```{r}
metadata <- 
  list("river" = list("river",
                      "River site used for collection",
                      c(SAC = "The Sacramento River", AM = "The American River")),
       "spp" = list("spp",
                    "Species common name", 
                    c(king = "King Salmon", ccho = "Coho Salmon")),
       "stg" = list("stg",
                    "Life Stage", 
                    c(parr = "third life stage", smolt = "fourth life stage")),
       "ct"  = list("ct",
                    "count", 
                    "number"))
```

### Coercion methods

Take this metadata and coerce it into a valid `attributeList`
A list of attributes based on this structure would look like:

```{r}

  setAs("list", "eml:attribute", 
        function(from){
         new("eml:attribute", 
             attributeName = from[[1]],
             attributeDefinition = from[[2]], 
             measurementScale = as(from[[3]], from[[4]])) 
        })
```

The trick is to decide what class measurementScale should be coerced into.  This should come from the data.frame class. Not clear if that's allowed in a standard coercion.  We could:

- [ ] Define the `setAs` method inside a function taking both the metadata list and the data table.

```{r}
detect_class <- function(dat, meta){
  for(i in length(dat)){
    meta[[i]][[4]] = map(class(dat[[i]])) 
  }
  meta 
}

map <- function(x){
  if(is(x, "numeric"))
    "eml:ratio"
  else if(is(x, "ordered"))
    "eml:ordinal"
  else if(is(x, "nominal"))
    "eml:nominal",
  else if(is(x, "POSIXlt") | is(x, "POSIXct"))
    "eml:dateTime"
  else if(is(x, "character"))
    "eml:nominal"
}
```

Then in any function that gets both dat and metadata, we can correct things:

```{r}
metadata <- detect_class(dat, metadata)
```
- [ ] Something else?

If/once we know it's a nominal, conversion is easy:

```{r}
setAs("character", "eml:nominal", 
      function(from)
        new("eml:nominal",
            nonNumericDomain = as(from, "eml:nonNumericDomain")))

setAs("character", "eml:textDomain", function(from){
  new("eml:textDomain", definition = from)
})

setAs("character", "eml:nonNumericDomain", 
      function(from){
        if (length(from) == 1)
          new("eml:nonNumericDomain",
              textDomain = as(from, "eml:textDomain"))
        else if(length(from) > 1)
          new("eml:nonNumericDomain", 
            enumeratedDomain = as(from, "ListOfCodeDefinition"))
      })

setAs("character", "ListOfCodeDefinition", 
      function(from){
        new("ListOfCodeDefinition", 
            lapply(names(from), function(name)
               new("eml:codeDefinition", code = name, definition = from[name])))
      })
```

Then convert the "unit metadata" for the river values just by:

```{r}
as(metadata$river[[3]], "nominal")
```

`nominal` is potential slot for `measurementScale`, which is a slot of an `attribute`.

Moving up the DOM tree from here is easy:

```{r}
attrList = new("eml:attributeList", as(metadata_list, "eml:attribute")
new("eml:dataSet", dataTable = new("eml:dataTable", attributeList = attrList) )
```

and so forth.  lots of permutations are possible.  We can define coercion methods and merge methods to make this all smooth and flexible.

## We either want to look at the data.frame column classes,
## Or guess from the length of the metadata

setAs("character", "eml:measurementScale",
      function(from){
        if(length(from) > 1) ## assume factor.  Not very robust!
          new("eml:measurementScale", 
              nominal = as(from, "eml:nominal"))
        else if(length(from) == 1) ## Assume a unit of 
      })

### 2013-08-12 inst/examples/S4_draft_methods.Rmd
Testing S4 based approach

@@
+```{r}
+require(reml)
+```
+
 ## What example functions would look like
@@
 ```{r}
+
 dat = data.frame(river = c("SAC",  "SAC",   "AM"),
@@
 
-Take this metadata and coerce it into a valid `attributeList`
-A list of attributes based on this structure would look like:
-
-```{r}
-
-  setAs("list", "eml:attribute", 
-        function(from){
-         new("eml:attribute", 
-             attributeName = from[[1]],
-             attributeDefinition = from[[2]], 
-             measurementScale = as(from[[3]], from[[4]])) 
-        })
-```
-
-The trick is to decide what class measurementScale should be coerced into.  This should come from the data.frame class. Not clear if that's allowed in a standard coercion.  We could:
-
-- [ ] Define the `setAs` method inside a function taking both the metadata list and the data table.
-
-```{r}
-detect_class <- function(dat, meta){
-  for(i in length(dat)){
-    meta[[i]][[4]] = map(class(dat[[i]])) 
-  }
-  meta 
-}
-
-map <- function(x){
-  if(is(x, "numeric"))
-    "eml:ratio"
-  else if(is(x, "ordered"))
-    "eml:ordinal"
-  else if(is(x, "nominal"))
-    "eml:nominal",
-  else if(is(x, "POSIXlt") | is(x, "POSIXct"))
-    "eml:dateTime"
-  else if(is(x, "character"))
-    "eml:nominal"
-}
-```
-
-Then in any function that gets both dat and metadata, we can correct things:
+Detect types
 
@@
 ```
-- [ ] Something else?
 
-If/once we know it's a nominal, conversion is easy:
+Coerce a factor definitions into a nominal,
 
 ```{r}
-setAs("character", "eml:nominal", 
-      function(from)
-        new("eml:nominal",
-            nonNumericDomain = as(from, "eml:nonNumericDomain")))
-
-setAs("character", "eml:textDomain", function(from){
-  new("eml:textDomain", definition = from)
-})
-
-setAs("character", "eml:nonNumericDomain", 
-      function(from){
-        if (length(from) == 1)
-          new("eml:nonNumericDomain",
-              textDomain = as(from, "eml:textDomain"))
-        else if(length(from) > 1)
-          new("eml:nonNumericDomain", 
-            enumeratedDomain = as(from, "ListOfCodeDefinition"))
-      })
-
-setAs("character", "ListOfCodeDefinition", 
-      function(from){
-        new("ListOfCodeDefinition", 
-            lapply(names(from), function(name)
-               new("eml:codeDefinition", code = name, definition = from[name])))
-      })
+as(metadata$river[[3]], "eml:nominal")
 ```
 
-Then convert the "unit metadata" for the river values just by:
+or coerce the entire list into an attribute node (includes the above coercion)
 
 ```{r}
-as(metadata$river[[3]], "nominal")
+as(metadata$river, "eml:attribute")
 ```
 
-`nominal` is potential slot for `measurementScale`, which is a slot of an `attribute`.
-
-Moving up the DOM tree from here is easy:
+Or all metadata lists at once:
 
 ```{r}
-attrList = new("eml:attributeList", as(metadata_list, "eml:attribute")
-new("eml:dataSet", dataTable = new("eml:dataTable", attributeList = attrList) )
+as(metadata, "eml:attributeList")
 ```
 
-and so forth.  lots of permutations are possible.  We can define coercion methods and merge methods to make this all smooth and flexible.
-
-## We either want to look at the data.frame column classes,
-## Or guess from the length of the metadata
-
-setAs("character", "eml:measurementScale",
-      function(from){
-        if(length(from) > 1) ## assume factor.  Not very robust!
-          new("eml:measurementScale", 
-              nominal = as(from, "eml:nominal"))
-        else if(length(from) == 1) ## Assume a unit of 
-      })
+```{r}
+new("eml:dataSet", dataTable = new("eml:dataTable", attributeList = as(metadata, "eml:attributeList")) )
+```

### 2013-08-12 inst/examples/S4_draft_methods.Rmd
Working on S4 methods

@@
 ```{r}
-eml_dataset <- function(x)
-  as(x, "eml:dataset")
 
@@
                  ct    = c(293L,    410L,    210L))
-```
-
-New metadata format
-
-```{r}
 metadata <- 
@@
 ```{r}
-new("eml:dataSet", dataTable = new("eml:dataTable", attributeList = as(metadata, "eml:attributeList")) )
+new("eml:dataset", dataTable = new("eml:dataTable", attributeList = as(metadata, "eml:attributeList")) )
 ```

### 2013-08-12 R/eml_write.R

@@ -47,9 +47,3 @@ eml_write = function(..., file = NULL){
 	}
 }
 
-eml_namespaces = c(eml = "eml://ecoinformatics.org/eml-2.1.1", 
-                   ds = "eml://ecoinformatics.org/dataset-2.1.1",
-                   xs = "http://www.w3.org/2001/XMLSchema",
-                   xsi = "http://www.w3.org/2001/XMLSchema-instance",
-                   stmml = "http://www.xml-cml.org/schema/stmml-1.1")
-

### 2013-09-03 inst/doc/vingette.Rmd
updated README

@@
 
+* **Note:** This package is still in active development and not yet submitted to CRAN.  Functions and documentation may be incomplete and subject to change.  
 * Maintainer: Carl Boettiger
-* Authors: Carl Boettiger, Matt Jones, Karthik Ram, Duncan Temple Lang
+* Authors: Carl Boettiger, Duncan Temple Lang, Karthik Ram, Matt Jones
 * License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
-* This is a dynamic document, generated by [knitr](http://yihui.name/knitr) using [vignette.Rmd](https://github.com/ropensci/reml/blob/master/inst/doc/vingette.Rmd) as the source.
+* [Issues](https://github.com/ropensci/reml/issues): Bug reports, feature requests, and development discussion.
 
-```{r include=FALSE}
+_This is a dynamic document, generated by [knitr](http://yihui.name/knitr) using [vignette.Rmd](https://github.com/ropensci/reml/blob/master/inst/doc/vingette.Rmd) as the source._
+
+```{r include=FALSE, cache=FALSE}
 source("~/.knitr_defaults.R")
+opts_chunk$set(cache=FALSE)
 ```
@@
 ```{r, results="hide", message=FALSE, warning=FALSE, eval=FALSE}
-install.packages("uuid",,'http://rforge.net/',type='source')
 library("devtools")
-install_github("XMLSchema", "omegahat")
 install_github("reml", "ropensci")
@@
 
-Configure general metadata you may want to reuse again and again through settings:
-
-```{r}
- eml$set(contact_givenName = "Carl", contact_surName = "Boettiger", contact_email = "cboettig@ropensci.org")
-```
-
 Consider some dataset as an R `data.frame`.
@@
 ```{r}
-  dat = data.frame(river=c("SAC", "SAC", "AM"),
+dat = data.frame(river = c("SAC", "SAC", "AM"),
                  spp   = c("king", "king", "ccho"),
@@
 
-Provide definitions for the columns.
-These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful.
+Provide a list giving the  __column header__ used in the data.frame, followed by a plain text definitions for the column, followed by a character vector giving the definition for the units:
+
+__column descriptions__: These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful.
 See "Advanced Use" for details on adding richer information, such as the method used to collect the data or set the geographic, taxanomic, or temporal coverage of an individual column.
 
-```{r}
-      col_metadata = c(river = "http://dbpedia.org/ontology/River",
-                       spp = "http://dbpedia.org/ontology/Species",
-                       stg = "Life history stage",
-                       ct = "count")
-```
-
-Define the units used in each column.
-For factors, this is a definition of the levels involved.
+__column units__:   For factors, this is a definition of the levels involved.
 For numeric data, specify the units from [this list](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-unitTypeDefinitions.html#StandardUnitDictionary).
@@
 ```{r}
-      unit_metadata =
-       list(river = c(SAC = "The Sacramento River", AM = "The American River"),
-            spp = c(king = "King Salmon", ccho = "Coho Salmon"),
-            stg = c(parr = "third life stage", smolt = "fourth life stage"),
-            ct = "number")
+metadata <- 
+  list("river" = list("river",
+                      "River site used for collection",
+                      c(SAC = "The Sacramento River", 
+                        AM = "The American River")),
+       "spp" = list("spp",
+                    "Species common name", 
+                    c(king = "King Salmon", 
+                      ccho = "Coho Salmon")),
+       "stg" = list("stg",
+                    "Life Stage", 
+                    c(parr = "third life stage", 
+                      smolt = "fourth life stage")),
+       "ct"  = list("ct",
+                    "count", 
+                    "number"))
 
@@
 
-The hard work is over, time to generate some EML.
+*for convenience, this metadata list can instead be constructed with the help of the reml wizard*.
+Just use `metadata <- metadata_wizard(dat)` to begin.
+While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.
 
 ```{r}
-eml_write(dat, col_metadata, unit_metadata, .title="A test EML file", file="my_eml_data.xml", file_description="Test data, only intended for testing")
+eml_write(dat, metadata, title = "reml example",  
+          description = "An example, intended for
+                              illustrative purposes only.",
+          creator = "Carl Boettiger <cboettig@gmail.com>",
+          file = "reml_example.xml")
 ```
 
-See the [EML generated](https://github.com/ropensci/reml/tree/master/inst/doc/my_eml_data.xml) by this example.
+See the [EML generated](https://github.com/ropensci/reml/tree/master/inst/doc/reml_example.xml) by this example.
+Note that if the `metadata` argument providing the definitions for the `data.frame` is not specified, `reml` will launch the wizard to prompt the user.
 
-#### In development.
+### Configuration
+
+Configure general metadata you may want to frequently reuse, avoiding having to specify things like the name of the creator or contact.
+
+```{r}
+eml_config(creator = list("Carl Boettiger <cboettig@ropensci.org>", "Karthik Ram"), contact = "Carl Boettiger <cboettig@ropensci.org>")
+```
 
@@
 
+_Note: this example assumes `rfigshare` is installed, even though it is not 
+required to use reml and run the examples in the other sections_
+
 <!-- We don't want to generate a DOI every time we run the vignette -->
-```{r publish_figshare, eval=FALSE} 
-id = eml_publish("my_eml_data.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML", destination="figshare")
+```{r publish_figshare} 
+id = eml_publish("reml_example.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML", destination="figshare")
 ```
 
-In return, figshare provides the object with a DOI, which is added to the EML.
+This creates a draft file visible only to the user configured in `rfigshare`.
+The document can be made (permanently) public using either the figshare web interface, the `rfigshare` package (using `fs_make_public(id)`) or just by adding the argument `visibility = TRUE` to the above `eml_publish` command.
+
+<!-- In return, figshare provides the object with a DOI, which is added to the EML.
 
@@
 doc <- eml_read("my_eml_data.xml")
-citation(doc)
+citation(doc) # not implemented yet
 ```
+-->
+
+<-- Too detailed for this vignette
 
@@
 e.g. if we use the figshare search tool.
+
+-->
 
@@
 
-Integrating Multiple EML files into a single data frame
+Integrating multiple EML files into a single data frame
 -------------------------------------------------------
@@
 See [Open issues in integrating EML](https://github.com/ropensci/reml/issues?labels=integrate+eml&page=1&state=open)
+
+Testing and development
+-----------------------
+
+While the dependencies for basic functionality are kept to a minimum, to access all the functions and tests implemented in `reml` you'll need several additional packages.
+
+```r
+install.packages(c("yaml", "knitr", "rfigshare", "testthat", "RCurl"))
+```
+
+Some of these additional packages are not yet on CRAN and may not be stable.
+
+```r
+install.packages("uuid",,'http://rforge.net/',type='source')
+install_github("RHTMLForms", "omegahat")
+install_github("XMLSchema", "omegahat")
+```
+
+Successful installation of these packages will require the `devtools` package (for `install_github`) and the ability to build packages from source.

### 2013-09-06 R/responsibleParty.R
added accessor methods for eml_read, see README

@@ -64,7 +64,7 @@ setAs("creator", "XMLInternalElementNode",   function(from) S4Toeml(from))
 setAs("person", "creator", function(from)
   as(as(from, "responsibleParty"), "creator"))
 setAs("creator", "person", function(from){
-   p <- as(from, "person")
+   p <- as(as(from, "responsibleParty"), "person")
    p$role = "cre" 
    p
   })
@@ -87,12 +87,25 @@ setAs("contact", "XMLInternalElementNode",   function(from) S4Toeml(from))
 setAs("person", "contact", function(from)
   as(as(from, "responsibleParty"), "contact"))
 setAs("contact", "person", function(from){
-   as(from, "person")
+   as(as(from, "responsibleParty"), "person")
   })
 setAs("character", "contact", function(from)
   as(as.person(from), "contact"))
 ## FIXME Should be something more intelligent that will let these
 ## inherit the method from responsibleParty?  callNextMethod() maybe??
 
+setAs("ListOfresponsibleParty", "person", function(from){
+      l <- lapply(from, as, "person")
+      out <- l[[1]]
+      for(p in l)
+        c(out, p)
+      out
+  })
 
-
+setAs("ListOfcreator", "person", function(from){
+    l <- lapply(from, as, "person")
+      out <- l[[1]]
+      for(p in l)
+        c(out, p)
+      out
+  })

### 2013-09-19 inst/doc/vignette.Rmd
added accessor methods for eml_read, see README
updating approach based on data.set class

--- 2013-09-03 inst/doc/vingette.Rmd
@@
 
-Consider some dataset as an R `data.frame`.
+`reml` now extends R's `data.frame` class by introducing the `data.set` class which includes additional metadata required by EML.
+A `data.set` can be created much like a `data.frame` by specifying additional arguments
 
 ```{r}
-dat = data.frame(river = c("SAC", "SAC", "AM"),
+dat = data.set(river = c("SAC",  "SAC",   "AM"),
                spp   = c("king",  "king", "ccho"),
                stg   = c("smolt", "parr", "smolt"),
-                 ct    =  c(293L, 410L, 210L))
+               ct    = c(293L,    410L,    210L),
+               col.defs = c("River site used for collection",
+                            "Species common name",
+                            "Life Stage", 
+                            "count of live fish in traps"),
+               unit.defs = list(c(SAC = "The Sacramento River", 
+                                  AM = "The American River"),
+                                c(king = "King Salmon", 
+                                  ccho = "Coho Salmon"),
+                                c(parr = "third life stage", 
+                                  smolt = "fourth life stage"),
+                                "number"))
+
 ```
 
-Provide a list giving the  __column header__ used in the data.frame, followed by a plain text definitions for the column, followed by a character vector giving the definition for the units:
-
-__column descriptions__: These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful.
+- `col.defs`: These are usually just plain text definitions, though a URI to a semantic definition can be particularly powerful.
 See "Advanced Use" for details on adding richer information, such as the method used to collect the data or set the geographic, taxanomic, or temporal coverage of an individual column.
 
-__column units__:   For factors, this is a definition of the levels involved.
+- `unit.defs`:   For factors, this is a definition of the levels involved.
 For numeric data, specify the units from [this list](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-unitTypeDefinitions.html#StandardUnitDictionary).
@@
 
-```{r}
-metadata <- 
-  list("river" = list("river",
-                      "River site used for collection",
-                      c(SAC = "The Sacramento River", 
-                        AM = "The American River")),
-       "spp" = list("spp",
-                    "Species common name", 
-                    c(king = "King Salmon", 
-                      ccho = "Coho Salmon")),
-       "stg" = list("stg",
-                    "Life Stage", 
-                    c(parr = "third life stage", 
-                      smolt = "fourth life stage")),
-       "ct"  = list("ct",
-                    "count", 
-                    "number"))
-
-```
-
-*for convenience, this metadata list can instead be constructed with the help of the reml wizard*.
-Just use `metadata <- metadata_wizard(dat)` to begin.
-While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.
+Alternatively, annotations can be added to an existing data frame, `data.set(my.data.frame, col.defs = ..., unit.defs = ...)`.
 
 ```{r}
-eml_write(dat, metadata, title = "reml example",  
+eml_write(dat, 
+          title = "reml example",  
           description = "An example, intended for
@@
 ```
+
+*for convenience, `dat` could simply be a `data.frame` and `reml` will launch it's metadata wizard to assist in constructing the metadata based on the data.frame provided*.
+While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.
 
@@
 
+```{r}
+obj <- eml_read("reml_example.xml")
+obj 
+```
+
+We can use various accessor functions to return the data and metadata elements in native R formats.
+
+```{r}
+dataTable(obj)
+```
+
+```{r}
+metadata <- attributeList(obj)
+```
+
+```{r}
+contact(obj)
+```
+
+Note that the contact has been coerced into R's built-in 'person' object:
+
+```{r}
+class(contact(obj))
+```
+
+```{r}
+citationInfo(obj)
+```
+
 #### In development

### 2013-09-19 R/attribute.R

@@ -2,18 +2,39 @@
 setClass("codeDefinition", 
          representation(code = "character",
                         definition = "character"))
+setAs("codeDefinition", "XMLInternalElementNode",   function(from) emlToS4(from))
+setAs("XMLInternalElementNode", "codeDefinition",  function(from) emlToS4(from))
+
+
 setClass("ListOfcodeDefinition", contains="list")
+
 setClass("enumeratedDomain", 
          representation(codeDefinition = "ListOfcodeDefinition"))
+setAs("enumeratedDomain", "XMLInternalElementNode",  function(from) emlToS4(from))
+setAs("XMLInternalElementNode", "enumeratedDomain", function(from) emlToS4(from))
+
 setClass("textDomain", 
          representation(definition = "character"))
+setAs("textDomain", "XMLInternalElementNode",   function(from) emlToS4(from))
+setAs("XMLInternalElementNode", "textDomain",  function(from) emlToS4(from))
+
 setClass("nonNumericDomain", 
          representation(enumeratedDomain = "enumeratedDomain", 
                         textDomain = "textDomain"))
+setAs("nonNumericDomain", "XMLInternalElementNode",   function(from) emlToS4(from))
+setAs("XMLInternalElementNode", "nonNumericDomain",  function(from) emlToS4(from))
+
+
 setClass("nominal", 
          representation(nonNumericDomain = "nonNumericDomain"))
+setAs("nominal", "XMLInternalElementNode",   function(from) emlToS4(from))
+setAs("XMLInternalElementNode", "nominal",  function(from) emlToS4(from))
+
 
 setClass("ordinal", contains="nominal")
+setAs("ordinal", "XMLInternalElementNode",   function(from) emlToS4(from))
+setAs("XMLInternalElementNode", "ordinal",  function(from) emlToS4(from))
+
 
 
 setClass("bounds", 
@@ -30,23 +51,35 @@ setAs("boundsGroup", "XMLInternalElementNode", function(from) S4Toeml(from))
 setClass("numericDomain", 
          representation(numberType = "character",
                         boundsGroup = "boundsGroup")) 
-         # Alternatively restrict this to one of: "natural", "whole", "integer", "real", though schema should enforce that
+setAs("numericDomain", "XMLInternalElementNode",   function(from) emlToS4(from))
+setAs("XMLInternalElementNode", "numericDomain",  function(from) emlToS4(from))
 
 setClass("unit",
          representation(standardUnit = "character", # make from a restricted subset
[…]

### 2013-11-01 inst/doc/vignette.Rmd
add support for writing older eml versions, e.g. 2.1.0
tweak vignette

@@
 obj <- eml_read("reml_example.xml")
-obj 
 ```
@@
 ```{r}
-dataTable(obj)
+dat <- dataTable(obj)
 ```
@@
 ```r
-install.packages("uuid",,'http://rforge.net/',type='source')
+install.packages("uuid")
 install_github("RHTMLForms", "omegahat")

### 2013-12-12 inst/doc/README.Rmd
updated basic tutorial to reflect current API

--- 2013-11-01 inst/doc/vignette.Rmd
@@
 
+We will also specify a default creator who will also be used as the contact person for EML files created in this session.
+
 ```{r} 
-eml_write(dat, 
-          title = "reml example",  
-          description = "An example, intended for
-                              illustrative purposes only.",
-          creator = "Carl Boettiger <cboettig@gmail.com>",
-          file = "reml_example.xml")
+eml_config(creator="Carl Boettiger <cboettig@gmail.com>")
 ```
 
-*for convenience, `dat` could simply be a `data.frame` and `reml` will launch it's metadata wizard to assist in constructing the metadata based on the data.frame provided*.
+While we will always be able to provide alternative or additional creators or contact person later, `eml_config` can store the defaults to save us the trouble of entering this data repeatedly.
+See `?eml_config` for more information and other configurable defaults.
+
+With this information in place, we have all the required metadata to generate a minimally valid EML file documenting this dataset.
+
+```{r}
+eml_write(dat, file = "reml_example.xml")
+```
+
+*for convenience, had we ommitted any essential metadata, such as providing only an unannotated `data.frame` in place of a `data.set`, `reml` will launch its metadata wizard to assist in constructing the metadata based on the data.frame provided*.
 While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.
@@
 See the [EML generated](https://github.com/ropensci/reml/tree/master/inst/doc/reml_example.xml) by this example.
-Note that if the `metadata` argument providing the definitions for the `data.frame` is not specified, `reml` will launch the wizard to prompt the user.
 
-### Configuration
-
-Configure general metadata you may want to frequently reuse, avoiding having to specify things like the name of the creator or contact.
+We can validate the resulting document to check that no required fields are missing and that no entry has been incorrectly formatted.
+This ensures that other software can correctly parse and interpret the metadata document:
 
 ```{r}
-eml_config(creator = list("Carl Boettiger <cboettig@ropensci.org>", "Karthik Ram"), contact = "Carl Boettiger <cboettig@ropensci.org>")
+eml_validate("reml_example.xml")
 ```
@@
 
+### Publishing to figshare
+
 In this example, we will publish the above
@@
 _Note: this example assumes `rfigshare` is installed, even though it is not 
-required to use reml and run the examples in the other sections_
+required to use reml and run the examples in the other sections.  See the 
+`rfigshare` documentation for authenticating `rfigshare` with your user account_
 
 <!-- We don't want to generate a DOI every time we run the vignette -->
+
 ```{r publish_figshare} 
-id = eml_publish("reml_example.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML", destination="figshare")
+eml_publish("reml_example.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML", destination="figshare")
 ```
@@
 
+### Publishing to the KNB
+
+We can also publish to the KNB (Knowledge Network for Biocomplexity, a data repository provided by NCEAS) using the `dataon` R package.
+KNB offers several advantages over alternative repositories such as figshare or [Dryad](http://datadryad.org) when publishing EML metadata.
+The KNB "speaks" EML, so it is unnecessary to provide repository-specific additional metadata such as description, category or tag -- all necessary information will be extracted from the EML.
+This is also much more thorough than other repositories: we can search the KNB network by specific taxonomic or geographic coverage, or even down at the level of individual attribute or column headings, rather than being limited to more generic metadata.
+Lastly, KNB is a member node of the DataONE repository, making the resulting data files immediately part of a much larger network for ecological, earth and environmental science data.
+
+_Note: this example assumes `dataone` is installed, even though it is not 
+required to use reml and run the examples in the other sections.  See the 
+`dataone` documentation for authenticating `dataone` with a user account_
+
+```{r publish_knb, eval=FALSE}
+eml_publish("reml_example.xml", destination="KNB", permanent=FALSE)
+```
+
+_NB:_ To avoid clutter, we should publish only real research data to the KNB production servers (by setting `permanent=TRUE`).
+By default, `eml_publish` assumes `permanent=FALSE` when sending data to the KNB, and will publish to the development server instead.
+The development server works as a testing platform, providing all the same features as the production platform but without permanent storage -- all data will be deleted regularly.
+The development server is not as stable as the production server, so if the command fails, please try again later.
+
 <!-- In return, figshare provides the object with a DOI, which is added to the EML.  
-
-```{r figshare_doi, eval=FALSE}
-doc <- eml_read("my_eml_data.xml")
-citation(doc) # not implemented yet
-```
 -->
@@
 
-Reading EML
------------
+Parsing and exploring an EML file
+----------------------------------
+
+Users will not only want to generate EML files, but also take advantage of the wealth of data already well documented in EML.
+The package provides many tools to facilitate the extraction of data from EML files into their native R formats, as well as utilities to access and manipulate the metadata provided.
+By having these tools avialble in the scripted R environment, it becomes possible to write functions that can be easily applied across large collections of files.
+In this way, researchers can scale their analyses across ever larger collections of datasets by automating the more tedious aspects of data discovery and integration.
 
@@
 
-We can use various accessor functions to return the data and metadata elements in native R formats.
+We can also read in a remote file by providing a URL or KNB object identifier (such as a DOI).
+
+The `eml_get` function provides us with easy access to many of the component elements of the metadata file.  See the documentation for a complete list.
 
 ```{r}
-dat <- dataTable(obj)
+dat <- eml_get(obj, "data.set")
 ```
@@
 ```{r}
-metadata <- attributeList(obj)
+eml_get(obj, "contact")
 ```
@@
 ```{r}
-contact(obj)
+eml_get(obj, "citation_info")
 ```
 
-Note that the contact has been coerced into R's built-in 'person' object:
+Advanced writing of EML
+-----------------------
 
-```{r}
-class(contact(obj))
-```
+**IN PROGRESS**
 
-```{r}
-citationInfo(obj)
-```
+Our minimal EML file barely scratches the surface of what is possible to do with EML.  In [Advanced writing of EML](), we construct a much richer EML file, including:
 
-#### In development
+- Constructing more complete lists of authors, publishers and contact.
+- Summarizing the geographic, temporal, and taxonomic coverage of the dataset
+- Reading in pages of methods descriptions from a Word document
+- Adding arbitrary additional metadata
+- Indicating the canonical citation to the paper that should be acknowledged when the data is re-used.  
+- Conversion between EML and other metadata formats, such as NCBII and ISO standards.
 
-See [Open issues in reading EML](https://github.com/ropensci/reml/issues?labels=read+eml&page=1&state=open)
+In so doing we will take a more modular approach that will allow us to build up our metadata from reusable components, while also providing a more fine-grained control over the resulting output fields and files.
 
-Integrating multiple EML files into a single data frame
--------------------------------------------------------
+Advanced parsing and manipulation of EML 
+----------------------------------------
 
-#### In development
+**IN PROGRESS**
 
-See [Open issues in integrating EML](https://github.com/ropensci/reml/issues?labels=integrate+eml&page=1&state=open)
+In [Advanced parsing and manipulation of EML](), we
+
+- Introduce how to access any EML element in R using the S4 subsetting mechanism
+- Demonstrate how to extract and manipulate semantic RDF triples from EML metadata
+- Illustrate how EML metadata extraction can be applied across a collection of files 
+- Illustrate how EML manipulation can be combined with queries across even larger distributed collections of EML data over the dataone network
+- Demonstrate merging of multiple datasets based on semantics
 
@@
 Successful installation of these packages will require the `devtools` package (for `install_github`) and the ability to build packages from source.
+
+***Suggestions, feature requests, bug reports and other feedback welcome!*** Please see our [issues tracker](https://github.com/ropensci/reml/issues)

### 2013-12-12 inst/doc/Advanced_writing_of_EML.Rmd
started Advanced writing tutorial, based on hf205.xml

In [Advanced writing of EML](), we construct richer EML files, including:

- Constructing more complete lists of authors, publishers and contact.
- Summarizing the geographic, temporal, and taxonomic coverage of the dataset
- Reading in pages of methods descriptions from a Word document
- Adding arbitrary additional metadata
- Indicating the canonical citation to the paper that should be acknowledged when the data is re-used.  
- Conversion between EML and other metadata formats, such as NCBII and ISO standards.

In so doing we will take a more modular approach that will allow us to build up our metadata from reusable components, while also providing a more fine-grained control over the resulting output fields and files.

### Creating hf205.xml

In this example, we will use R to re-generate the EML metadata originally published by [Ellison _et al_ (2012)]() through the Harvard Forest Long Term Ecological Research Center.  <!-- The original metadata file was created in association with the publication in PNAS based on a Microsoft Word document template that Harvard Forest provides to the academic researchers.  Metadata from this template is then read off by hand and an EML file is generated using a combination of a commercial XML editing platform (Oxygen) for commonly used higher-level elements, and the Java platform `Morpho` provided by the EML development team for lower level attribute metadata. -->

#### Overview of the EML heirarchy

The EML metadata language

We begin by reading in the CSV file providing the raw data that is to be annotated.  A copy of this file is distributed for demonstration purposes as part of the R package.

```{r}
 dat <- read.csv(system.file("examples", "hf205-01-TPexp1.csv", package="reml"))
```

As in our [Basic tutorial](), essential metadata describes the columns and units used in the CSV file.

```{r}
col.defs <- c("run.num" = "which run number (=block). Range: 1 - 6. (integer)",
              "year" = "year, 2012",
              "day" = "Julian day. Range: 170 - 209.",
              "hour.min" = "hour and minute of observation. Range 1 - 2400 (integer)",
              "i.flag" =  "is variable Real, Interpolated or Bad (character/factor)",
              "variable" = "what variable being measured in what treatment (character/factor).",
              "value.i" = "value of measured variable for run.num on year/day/hour.min.")
```

```{r}
unit.defs = list("which run number", 
                 "YYYY",
                 "DDD",
                 "hhmm", 
                 c(R = "real", I = "interpolated", B = "bad"),
                 c(control = "no prey added", 
                   low = "0.125 mg prey added ml-1 d-1",
                   med.low = "0,25 mg prey added ml-1 d-1",    # ERROR 
                   med.high = "0.5 mg prey added ml-1 d-1",
                   high = "1.0 mg prey added ml-1 d-1",
                   air.temp = "air temperature measured just above all plants (1 thermocouple)",
                   water.temp = "water temperature measured within each pitcher",
                   par = "photosynthetic active radiation (PAR) measured just above all plants (1 sensor)"),
                 c(control = "% dissolved oxygen",
                   low = "% dissolved oxygen",
                   med.low = "% dissolved oxygen",
                   med.high = "% dissolved oxygen",
                   high  = "% dissolved oxygen",
                   air.temp = "degrees C",
                   water.temp = "degrees C",
                   par = "micromoles m-1 s-1"))
```

We attach this essential metadata directly to the `data.frame` to make a `data.set` object (a data.frame with basic metadata).

```{r}
dat <- data.set(dat, col.defs = col.defs, unit.defs = unit.defs)
```

Rather than write this out directly to an EML file with `eml_write` like we did in the basic example, we'll first assemble more metadata information.

As a _metadata_ format, we do not encode the actual data file itself into the EML object.  Instead, the data is written out using some existing standard file format, which is merely _described by_ the EML file. The leading recommendation for exchanging and preserving tabular data is to use a plain text csv (comma separated values) file.  The function `eml_dataTable` will write out the data frame to a csv file, linking the filename to the metadata. We can optionally provide other information, such as description and name for the output file (if none is provided the unique id generated for this element will be used).

```{r}
dataTable <- reml:::eml_dataTable(dat, 
                           description = "Metadata documentation for S1.csv", 
                           file = "S1.csv")
```

CSV file conventions can differ across platforms and continents. To remove any ambiguity, EML explicitly describes how the csv file is constructed (noting that it is actually comma delimited as the name implies and not tab delimited, and noting other details such as the number of rows, the character used to indicate and end of a row, and so forth).  Because R is writing the csv file itself after all, it already knows all of these details and we are able to automatically fill out this section of the metadata.  This information is summarized in the `dataTable@physical` element.

One of the primary benefits in constructing EML in a scripted language such as R rather than in a GUI interface such as `morpho` is the ability to easily save and reuse components.  Here is an EML component that Harvard Forest will use a lot: it's address

```{r}
HF_address <- new("address", 
                  deliveryPoint = "324 North Main Street",
                  city = "Petersham",
                  administrativeArea = "MA",
                  postalCode = "01366",
                  country = "USA") 
```

We use the `new` function to construct an object of class `address`.  This R object has "slots" corresponding to each of the elements that makes up an address in EML.  We can define the values for these slots all at once as shown, or we can add and edit them after the object is created by using R's S4 subsetting operator, `@`.  For instance, we could change the street address with: `HF_address@deliveryPoint <- "some new address"`. (Some users might find the equivalent notation, `slot(HF_address, "deliveryPoint") <- "some new address"` to be more intuitive).  We need not specify any additional values when first calling `new("address")`, and we can use `slotNames(HF_address)` to see a list of possible fields.

This element is now available for reuse as we create other metadata.  For instance, we will define the publisher of the data by providing a name and an address:

```{r}
publisher <- new("publisher", 
    organizationName = "Harvard Forest",
    address = HF_address)
```

Note that the `address` slot is not simply a text field, but is in fact an `address` S4 object we created above.  This object preserves the structure defined by the EML schema, which ensures that software programs can consistently interpret the values we write into the metadata file.

We can reuse data formats native to R as well.  For instance, R comes with an inherent notion of a "person" as something with a given name, a family name, an email address, and so forth.  R recognizes a string written like this:

```{r}
aaron <- as.person("Aaron Ellison <fakeaddress@email.com>")
```

can be interpreted as person's first and last name and email address.  The EML R package also recognizes this syntax, and can coerce this into the various 'person' or 'responsibleParty' roles defined in the EML schema.  Here we construct a contact node using coercion (the `as` function) from the `person` object

```{r}
contact <- as(aaron, "contact")
contact@address = HF_address
contact@organizationName = "Harvard Forest"
contact@phone = "000-000-0000"
```

We could have entered each slot manually using `new("contact")` as we did for the above elemnts.

Multiple individuals can be listed as the creator (author) of data.  Here we use coercion (`as`) to turn strings into creator nodes, and then use R's concatenate operator, `c`, to combine them.

```{r}
creator <- c(as("Aaron Ellison", "creator"), as("Nicholas Gotelli", "creator"))
```

Yet another way to create person objects is to use the helper function `eml_person`.  This function recongnizes the plain-text string conventions already used by R's person objects (or the person objects themselves, see ?person).   Here we add other researchers as contributors to the data.  The `eml_person` function automatically decides whether to return a `contact`, `creator` list, or `associatedParty`, based on the additional information provided (such as an email address in angle braces, contributor role in square brackets).

```{r}
other_researchers <- eml_person("Benjamin Baiser [ctb]", "Jennifer Sirota [ctb]") # creates a 'ListOfassociatedParty' (roles != cre)
```

Again, we could just as well have created these nodes using any of the other mechanisms shown above.

Some metadata is just a string:

```{r}
pubDate <- "2012" 
title <- "Thresholds and Tipping Points in a Sarracenia Microecosystem at Harvard Forest since 2012"
```

When metadata fields have more complicated types, the package provides constructor functions to help build these elements succinctly.  For instance, while we could add our keyword sets manually using only `new`:

```{r eval=FALSE}
keys <-
  c(new("keywordSet", 
      keywordThesaurus = "LTER controlled vocabulary",
      keyword = c(new("keyword", keyword="bacteria"), new("keyword", keyword="carnivorous plants"), ...)
    ),
   new("keywordSet", 
      keywordThesaurus = "LTER core area",
      keyword = ...) 
```

and so forth, it is more compact to use the provided constructor function:

```{r}
keys <- eml_keyword(list("LTER controlled vocabulary" = c("bacteria", "carnivorous plants", "genetics", "thresholds"),
                                     "LTER core area" = c("populations", "inorganic nutrients", "disturbance"),
                                        "HFR default" = c("Harvard Forest", "HFR", "LTER", "USA")))
```

A similar situation arises with describing the coverage metadata element. 
```{r}
coverage <- eml_coverage(scientific_names = "Sarracenia purpurea", 
                         dates            = c('2012-06-01', '2013-12-31'),
                         geographic_description = "Harvard Forest Greenhouse, Tom Swamp Tract (Harvard Forest)", 
                         NSEWbox          = c( 42.55,  42.42, -72.1, -72.29, 160, 330))
```

In this case, the helper function supports only the more common types of coverage, such as listing taxonomic coverage at the species name rather than higher taxonomic level, or providing only calendar dates instead of, say, geological epochs.  For these alternative constructions, users will call the `new` constructors directly, see `?temporalCoverage`, `?taxonomicCoverage` and related functions for details.

The remaining metadata provided in hf205.xml consists primarily of text blocks.  We might write small text blocks directly into our script:

```{r}
abstract <- "The primary goal of this project is to determine
  experimentally the amount of lead time required to prevent a state
  change. To achieve this goal, we will (1) experimentally induce state
  changes in a natural aquatic ecosystem - the Sarracenia microecosystem;
  (2) use proteomic analysis to identify potential indicators of states
  and state changes; and (3) test whether we can forestall state changes
  by experimentally intervening in the system. This work uses state-of-the
  art molecular tools to identify early warning indicators in the field
  of aerobic to anaerobic state changes driven by nutrient enrichment
  in an aquatic ecosystem. The study tests two general hypotheses: (1)
  proteomic biomarkers can function as reliable indicators of impending
  state changes and may give early warning before increasing variances
  and statistical flickering of monitored variables; and (2) well-timed
  intervention based on proteomic biomarkers can avert future state changes
  in ecological systems."  
```

or read them in from some externa plain text file such as `abstract <- readLines("abstract.txt")`.

```{r}
rights <- "This dataset is released to the public and may be freely
  downloaded. Please keep the designated Contact person informed of any
  plans to use the dataset. Consultation or collaboration with the original
  investigators is strongly encouraged. Publications and data products
  that make use of the dataset must include proper acknowledgement. For
  more information on LTER Network data access and use policies, please
  see: http://www.lternet.edu/data/netpolicy.html."
```

For larger blocks of text we might rather write these Word.  We can read them in from there, with the help of a few additional R packages.

```{r}
library(RWordXML)
library(XML)
f2 <- wordDoc(system.file("examples", "methods.docx", package="reml"))
doc <- f2[[getDocument(f2)]]
txt <- xpathSApply(doc, "//w:t", xmlValue)
## FIXME add <title> <section> and <para> blocking back: 
method_description <- paste(txt, collapse = "\n\n") 
```

We can then construct a methods node containing this text as the description:

```{r}
methods <- new("methods")
methods@methodsStep@description <- method_description
```

One of the simplest ways and most powerful ways to generate EML metadata is to resuse a section from an existing EML file directly.

```{r}
additionalMetadata <- hf205@additionalMetadata # extracted from previous eml file
```

```{r}
dataset <- new("dataset", 
                title = title,
                creator = creator,
                contact = contact,
                pubDate = pubDate,
                associatedParty = other_researchers,
                keywordSet = keywords,
                coverage = coverage,
                methods = methods,
                dataTable = c(dataTable))

```

```{r}
eml <- new("eml", 
            dataset = dataset,
            additionalMetadata = additionalMetadata)
```

This `eml` object we have just created now has the complete R representation of the EML data.  This is the same class of object returned by `eml_read`, and consequently we can use all of the read methods provided to explore or manipulate it.

Finally we are ready to write out our EML object to an XML file.

```{r}
eml_write(eml, "hf205_from_reml.xml")
```

When we used the `eml_write()` function for the first time in the [Basic tutorial](), we simply passed a `data.set` object.  Internally the function made the remaining calls to `new("dataTable")`, `new("dataset")`, and `new("eml")` for us.  This time, the `eml_write` function can recognize that this is already an `eml` object, which it immediately converts to the XML representation and writes out as a file.  (If no file name is provided, the XML text will be returned to the console).

As before, we can now validate our EML document to ensure all the information has been formatted correctly:

```{r}
eml_validate("hf205_from_reml.xml")
```

While it is difficult to create invalid EML using the constructor functions, it is not impossible so it never hurts to check.  As illustrated in the basic tutorial, we can now publish this file automatically to a central repository such as the KNB.

### 2013-12-13 R/dataTable_methods.R
tweak default field values after simplifying eml_write workflow, tests pass with validation

@@ -9,6 +9,8 @@ eml_dataTable <- function(dat,
                           description = character(0), 
                           filename = character(0), ...){
 
+## FIXME title should be called entityName, or maybe just "name"
+
   if(is(dat, "data.set")) 
     meta <- get_metadata(dat)
   if(is.null(meta))
@@ -19,7 +21,9 @@ eml_dataTable <- function(dat,
     filename <- paste0(gsub(" ", "_", title), ".csv")
   id <- reml_id()
   if(length(filename) == 0)
-    filename <- paste0(id,".csv")
+    filename <- paste0(id[["id"]], ".csv")
+  if(length(title) == 0)
+    title <- filename 
 
 
   meta <- detect_class(dat, meta)

### 2013-12-13 inst/doc/Advanced_writing_of_EML.Rmd

@@
 
+```{r include=FALSE, cache=FALSE}
+source("~/.knitr_defaults.R")
+```
+
+### Overview of the EML heirarchy
+
+A basic knowledge of the components of an EML metadata file is essential to being able to take full advantage of the language.
+While more complete information can be found in the official schema documentation, here we provide a general overview of commonly used metadata elements most relevant to describing data tables.
+
+This schematic shows each of the metadata elements we will generate.
+Most these elements have sub-components (e.g.
+a 'publisher' may have a name, address, and so forth) which are not shown for simplicity.
+Other optional fields we will not be generating in this example are also not shown.
+
+```yaml
+- eml
+  - dataset
+    - creator
+    - contact
+    - publisher
+    - title
+    - pubDate
+    - keywords
+    - abstract 
+    - intellectualRights
+    - methods
+    - coverage
+    - dataTable
+      - physical
+      - attributeList
+  - additionalMetadata
+```
+
+The EML R package allows us to assemble these pieces bit by bit, allowing us to work either from the top down or bottom up.
+
+Our `data.set` object captures all of the information required to generate the `dataTable` element: the `data.frame`, which will be written to an external CSV file descibed by the `physical` node, and the `attributeList`, containing the column and unit metadata descriptions.
+
 ### Creating hf205.xml
@@
 
-#### Overview of the EML heirarchy
-
-The EML metadata language
+```{r}
+library(reml)
+```
 
@@
 ```{r}
+hf205 <- eml_read(system.file("examples", "hf205.xml", package="reml"))
 additionalMetadata <- hf205@additionalMetadata # extracted from previous eml file
 ```
+
+We now have all of the elements we wished to define for the `dataset` metadata. We can assemble them into a `datset` node as follows:
 
@@
                 associatedParty = other_researchers,
-                keywordSet = keywords,
+                keywordSet = keys,
                 coverage = coverage,
@@
 ```
+
+Note that because a `dataset` can have multiple dataTable elements, we use the same concatenate notation illustrated earlier, `c(dataTable)`, even though only one element is provided.
+This lets R known that the resulting object is a `ListOfdataTable`, not an individual `dataTable` file.
+This convention should be followed whenever assigning a value to a slot that can take more than one element of that name.
+
+With most of the metadata now in place, we can construct the top-level element:
 
@@
 
+This call automatically adds a few additional bits of information, such as a unique identifier `packageId` for the metadata file.
 This `eml` object we have just created now has the complete R representation of the EML data.
@@
 While it is difficult to create invalid EML using the constructor functions, it is not impossible so it never hurts to check.
+
 As illustrated in the basic tutorial, we can now publish this file automatically to a central repository such as the KNB.

### 2013-12-15 inst/doc/Advanced_writing_of_EML.Rmd

@@
 ```{r}
-dat <- read.csv(system.file("examples", "hf205-01-TPexp1.csv", package="reml"))
+dat <- read.csv(system.file("examples", "hf205-01-TPexp1.csv", package="reml"), colClasses = 
+                  c("factor", "Date", "Date", "Date", "factor", "factor", "factor"))
 ```
@@
 ```{r}
-eml_write(eml, "hf205_from_reml.xml")
+eml_write(eml, file="hf205_from_reml.xml")
 ```

### 2014-01-02 inst/doc/Advanced_parsing_of_EML.Rmd
advanced parsing tutorial draft

In [Advanced parsing and manipulation of EML](https://github.com/ropensci/reml/blob/master/inst/doc/Advanced_parsing_of_EML.md), we

- Introduce how to access any EML element in R using the S4 subsetting mechanism
- Illustrate how EML metadata extraction can be applied across a collection of files 
- Illustrate how EML manipulation can be combined with queries across even larger distributed collections of EML data over the dataone network
- Demonstrate how to extract and manipulate semantic RDF triples from EML metadata
- Demonstrate merging of multiple datasets based on semantics

EML (the Ecological Metadata Language) is formatted using XML (eXtensible Markup Language) acording to a specified XML Schema.  XML uses a heirarchical or treed structure to represent information.  Different types of information are identified by name-value pairs.  Nodes specifiy names such as `<title>` or `<creator>` in angle brackets.  The value is enclosed between opening and closing elements, such as `<title>This is the title</title>`.  The heirarichal structure indicates components of the element, such as:

```xml
<individualName>
  <givenName>John</givenName>
  <surName>Smith</surName>
</individualName>
```

Nodes can also specify information as attributes, such as `<individualName id=45>`, indicating the node has an `id` with the value `45`.  In EML attributes are rarely used for information other than identifiers on the nodes.

The EML package represents nodes defined in the EML schema as R objects using R's S4 object mechanism. We can create a new object corresponding to any node defined in the EML Schema using the `new()` function; such as:

```{r}
node <- new("individualName")
node
```

Showing our object in the console reveals that it has three slots, `salutation`, `givenName`, and `surName`, each currently with the value of a length zero character string.  We can assign values to the slots using the S4 subsetting operator `@` (most R users may be more familiar with subsetting S3 objects or lists by using the `$` operator)

```{r}
node@givenName <- John
```

Alternatively, we can use the `slot` function:

```{r}
slot(node, "givenName") <- John
```

Each of these operations can also be used (without the assignment operator `<-`) to access the value of the slot, e.g.

```{r}
node@givenName
```

### 2014-01-04 R/eml.R
updated documentation, test suite passing

@@ -5,6 +5,8 @@
 #' @include protocol.R
 #' @include reml_id.R
 
+
+#' @import methods
 setClass("metadata", contains="XMLInternalElementNode")
 setAs("metadata", "XMLInternalElementNode", function(from) newXMLNode("metadata", from))
 setAs("XMLInternalElementNode", "metadata", function(from) new("metadata", from))

### 2014-01-04 R/literature.R

@@ -824,7 +824,7 @@ setAs("bibentry", "map", function(from){
 setClass("audio_visual_slots",
          slots = c(publisher = "publisher",
                    publicationPlace = "character",
-                   performer = "listOfperformer",
+                   performer = "ListOfperformer",
                    ISBN = "character")
          )
 

### 2014-01-04 R/reml_id.R

@@ -1,5 +1,7 @@
+
+#' @import uuid 
 reml_id <- function(prefix="reml_") {
-  success <- require(uuid)
+  success <- 1 # require(uuid)  ## uuid on CRAN, let's just import it and be done with
     if(success){
       id <- paste0("urn:uuid:", uuid::UUIDgenerate())
       scope <- "system"

### 2014-01-05 inst/doc/Advanced_writing_of_EML.Rmd
troubleshooting various minor bugs arising from recent changes not caught earlier. Avoid warnings from collate list, validation errors from missing contact when only creator config provided, tweaks to publish_knb().  Still need some tweaks to eml_read to handle identifiers directly.  Still testing on tutorials and vignettes.

@@
 ```{r}
-dat <- read.csv(system.file("examples", "hf205-01-TPexp1.csv", package="reml"), colClasses = 
-                  c("factor", "Date", "Date", "Date", "factor", "factor", "factor"))
+f <- system.file("examples", "hf205-01-TPexp1.csv", package="reml")
+dat <- read.csv(f, 
+                colClasses = "factor", # c("factor", "Date", "Date", "Date", 
+                             #  "factor", "factor", "factor")
+                )
 ```
@@
 Here we add other researchers as contributors to the data.
-The `eml_person` function automatically decides whether to return a `contact`, `creator` list, or `associatedParty`, based on the additional information provided (such as an email address in angle braces, contributor role in square brackets).
+The `eml_person` function automatically decides whether to return a `contact`, `creator` list, or `associatedParty`, based on the additional information provided (such as an email address in angle braces, contributor role, `ctb` in square brackets).
 
 ```{r}
-other_researchers <- eml_person("Benjamin Baiser [ctb]", "Jennifer Sirota [ctb]") # creates a 'ListOfassociatedParty' (roles != cre)
+other_researchers <- eml_person("Benjamin Baiser [ctb]", 
+                                "Jennifer Sirota [ctb]") 
 ```
@@
 pubDate <- "2012" 
-title <- "Thresholds and Tipping Points in a Sarracenia Microecosystem at Harvard Forest since 2012"
+title <- "Thresholds and Tipping Points in a Sarracenia 
+          Microecosystem at Harvard Forest since 2012"
 ```
@@
       keywordThesaurus = "LTER controlled vocabulary",
-      keyword = c(new("keyword", keyword="bacteria"), new("keyword", keyword="carnivorous plants"), ...)
+      keyword = c(new("keyword", keyword="bacteria"), 
+                  new("keyword", keyword="carnivorous plants"), 
+                  ...)
     ),
@@
 ```{r}
-keys <- eml_keyword(list("LTER controlled vocabulary" = c("bacteria", "carnivorous plants", "genetics", "thresholds"),
-                                     "LTER core area" = c("populations", "inorganic nutrients", "disturbance"),
-                                        "HFR default" = c("Harvard Forest", "HFR", "LTER", "USA")))
+keys <- eml_keyword(list(
+ "LTER controlled vocabulary" = c("bacteria", 
+                                  "carnivorous plants", 
+                                  "genetics", 
+                                  "thresholds"),
+             "LTER core area" = c("populations", 
+                                  "inorganic nutrients", 
+                                  "disturbance"),
+                "HFR default" = c("Harvard Forest", 
+                                  "HFR", 
+                                  "LTER", 
+                                  "USA")))
 ```
@@
 ```{r}
-coverage <- eml_coverage(scientific_names = "Sarracenia purpurea", 
+coverage <- eml_coverage(
+  scientific_names = "Sarracenia purpurea", 
   dates            = c('2012-06-01', '2013-12-31'),
-                         geographic_description = "Harvard Forest Greenhouse, Tom Swamp Tract (Harvard Forest)", 
+  geographic_description = "Harvard Forest Greenhouse, 
+                            Tom Swamp Tract (Harvard Forest)", 
   NSEWbox          = c( 42.55,  42.42, -72.1, -72.29, 160, 330))

### 2014-01-05 inst/doc/README.Rmd
tweak default field values after simplifying eml_write workflow, tests pass with validation
troubleshooting various minor bugs arising from recent changes not caught earlier. Avoid warnings from collate list, validation errors from missing contact when only creator config provided, tweaks to publish_knb().  Still need some tweaks to eml_read to handle identifiers directly.  Still testing on tutorials and vignettes.

@@
 source("~/.knitr_defaults.R")
-opts_chunk$set(cache=FALSE)
 ```
@@
 ```{r publish_knb, eval=FALSE}
-eml_publish("reml_example.xml", destination="KNB", permanent=FALSE)
+ids <- eml_publish("reml_example.xml", destination="KNB", permanent=FALSE)
 ```
@@
 The development server is not as stable as the production server, so if the command fails, please try again later.
+
+<!--
+Test data shows up on the test server, followed by the identifer, at `paste0("(https://mn-demo-5.test.dataone.org/knb/d1/mn/v1/object/", ids[["eml"]])`
+
+This returns the object identifiers for the EML metadata, the csv file, and the package itself.
+We can read in the EML directly from its unique identifier using `eml_read`.
+We must wait about four minutes first for the entry to be completely registered in the dataone repository.
+(Note that because we are using the development server, `permanent=FALSE`, this file will dissapear after a few days.)
+
+```{r}
+Sys.sleep(60*4)
+eml <- eml_read(paste0("(https://mn-demo-5.test.dataone.org/knb/d1/mn/v1/object/", ids[["eml"]]))
+```
+-->
 

### 2014-01-05 R/eml_read.R

@@ -23,6 +23,8 @@ eml_read <- function(file,  ...){
   if(is(file, "XMLInternalDocument")){  # if already parsed XML, e.g. from d1_get
     doc <- file
 
+  
+
   ## Local Path
   } else if(file.exists(file)){ # if is a local file path...
     doc <- xmlParse(file=file, ...)
@@ -35,7 +37,9 @@ eml_read <- function(file,  ...){
         doc <- content(GET(file), "parsed", "text/xml") 
       
       ## Identifier 
-      } else {   ## Assume object is a DataONE identifer(?)
+      } else {   ## FIXME cannot assume this endpoint. Better to use dataone function  
+#        getD1Object(cli, file)
+#        use  dataone::EMLParser -- but returns XMLNode instead of an XMLInternalNode (C pointer...) 
         doc <- content(GET(paste0("https://cn.dataone.org/cn/v1/object/", file)), "parsed", "text/xml") # content type actually is detected.
       }
   } else {

### 2014-01-16 inst/doc/README.Rmd
editing constructor for eml, avoid defining ID in eml prototype as the id is then not unique byt re-used through-out the session
Fixes #88 upstream (inst/doc/README.Rmd)

@@
 
-This creates a draft file visible only to the user configured in `rfigshare`.
-The document can be made (permanently) public using either the figshare web interface, the `rfigshare` package (using `fs_make_public(id)`) or just by adding the argument `visibility = TRUE` to the above `eml_publish` command.
+This creates a draft file visible only to the user configured in
+`rfigshare`.  The document can be made (permanently) public using
+either the figshare web interface, the `rfigshare` package (using
+`fs_make_public(id)`) or just by adding the argument `visibility = TRUE`
+to the above `eml_publish` command.
 
@@
 
-We can also publish to the KNB (Knowledge Network for Biocomplexity, a data repository provided by NCEAS) using the `dataon` R package.
-KNB offers several advantages over alternative repositories such as figshare or [Dryad](http://datadryad.org) when publishing EML metadata.
-The KNB "speaks" EML, so it is unnecessary to provide repository-specific additional metadata such as description, category or tag -- all necessary information will be extracted from the EML.
-This is also much more thorough than other repositories: we can search the KNB network by specific taxonomic or geographic coverage, or even down at the level of individual attribute or column headings, rather than being limited to more generic metadata.
-Lastly, KNB is a member node of the DataONE repository, making the resulting data files immediately part of a much larger network for ecological, earth and environmental science data.
+We can also publish to the KNB (Knowledge Network for Biocomplexity,
+a data repository provided by NCEAS) using the `dataone` R package.
+KNB offers several advantages over alternative repositories such
+as figshare or [Dryad](http://datadryad.org) when publishing EML
+metadata. The KNB "speaks" EML, so it is unnecessary to provide
+repository-specific additional metadata such as description, category or
+tag -- all necessary information will be extracted from the EML. This is
+also much more thorough than other repositories: we can search the KNB
+network by specific taxonomic or geographic coverage, or even down at
+the level of individual attribute or column headings, rather than being
+limited to more generic metadata.  Lastly, KNB is a member node of the
+DataONE repository, making the resulting data files immediately part of a
+much larger network for ecological, earth and environmental science data.
 
@@
 
-```{r publish_knb, eval=FALSE}
+```{r publish_knb}
 ids <- eml_publish("reml_example.xml", destination="KNB", permanent=FALSE)
@@
 
-_NB:_ To avoid clutter, we should publish only real research data to the KNB production servers (by setting `permanent=TRUE`).
-By default, `eml_publish` assumes `permanent=FALSE` when sending data to the KNB, and will publish to the development server instead.
-The development server works as a testing platform, providing all the same features as the production platform but without permanent storage -- all data will be deleted regularly.
-The development server is not as stable as the production server, so if the command fails, please try again later.
+_NB:_ To avoid clutter, we should publish only real research data to
+the KNB production servers (by setting `permanent=TRUE`).  By default,
+`eml_publish` assumes `permanent=FALSE` when sending data to the KNB,
+and will publish to the development server instead.  The development
+server works as a testing platform, providing all the same features as
+the production platform but without permanent storage -- all data will
+be deleted regularly.  The development server is not as stable as the
+production server, so if the command fails, please try again later.
 
@@
 
-<-- Too detailed for this vignette
+<--! Too detailed for this vignette
 
@@
 
-Users will not only want to generate EML files, but also take advantage of the wealth of data already well documented in EML.
-The package provides many tools to facilitate the extraction of data from EML files into their native R formats, as well as utilities to access and manipulate the metadata provided.
-By having these tools avialble in the scripted R environment, it becomes possible to write functions that can be easily applied across large collections of files.
-In this way, researchers can scale their analyses across ever larger collections of datasets by automating the more tedious aspects of data discovery and integration.
+Users will not only want to generate EML files, but also take advantage of
+the wealth of data already well documented in EML.  The package provides
+many tools to facilitate the extraction of data from EML files into
+their native R formats, as well as utilities to access and manipulate
+the metadata provided. By having these tools avialble in the scripted R
+environment, it becomes possible to write functions that can be easily
+applied across large collections of files. In this way, researchers
+can scale their analyses across ever larger collections of datasets by
+automating the more tedious aspects of data discovery and integration.
 

### 2014-01-16 inst/doc/README.Rmd
Updates readme
o Adds me as author to the rmd file
o builds the readme

--- 2014-01-05 inst/doc/README.Rmd
@@
 * Maintainer: Carl Boettiger
-* Authors: Carl Boettiger, Duncan Temple Lang, Karthik Ram, Matt Jones
+* Authors: Carl Boettiger, Claas-Thido Pfaff, Duncan Temple Lang, Karthik Ram, Matt Jones
 * License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
@@
 _This is a dynamic document, generated by [knitr](http://yihui.name/knitr) using [vignette.Rmd](https://github.com/ropensci/reml/blob/master/inst/doc/vingette.Rmd) as the source._
-
-```{r include=FALSE, cache=FALSE}
-source("~/.knitr_defaults.R")
-```
 

### 2014-01-20 inst/doc/Advanced_writing_of_EML.Rmd
update tutorial

@@
 A similar situation arises with describing the coverage metadata element.
+While we could call `new("coverage")` and build the element up slot by slot, most users will find the helper function more convenient:
+
 ```{r}
@@
 ## FIXME add <title> <section> and <para> blocking back: 
-method_description <- paste(txt, collapse = "\n\n") 
+method <- paste(txt, collapse = "\n\n") 
 ```
@@
 ```{r}
-methods <- new("methods")
-methods@methodsStep@description <- method_description
+methods <- new("methods", methodStep = c(new("methodStep", description = method)))
 ```
@@
 
-We now have all of the elements we wished to define for the `dataset` metadata. We can assemble them into a `datset` node as follows:
+We now have all of the elements we wished to define for the `dataset` metadata.
+A `dataset` can include yet more elements than we show here (see a list with `slotNames(new("dataset"))`, or see `?dataset` or the [EML documentation](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-dataset.html#DatasetType) for details).
+We can assemble them into a `dataset` node as follows:
 
@@
                 pubDate = pubDate,
+                intellectualRights = rights,
+                abstract = abstract,
                 associatedParty = other_researchers,
@@
 eml <- new("eml",
+            packageId = uuid::UUIDgenerate(),
+            system = "uuid", # type of identifier
             dataset = dataset,
@@
 
-This call automatically adds a few additional bits of information, such as a unique identifier `packageId` for the metadata file.
 This `eml` object we have just created now has the complete R representation of the EML data.
 This is the same class of object returned by `eml_read`, and consequently we can use all of the read methods provided to explore or manipulate it.
+
+## `eml` helper function
+
+The helper function `eml` provides a shortcut to this step-by-step construction while retaining most of the expressiveness.
+
+```{r}
+eml     <- eml( dataset = dat,
+                title = title,
+                creator = creator,
+                contact = contact,
+                pubDate = pubDate,
+                associatedParty = other_researchers,
+                intellectualRights = rights,
+                abstract = abstract,
+                keywordSet = keys,
+                coverage = coverage,
+                methods = method,
+                additionalMetadata = additionalMetadata
+              )
+```
+
+The helper function handles certain steps automatically, such as the creation of a `uuid` for the package identifier.
+The required fields `creator` and `contact` will use values set by `eml_config` or can take text strings or R `person` objects as arguments, as well as eml's `creator` or `contact` classes.
+In this case we pass a `dataTable` (permitting additional customization we used above such as the description field).
+Alternatively this could be the more basic `data.set` object as the first argument (`dataset = dat`) or an already constructed EML `dataset` class.
+
+Because package identifer, creator and contact fields are required for valid EML, this constructor function is often more convenient then the stepwise process shown above.
+Typically a single call to this function can provide all the metadata needed to properly annotate a `data.set`.
+This is similar to the behavior of the `eml_write` function, but returns the R eml object rather than writing out to an XML file.
+Additional fields can then be added or manipulated using the standard R S4 subsetting methods (see Advanced Parsing tutorial).
 
@@
 When we used the `eml_write()` function for the first time in the [Basic tutorial](), we simply passed a `data.set` object.
-Internally the function made the remaining calls to `new("dataTable")`, `new("dataset")`, and `new("eml")` for us.
+Internally the function made the remaining calls to `new("dataTable")`, `new("dataset")`, and `eml()` for us.
 This time, the `eml_write` function can recognize that this is already an `eml` object, which it immediately converts to the XML representation and writes out as a file.

### 2014-01-20 inst/doc/README.Rmd
updated README

--- 2014-01-16 inst/doc/README.Rmd
@@
 
-### Publishing to the KNB
-
-We can also publish to the KNB (Knowledge Network for Biocomplexity,
-a data repository provided by NCEAS) using the `dataone` R package.
-KNB offers several advantages over alternative repositories such
-as figshare or [Dryad](http://datadryad.org) when publishing EML
-metadata. The KNB "speaks" EML, so it is unnecessary to provide
-repository-specific additional metadata such as description, category or
-tag -- all necessary information will be extracted from the EML. This is
-also much more thorough than other repositories: we can search the KNB
-network by specific taxonomic or geographic coverage, or even down at
-the level of individual attribute or column headings, rather than being
-limited to more generic metadata.  Lastly, KNB is a member node of the
-DataONE repository, making the resulting data files immediately part of a
-much larger network for ecological, earth and environmental science data.
-
-_Note: this example assumes `dataone` is installed, even though it is not 
-required to use reml and run the examples in the other sections.  See the 
-`dataone` documentation for authenticating `dataone` with a user account_
-
-```{r publish_knb}
-ids <- eml_publish("reml_example.xml", destination="KNB", permanent=FALSE)
-```
-
-_NB:_ To avoid clutter, we should publish only real research data to
-the KNB production servers (by setting `permanent=TRUE`).  By default,
-`eml_publish` assumes `permanent=FALSE` when sending data to the KNB,
-and will publish to the development server instead.  The development
-server works as a testing platform, providing all the same features as
-the production platform but without permanent storage -- all data will
-be deleted regularly.  The development server is not as stable as the
-production server, so if the command fails, please try again later.
-
-<!--
-Test data shows up on the test server, followed by the identifer, at `paste0("(https://mn-demo-5.test.dataone.org/knb/d1/mn/v1/object/", ids[["eml"]])`
-
-This returns the object identifiers for the EML metadata, the csv file, and the package itself.
-We can read in the EML directly from its unique identifier using `eml_read`.
-We must wait about four minutes first for the entry to be completely registered in the dataone repository.
-(Note that because we are using the development server, `permanent=FALSE`, this file will dissapear after a few days.)
-
-```{r}
-Sys.sleep(60*4)
-eml <- eml_read(paste0("(https://mn-demo-5.test.dataone.org/knb/d1/mn/v1/object/", ids[["eml"]]))
-```
--->
-
-<!-- In return, figshare provides the object with a DOI, which is added to the EML.  
--->
-
-<--! Too detailed for this vignette
-
-For the sake of completeness, the explicitly figshare metadata is also written into a 
-dedicated `<additionalMetadata>` node with the attribute `id = figshare`.  This allows
-us to distinguish between metadata that is an is not available to figshare's indexing,
-e.g. if we use the figshare search tool.
-
--->
-
 #### In development

### 2014-01-20 inst/doc/knb.Rmd
knb tutorial separately (since it must handle authentication)

### Publishing to the KNB

We can also publish to the KNB (Knowledge Network for Biocomplexity,
a data repository provided by NCEAS) using the `dataone` R package.
KNB offers several advantages over alternative repositories such
as figshare or [Dryad](http://datadryad.org) when publishing EML
metadata. The KNB "speaks" EML, so it is unnecessary to provide
repository-specific additional metadata such as description, category or
tag -- all necessary information will be extracted from the EML. This is
also much more thorough than other repositories: we can search the KNB
network by specific taxonomic or geographic coverage, or even down at
the level of individual attribute or column headings, rather than being
limited to more generic metadata.  Lastly, KNB is a member node of the
DataONE repository, making the resulting data files immediately part of a
much larger network for ecological, earth and environmental science data.

_Note: this example assumes `dataone` is installed, even though it is not 
required to use reml and run the examples in the other sections.  See the 
`dataone` documentation for authenticating `dataone` with a user account_

```{r publish_knb}
ids <- eml_publish("reml_example.xml", destination="KNB", permanent=FALSE)
```

_NB:_ To avoid clutter, we should publish only real research data to
the KNB production servers (by setting `permanent=TRUE`).  By default,
`eml_publish` assumes `permanent=FALSE` when sending data to the KNB,
and will publish to the development server instead.  The development
server works as a testing platform, providing all the same features as
the production platform but without permanent storage -- all data will
be deleted regularly.  The development server is not as stable as the
production server, so if the command fails, please try again later.

<!--
Test data shows up on the test server, followed by the identifer, at `paste0("(https://mn-demo-5.test.dataone.org/knb/d1/mn/v1/object/", ids[["eml"]])`

This returns the object identifiers for the EML metadata, the csv file, and the package itself.  We can read in the EML directly from its unique identifier using `eml_read`. We must wait about four minutes first for the entry to be completely registered in the dataone repository.  (Note that because we are using the development server, `permanent=FALSE`, this file will dissapear after a few days.)

```{r}
Sys.sleep(60*4)
eml <- eml_read(paste0("(https://mn-demo-5.test.dataone.org/knb/d1/mn/v1/object/", ids[["eml"]]))
```
-->

<!-- In return, figshare provides the object with a DOI, which is added to the EML.  
-->

<--! Too detailed for this vignette

For the sake of completeness, the explicitly figshare metadata is also written into a 
dedicated `<additionalMetadata>` node with the attribute `id = figshare`.  This allows
us to distinguish between metadata that is an is not available to figshare's indexing,
e.g. if we use the figshare search tool.

-->

### 2014-01-21 inst/doc/Advanced_parsing_of_EML.Rmd
update tutorial
rename package as EML instead of reml, closes #63

@@
-In [Advanced parsing and manipulation of EML](https://github.com/ropensci/reml/blob/master/inst/doc/Advanced_parsing_of_EML.md), we
+```{r include=FALSE, cache=FALSE}
+source("~/.knitr_defaults.R")
+```
+
+In [Advanced parsing and manipulation of EML](https://github.com/ropensci/EML/blob/master/inst/doc/Advanced_parsing_of_EML.md), we
 
@@
 
+```{r}
+library(EML)
+```
+
 The EML package represents nodes defined in the EML schema as R objects using R's S4 object mechanism.
@@
 ```{r}
-node@givenName <- John
+node@givenName <- "John"
 ```
@@
 ```{r}
-slot(node, "givenName") <- John
+slot(node, "givenName") <- "John"
 ```

### 2014-01-21 inst/doc/README.Rmd
rename package as EML instead of reml, closes #63

--- 2014-01-20 inst/doc/README.Rmd
@@
-reml
+EML
 ====
@@
 * Maintainer: Carl Boettiger
-* Authors: Carl Boettiger, Duncan Temple Lang, Karthik Ram, Matt Jones
+* Authors: Carl Boettiger, Claas-Thido Pfaff, Duncan Temple Lang, Karthik Ram, Matt Jones
 * License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
-* [Issues](https://github.com/ropensci/reml/issues): Bug reports, feature requests, and development discussion.
+* [Issues](https://github.com/ropensci/EML/issues): Bug reports, feature requests, and development discussion.
 
-_This is a dynamic document, generated by [knitr](http://yihui.name/knitr) using [vignette.Rmd](https://github.com/ropensci/reml/blob/master/inst/doc/vingette.Rmd) as the source._
+_This is a dynamic document, generated by [knitr](http://yihui.name/knitr) using [vignette.Rmd](https://github.com/ropensci/EML/blob/master/inst/doc/vingette.Rmd) as the source._
 
@@
 library("devtools")
-install_github("reml", "ropensci")
+install_github("EML", "ropensci")
 ```
@@
 ```{r, results="hide", message=FALSE, warning=FALSE}
-library("reml")
+library("EML")
 ```
@@
 
-`reml` now extends R's `data.frame` class by introducing the `data.set` class which includes additional metadata required by EML.
+`EML` now extends R's `data.frame` class by introducing the `data.set` class which includes additional metadata required by EML.
 A `data.set` can be created much like a `data.frame` by specifying additional arguments
@@
 ```{r}
-eml_write(dat, file = "reml_example.xml")
+eml_write(dat, file = "EML_example.xml")
 ```
 
-*for convenience, had we ommitted any essential metadata, such as providing only an unannotated `data.frame` in place of a `data.set`, `reml` will launch its metadata wizard to assist in constructing the metadata based on the data.frame provided*.
+*for convenience, had we ommitted any essential metadata, such as providing only an unannotated `data.frame` in place of a `data.set`, `EML` will launch its metadata wizard to assist in constructing the metadata based on the data.frame provided*.
 While this may be helpful starting out, regular users will find it faster to define the columns and units directly in the format above.
 
-See the [EML generated](https://github.com/ropensci/reml/tree/master/inst/doc/reml_example.xml) by this example.
+See the [EML generated](https://github.com/ropensci/EML/tree/master/inst/doc/EML_example.xml) by this example.
 
@@
 ```{r}
-eml_validate("reml_example.xml")
+eml_validate("EML_example.xml")
 ```
 
-* See [Open issues in writing EML](https://github.com/ropensci/reml/issues?labels=write+eml&page=1&state=open)
+* See [Open issues in writing EML](https://github.com/ropensci/EML/issues?labels=write+eml&page=1&state=open)
 
@@
 _Note: this example assumes `rfigshare` is installed, even though it is not 
-required to use reml and run the examples in the other sections.  See the 
+required to use EML and run the examples in the other sections.  See the 
 `rfigshare` documentation for authenticating `rfigshare` with your user account_
@@
 ```{r publish_figshare} 
-eml_publish("reml_example.xml", description="Example EML file from reml", categories = "Ecology", tags = "EML", destination="figshare")
+eml_publish("EML_example.xml", description="Example EML file from EML", categories = "Ecology", tags = "EML", destination="figshare")
 ```
@@
 
-See [Open issues in publishing EML](https://github.com/ropensci/reml/issues?labels=publish+eml&page=1&state=open)
+See [Open issues in publishing EML](https://github.com/ropensci/EML/issues?labels=publish+eml&page=1&state=open)
 
@@
 ```{r}
-obj <- eml_read("reml_example.xml")
+obj <- eml_read("EML_example.xml")
 ```
@@
 
-While the dependencies for basic functionality are kept to a minimum, to access all the functions and tests implemented in `reml` you'll need several additional packages.
+While the dependencies for basic functionality are kept to a minimum, to access all the functions and tests implemented in `EML` you'll need several additional packages.
 
@@
 
-***Suggestions, feature requests, bug reports and other feedback welcome!*** Please see our [issues tracker](https://github.com/ropensci/reml/issues)
+***Suggestions, feature requests, bug reports and other feedback welcome!*** Please see our [issues tracker](https://github.com/ropensci/EML/issues)

### 2014-01-21 inst/doc/knb.Rmd
rename package as EML instead of reml, closes #63

@@
 _Note: this example assumes `dataone` is installed, even though it is not 
-required to use reml and run the examples in the other sections.  See the 
+required to use EML and run the examples in the other sections.  See the 
 `dataone` documentation for authenticating `dataone` with a user account_
@@
 ```{r publish_knb}
-ids <- eml_publish("reml_example.xml", destination="KNB", permanent=FALSE)
+ids <- eml_publish("EML_example.xml", destination="KNB", permanent=FALSE)
 ```

### 2014-01-21 inst/doc/one-simple-way.Rmd
rename package as EML instead of reml, closes #63

@@
 
-Use `reml`.
+Use `EML`.
 
@@
 
-Sharing is as easy as `publish_reml("my_eml_data.xml")`
+Sharing is as easy as `publish_EML("my_eml_data.xml")`
 
@@
 
-`reml` can automatically push our data to a variety of established repositories, as illustrated in this example.
+`EML` can automatically push our data to a variety of established repositories, as illustrated in this example.
 
@@
 
-We see by default `reml` has provided us with the liberal CC0 license for the data, consistent with major repositories such as Dryad.  Other options can be specified if needed.
+We see by default `EML` has provided us with the liberal CC0 license for the data, consistent with major repositories such as Dryad.  Other options can be specified if needed.
 

### 2014-01-21 inst/tests/knb_test.R

@@ -2,7 +2,7 @@ context("Publish to KNB")
 
 
 
-library(reml)
+library(EML)
 dat = data.set(river = factor(c("SAC",  "SAC",   "AM")),
                spp   = c("Oncorhynchus tshawytscha",  "Oncorhynchus tshawytscha", "Oncorhynchus kisutch"),
                stg   = ordered(c("smolt", "parr", "smolt"), levels=c("parr", "smolt")), # levels indicates increasing level, eg. parr < smolt
@@ -29,7 +29,7 @@ eml_write(dat, contact = "Carl Boettiger <cboettig@ropensci.org>", file = "test.
 
 test_that("We can publish data to the KNB", {
   nodeid = "urn:node:mnDemo5" # A Development server for testing
-  pid <- reml:::eml_knb("test.xml", 
+  pid <- EML:::eml_knb("test.xml", 
                  mn_nodeid = nodeid,                   
                  cli = D1Client("DEV", nodeid)) # Use dev mode client
 

### 2014-01-21 inst/tests/test_literature_module.R

@@ -1,6 +1,6 @@
 context("Testing the literature module")
 
-require(reml)
+require(EML)
 
 # Create litrature filds for reuse
 a_creator = new("ListOfcreator", list(new("creator", individualName = new("individualName", givenName = "Claas-Thido", surName = "Pfaff")))) 

### 2014-01-21 inst/tests/test_measurementScale.R

@@ -2,10 +2,10 @@ context("testing measurement scale detection")
 
 
 
-require(reml)
+require(EML)
 data(ex2)
 
-s4 <- reml:::eml(dat, 
+s4 <- EML:::eml(dat, 
                  title = "title", 
                  creator = "Carl Boettiger <cboettig@gmail.com>")
 
@@ -23,7 +23,7 @@ s4 <- reml:::eml(dat,
 ##
 
 
-## See https://github.com/ropensci/workshops-norwich-2013-09/blob/0f8738b50d5613526fa9dd531698649c9ad18490/04-reproducible-research/reml.Rmd for an example with different fields (including Date)
+## See https://github.com/ropensci/workshops-norwich-2013-09/blob/0f8738b50d5613526fa9dd531698649c9ad18490/04-reproducible-research/EML.Rmd for an example with different fields (including Date)
 
 ## Write more simple data cases showing this diversity of data types..
 

### 2014-01-21 inst/tests/test_party.R

@@ -1,11 +1,11 @@
 context("party.R")
 
-require(reml)
+require(EML)
 
-test_that("Coercions work with reml:::contact_creator function", {
+test_that("Coercions work with EML:::contact_creator function", {
 
 
-  b <- reml:::contact_creator(contact = "Carl Boettiger <cboettig@gmail.com>")
+  b <- EML:::contact_creator(contact = "Carl Boettiger <cboettig@gmail.com>")
 
   expect_is(b$creator, "ListOfcreator")
   expect_is(b$contact, "contact")
@@ -15,14 +15,14 @@ test_that("Coercions work with reml:::contact_creator function", {
 
 
 
-  b <- reml:::contact_creator(creator = "Carl Boettiger <cboettig@gmail.com>")
+  b <- EML:::contact_creator(creator = "Carl Boettiger <cboettig@gmail.com>")
   expect_is(b$creator, "ListOfcreator")
   expect_is(b$contact, "contact")
   expect_identical(format(as(b$contact, "person")), "Carl Boettiger <cboettig@gmail.com>")
   expect_identical(b$contact@individualName@surName, "Boettiger")
   expect_identical(b$creator[[1]]@individualName@surName, "Boettiger")
 
-  b <- reml:::contact_creator(creator = "Carl Boettiger <cboettig@gmail.com>",
+  b <- EML:::contact_creator(creator = "Carl Boettiger <cboettig@gmail.com>",
                        contact = "Carl Boettiger <cboettig@gmail.com>")
   expect_is(b$creator, "ListOfcreator")
   expect_is(b$contact, "contact")
@@ -30,7 +30,7 @@ test_that("Coercions work with reml:::contact_creator function", {
   expect_identical(b$contact@individualName@surName, "Boettiger")
   expect_identical(b$creator[[1]]@individualName@surName, "Boettiger")
 
-  b <- reml:::contact_creator(creator = as("Carl Boettiger <cboettig@gmail.com>", "creator"),
+  b <- EML:::contact_creator(creator = as("Carl Boettiger <cboettig@gmail.com>", "creator"),
                        contact = "Carl Boettiger <cboettig@gmail.com>")
   expect_is(b$creator, "ListOfcreator")
   expect_is(b$contact, "contact")
@@ -38,7 +38,7 @@ test_that("Coercions work with reml:::contact_creator function", {
   expect_identical(b$contact@individualName@surName, "Boettiger")
[…]

### 2014-01-21 inst/tests/test_read.R

@@ -3,8 +3,8 @@ context("Parsing")
 ## test basic reading of XML into S4
 
 library(XML)
-require(reml)
-f <- system.file("examples", "hf205.xml",  package="reml")
+require(EML)
+f <- system.file("examples", "hf205.xml",  package="EML")
 
 test_that("we can parse XML into S4", {
 
@@ -21,9 +21,9 @@ test_that("we can read in S4 using eml_read", {
 })
 
 
-require(reml)
+require(EML)
 data(ex2)
-S4obj <- reml:::eml(dat, 
+S4obj <- EML:::eml(dat, 
                     title = "title", 
                     creator = "Carl Boettiger <cboettig@gmail.com>")
 

### 2014-01-21 inst/doc/README.Rmd
update installation of suggested packages notes

@@
 
-While the dependencies for basic functionality are kept to a minimum, to access all the functions and tests implemented in `EML` you'll need several additional packages.
+While the dependencies for basic functionality are kept to a minimum, to access all the functions and tests implemented in `EML` you'll need several additional packages from CRAN.
 
 ```r
-install.packages(c("yaml", "knitr", "rfigshare", "testthat", "RCurl"))
+install.packages(c("knitr", "rfigshare", "testthat", "RCurl", "dataone"))
 ```
@@
 ```r
-install.packages("uuid")
 install_github("RHTMLForms", "omegahat")

### 2014-01-21 R/eml_methods.R
complete migration to EML name

@@ -21,8 +21,8 @@
 #' @export 
 eml <- function(dataset = NULL,
                 title = "metadata",
-                creator = get("defaultCreator", envir=remlConfig),
-                contact = get("defaultContact", envir=remlConfig),
+                creator = get("defaultCreator", envir=EMLConfig),
+                contact = get("defaultContact", envir=EMLConfig),
                 coverage = eml_coverage(scientific_names = NULL,
                                         dates = NULL,
                                         geographic_description = NULL,
@@ -40,7 +40,7 @@ eml <- function(dataset = NULL,
   ## Coerce character string persons into EML representations
 
 
-  uid <- reml_id()
+  uid <- EML_id()
   who <- contact_creator(contact = contact, 
                          creator = creator)
 
@@ -223,8 +223,8 @@ get_col_classes <- function(attrs){
 
 
 
-contact_creator <- function(contact = get("defaultContact", envir=remlConfig), 
-                            creator = get("defaultCreator", envir=remlConfig)){
+contact_creator <- function(contact = get("defaultContact", envir=EMLConfig), 
+                            creator = get("defaultCreator", envir=EMLConfig)){
 
    ## Get a contact first 
   if(is.null(contact) || length(contact) == 0 || isEmpty(contact)){ # IF no contact given... 

### 2014-01-21 R/eml_write.R

@@ -5,12 +5,12 @@
 #' @aliases eml_write write.eml 
 #' @import XML
 #' @export eml_write write.eml
-#' @include reml_environment.R
+#' @include EML_environment.R
 eml_write <- function(dat,
                       file = NULL,
                       title = "metadata",
-                      creator = get("defaultCreator", envir=remlConfig), 
-                      contact = get("defaultContact", envir=remlConfig),
+                      creator = get("defaultCreator", envir=EMLConfig), 
+                      contact = get("defaultContact", envir=EMLConfig),
                       ...){
 
   ## dat Types we can handle by coercion

### 2014-01-21 R/figshare.R

@@ -11,7 +11,7 @@
 #' @return the figshare id
 #' @details figshare requires authentication. See rfigshare 
 #' [tutorial]() for help on configuring this.  
-#' Arguments for figshare are optional, if not provided reml will attempt to
+#' Arguments for figshare are optional, if not provided EML will attempt to
 #' extract appropriate values from the EML.  If no values are found for 
 #' essential metadata (title, description, category), then figshare object
 #' will be still be created as a draft but cannot be published. 
@@ -24,8 +24,8 @@
 #' @import XML
 #' @examples 
 #' \dontrun{
-#'    file <- system.file("doc", "my_eml_data.xml", package="reml")
-#'    id <- eml_figshare(file, description="Example EML file from reml", categories = "Ecology", tags = "EML")
+#'    file <- system.file("doc", "my_eml_data.xml", package="EML")
+#'    id <- eml_figshare(file, description="Example EML file from EML", categories = "Ecology", tags = "EML")
 #' }
 eml_figshare <- function(file, title = NULL, description = NULL, 
                          categories = NULL, tags = NULL, links = NULL, 

### 2014-01-21 data/ex4.R

@@ -1,4 +1,4 @@
-library(reml)
+library(EML)
 dat = data.set(river = factor(c("SAC",  "SAC",   "AM")),
                spp   = c("Oncorhynchus tshawytscha",  "Oncorhynchus tshawytscha", "Oncorhynchus kisutch"),
                stg   = ordered(c("smolt", "parr", "smolt"), levels=c("parr", "smolt")), # levels indicates increasing level, eg. parr < smolt

### 2014-01-21 inst/doc/README.Rmd
update installation instructions for suggested packages

@@
 ```r
-install_github("RHTMLForms", "omegahat")
-install_github("XMLSchema", "omegahat")
+# install_github("Sxslt", "omegahat")
+# install_github("XMLSchema", "omegahat") 
+# install_github("RHTMLForms", "omegahat")
+
+install_github("RWordXML", "duncantl") # fails, clone and install locally
+
+install.packages("Sxslt", repos="http://www.omegahat.org/R", type="source")
+install.packages("RHTMLForms", repos="http://www.omegahat.org/R", type="source")
+install.packages("XMLSchema", repos="http://www.omegahat.org/R", type="source")
+
 ```

### 2014-01-28 inst/doc/pubs/reml.Rmd
In sync with new renamed repo now

--- 2013-07-10 inst/doc/reml.Rmd
@@
 
-1. Reuse is essential
-2. Reuse requires good metadata standards
-3. reml provides an easy way to utilize the Ecological Metadata Language to generate, access, and publish data.
+1. Following metadata standards facilitates reuse
+2. data 
+3. EML provides an easy way to utilize the Ecological Metadata Language to generate, access, and publish data.
 
@@
 
-The `reml` package is built for the Ecological Markup Language schema, [version 2.1.1], using the R software environment and the XML package for parsing and generating XML [Temple Lang 2013, XML].
+The `EML` package is built for the Ecological Markup Language schema, [version 2.1.1], using the R software environment and the XML package for parsing and generating XML [Temple Lang 2013, XML].
 
@@
 
-The `reml` package can perform three levels of validation on EML files written from or read into R: validating the XML, validating against the EML 2.1.1 Schema, and performing
+The `EML` package can perform three levels of validation on EML files written from or read into R: validating the XML, validating against the EML 2.1.1 Schema, and performing
 
-Several mechanisms are in place to evaluate the quality of the `reml` package functions themselves.
-In addition to the examples from the documentation and the automated package checks provided for the R system, the `reml` package includes a complete test suite using the testthat package [Wickham 2011].
+Several mechanisms are in place to evaluate the quality of the `EML` package functions themselves.
+In addition to the examples from the documentation and the automated package checks provided for the R system, the `EML` package includes a complete test suite using the testthat package [Wickham 2011].
 
@@
 
-The `reml` package is available through the CRAN repository under a CC0 license.
-At the time of writing, the current version is `r packageDescription("reml", fields="Version")`.
-The current development branch can be found on Github at [https://github.com/ropensci/reml].
+The `EML` package is available through the CRAN repository under a CC0 license.
+At the time of writing, the current version is `r packageDescription("EML", fields="Version")`.
+The current development branch can be found on Github at [https://github.com/ropensci/EML].
 See the NEWS file for changes in recent versions.
@@
 
-There are many promising directions for further development of the `reml` package.
+There are many promising directions for further development of the `EML` package.
 The project's Milestones page provides a list of steps currently planned or under consideration.
@@
 
-This project was supported in part by a grant from the Alfred P Sloan Foundation (CB and KR), NSF Grant DBI-1306697 (CB), ... `reml` is part of the rOpenSci project, [http://ropensci.org].
+This project was supported in part by a grant from the Alfred P Sloan Foundation (CB and KR), NSF Grant DBI-1306697 (CB), ... `EML` is part of the rOpenSci project, [http://ropensci.org].
 

### 2014-01-28 R/eml_config.R

@@ -11,14 +11,15 @@
 #' @param creator a character string, person object, or creator object
 #' @param contact  a character string, person object, or contact object
 #' @param defaultRights a text string specifying the default intellectual rights text that should be indicated.  
-#' @return Nothing, but updates the remlConfig environment where this information is stored.  
+#' @return Nothing, but updates the EMLConfig environment where this information is stored.  
 #' @details Configuration settings will be used as the default values when creating a new eml file or dataset object
 #'   unless alternative values are provided at the time of creation.  Configuration settings last only for the duration
-#'   of the current R session.  Add a call to `reml::eml_config` in your .Rprofile to have these settings used by
+#'   of the current R session.  Add a call to `EML::eml_config` in your .Rprofile to have these settings used by
 #'   default in each R session.  Repeating a call to eml_config will reset the settings. The default settings of `eml_config`
 #'   are not applied if the function is not called explicitly.  
 #' @include reml_environment.R
 #' @include party.R 
+#' @import methods
 #' @export
 eml_config <- function(creator = "", contact = "", defaultRights = "CC0"){
   if(defaultRights == "CC0")
@@ -29,15 +30,15 @@ eml_config <- function(creator = "", contact = "", defaultRights = "CC0"){
     defaultContact <- as(creator[[1]], "contact")
  else 
     defaultContact <- as(contact, "contact")
-  assign('defaultRights', defaultRights, envir=remlConfig)
-  assign('defaultCreator', defaultCreator, envir=remlConfig)
-  assign('defaultContact', defaultContact, envir=remlConfig)
+  assign('defaultRights', defaultRights, envir=EMLConfig)
+  assign('defaultCreator', defaultCreator, envir=EMLConfig)
+  assign('defaultContact', defaultContact, envir=EMLConfig)
 }
 eml_config()
 
 
 eml_get_config <- function(x){
-  get(x, envir = remlConfig)
+  get(x, envir = EMLConfig)
 }
 
 

### 2014-01-28 R/eml_write.R

@@ -2,6 +2,14 @@
 
 #' Top-level API function for writing eml 
 #' @param dat either a data.frame, data.set, or eml object
+#' @param file the filename for the output EML.  If NULL, will return the 
+#' XML text as a character string.  
+#' @param title a title for the EML metadata file, also used in naming the csv file.  
+#' @param creator an EML creator (or ListOfcreator for multiple creators) object, 
+#' R person object, or text string recognizable as a name or vector of names.  
+#' @param contact an EML contact, R person object with email address, or text
+#' string with email in angle brackets.  
+#' @param ... additional arguments to the dataset constructor, see ?dataset
 #' @aliases eml_write write.eml 
 #' @import XML
 #' @export eml_write write.eml
@@ -9,8 +17,9 @@
 eml_write <- function(dat,
                       file = NULL,
                       title = "metadata",
-                      creator = get("defaultCreator", envir=remlConfig), 
-                      contact = get("defaultContact", envir=remlConfig)){
+                      creator = get("defaultCreator", envir=EMLConfig), 
+                      contact = get("defaultContact", envir=EMLConfig),
+                      ...){
 
   ## dat Types we can handle by coercion
   if(is(dat, "eml")) {
@@ -23,7 +32,11 @@ eml_write <- function(dat,
   } else { ## assemble minimal valid information  
 
 
-    s4 <- eml(dat = dat, title = title, creator = creator, contact = contact)
+    s4 <- eml(dat = dat, 
+              title = title, 
+              creator = creator, 
+              contact = contact,
+              ...)
   }
   xml <- as(s4, "XMLInternalElementNode")
   saveXML(xml, file = file)

### 2014-01-28 R/knb.R

@@ -1,21 +1,25 @@
 #' Upload EML to knb
-#' 
+#'
+#' Upload EML to knb
+#' @param file the filename of the EML file to upload
+#' @param permanent logical, indicate whether we should use the production server.
+#' default is FALSE, implies use of the Demo server, overriding any arguments set
+#' to mn_nodeid and cli.  
+#' @param public logical, make object public?  Otherwise only authorized users may access file.
+#' @param mn_nodeid The server ID to which we upload (if permanent is FALSE).  The default
+#' path is to the KNB, which can index all EML metadata through metacat. 
+#' @param cli the dataone client object. See dataone::D1Client 
 #' @details knb requires authentication. 
+#' @return the EML identifier, csv file identifier, and dataone package identifier 
+#'  (a metadata object associating the two files in the dataone repository).  
 #' @import XML
-#' @examples 
-#' \dontrun{
-
-# Initialize a client to interact with DataONE
-
-
-#' }
 eml_knb <- function(file,
                     permanent = FALSE,
                     public = TRUE,
                     mn_nodeid = "urn:node:KNB",
                     cli = D1Client("PROD", mn_nodeid)){
 
-
+  # require call leads to a warning when package is not on the SUGGESTS list.  A problem for TRAVIS, not for CRAN.  
   success <- require(dataone)
   if(!success){
     message("dataone package not found. Attempting to install")

### 2014-01-28 R/xsl_coercions.R

@@ -1,11 +1,23 @@
 
 
 
-# file <- system.file("examples", "hf205.xml", package = "reml")
-#' @export
-get_rdf <- function(file){
-  require(Sxslt)
+#' get rdf
+#'   
+#' get rdf
+#' @param eml an eml file, file name, or XML content (text string). See xsltApplyStyleSheet
+#' @return an rdf file containing metadata that could be rendered in Dublin Core.  
+# @export
+# @examples
+# file <- system.file("examples", "hf205.xml", package = "EML")
+# get_rdf(file) # not working yet
+
+get_rdf <- function(eml){
+#   require(Sxslt) # Wait until package is available on CRAN to formally depend on it.  
+  # FIXME stylesheet not working yet
   # FIXME detect EML version first in selecting which stylesheet to use.
-  to_rdf <- system.file("xsl", "eml210toDublinCore.xsl", package="reml")
-  rdf <- xsltApplyStyleSheet(file, to_rdf)
+  if(is(eml, "eml"))
+    eml <- eml_write(eml)
+
+  to_rdf <- system.file("xsl", "eml210toDublinCore.xsl", package="EML")
+  rdf <- xsltApplyStyleSheet(eml, to_rdf)
 }

### 2014-01-28 inst/examples/hf205.R

@@ -4,12 +4,12 @@
 
 ## Illustrate reading in based on identifier and based on URL
 
-f <- system.file("examples", "hf205.xml", package="reml")
+f <- system.file("examples", "hf205.xml", package="EML")
 hf205 <- read.eml(f)
 
 dat <- eml_get(hf205, "data.set")
 
-# dat <- read.csv(system.file("examples", "hf205-01-TPexp1.csv", package="reml"))
+# dat <- read.csv(system.file("examples", "hf205-01-TPexp1.csv", package="EML"))
 
 
 
@@ -149,7 +149,7 @@ rights <- "This dataset is released to the public and may be freely
 ## Read in methods description from an Word file.  
 library(RWordXML)
 library(XML)
-f2 <- wordDoc(system.file("examples", "methods.docx", package="reml"))
+f2 <- wordDoc(system.file("examples", "methods.docx", package="EML"))
 doc <- f2[[getDocument(f2)]]
 txt <- xpathSApply(doc, "//w:t", xmlValue)
 ## FIXME add <title> <section> and <para> blocking back: 
@@ -165,7 +165,7 @@ additionalMetadata <- hf205@additionalMetadata # extracted from previous eml fil
 ## Construct the dataTable #########
 
 
-dataTable <- reml:::eml_dataTable(dat, 
+dataTable <- EML:::eml_dataTable(dat, 
                            title = "Supplemental Data, table 1", 
                            description = "Metadata documentation for S1.csv", 
                            file = "S1.csv")
@@ -186,6 +186,6 @@ eml <- new("eml",
             dataset = dataset,
             additionalMetadata = additionalMetadata)
 
-write.eml(eml, "hf205_from_reml.xml")
+write.eml(eml, "hf205_from_EML.xml")
 
 

### 2014-01-28 inst/tests/test_hf205.R

@@ -1,13 +1,0 @@
-context("Identical read in and write out of an existing EML file")
-
-## test basic reading of XML into S4
-
-
-test_that("we can read in and write out the file", {
-  require(reml)
-  f <- system.file("examples", "hf205.xml",  package="reml")
-  eml <- read.eml(f)
-  ex <- eml_write(eml) #  version is preserved 
-  eml_validate(ex)
-})
-

### 2014-01-28 inst/tests/test_measurementScale.R

@@ -1,30 +1,0 @@
-context("testing measurement scale detection")
-
-
-
-require(reml)
-data(ex2)
-
-s4 <- reml:::eml(dat, 
-                 title = "title", 
-                 creator = "Carl Boettiger <cboettig@gmail.com>")
-
-
-## check that a factor becomes nominal/enumeratedDomain with levels defined
-
-## check that an ordered becomes an ordinal/enumeratedDomain
-
-## check that a numeric becomes a ratio
-
-## check that a Date becomes a dateTime
-
-## check that a character can become a nominal/textDefinition
-
-##
-
-
-## See https://github.com/ropensci/workshops-norwich-2013-09/blob/0f8738b50d5613526fa9dd531698649c9ad18490/04-reproducible-research/reml.Rmd for an example with different fields (including Date)
-
-## Write more simple data cases showing this diversity of data types..
-
-

### 2014-01-28 inst/tests/test_publish.R

@@ -1,33 +1,0 @@
-context("publish")
-
-test_that("We can publish a draft to figshare", {
-
-  library(reml)
-  library(XML)
-
-  f <- system.file("examples", "hf205.xml",  package="reml")
-  csv <- system.file("examples", "hf205-01-TPexp1.csv",  package="reml")
-
-  # move files to working directory
-  saveXML(xmlParse(f), "hf205.xml")
-  write.csv(read.csv(csv), "hf205-01-TPexp1.csv")
-
-  ## Publish the data to figshare (as a draft)
-  id <- eml_publish("hf205.xml", categories="Ecology", destination="figshare")
-
-  ## Confirm that the appropriate metadata has been written to figshare
-  library(rfigshare)
-  details <- fs_details(id, mine=TRUE)
-  expect_that(details$categories[[1]]$name, equals("Ecology"))
-
-  ## Confirm that the EML contains the figshare metadata 
-  doc <- xmlParse("figshare_hf205.xml") # check uploaded copy, since cannot download draft/private file
-  eml_cat <- xpathSApply(doc, "//additionalMetadata[@id = 'figshare']/metadata/keywordSet/keyword", xmlValue)
-  expect_that(eml_cat, equals("Ecology"))
-
-  ## Clean up 
-  fs_delete(id)
-  unlink("hf205-01-TPexp1.csv")
-  unlink("hf205.xml")
-  unlink("figshare_hf205.xml")
-})

### 2014-01-28 tests/test-all.R

@@ -1,0 +1,7 @@
+if (packageVersion("testthat") >= "0.7.1.99") {
+    library(testthat)
+  test_check("EML")
+} else {
+  library(testthat)
+  test_package("EML")
+}

### 2014-01-28 tests/testthat/test_extract.R

@@ -1,0 +1,72 @@
+context("extract")
+
+
+## FIXME add unlink commands to remove/cleanup any files created by tests (e.g. the .csv and .xml files)
+
+library(EML)
+dat = data.set(river = factor(c("SAC",  "SAC",   "AM")),
+               spp   = c("Oncorhynchus tshawytscha",  "Oncorhynchus tshawytscha", "Oncorhynchus kisutch"),
+               stg   = ordered(c("smolt", "parr", "smolt"), levels=c("parr", "smolt")), # levels indicates increasing level, eg. parr < smolt
+               ct    = c(293L,    410L,    210L),
+               day   = as.Date(c("2013-09-01", "2013-09-1", "2013-09-02")),
+               stringsAsFactors = FALSE, 
+               col.defs = c("River site used for collection",
+                            "Species scientific name",
+                            "Life Stage", 
+                            "count of live fish in traps",
+                            "day traps were sampled (usually in morning thereof)"),
+               unit.defs = list(c(SAC = "The Sacramento River",                         # Factor 
+                                  AM = "The American River"),
+                                "Scientific name",                                      # Character string (levels not defined)
+                                c(parr = "third life stage",                            # Ordered factor 
+                                  smolt = "fourth life stage"),
+                                c(unit = "number", precision = 1, bounds = c(0, Inf)),  # Integer
+                                c(format = "YYYY-MM-DD", precision = 1)))               # Date
+
+
+S4obj <- EML:::eml(dat, title = "title", 
+             creator = "Carl Boettiger <cboettig@gmail.com>")
+
+
+
+test_that("we can extract the data unaltered (using method for object 'physical')", {
+  out_dat <- EML:::extract(S4obj@dataset@dataTable[[1]]@physical, using=EML:::col_classes(S4obj)) ## Using is ignored...
+
+ 
+  expect_equal(dat[["river"]], out_dat[["river"]])
+
+
+#  expect_equal(dat, out_dat) ## FIXME not quite, classes differ 
+ # expect_identical(dat, out_dat) ## FIXME should be identical, but isn't because integers are cast as eml:ratio which is cast as numeric.  Need to add mechanism for integers EML::detect_class
+})
+
+
+test_that("we can extract from alternative paths", {
[…]

### 2014-01-28 tests/testthat/test_validate.R

@@ -1,0 +1,35 @@
+context("external validator")
+
+
+test_that("External validator works using RHTMLForms", {
+  library(RHTMLForms)
+  ff = getHTMLFormDescription("http://knb.ecoinformatics.org/emlparser/")
+#eml_validate = createFunction(ff[[2]], reader = processValidationResponse) # failing...
+
+# printing eml_validate to include as source code below currently
+# loses the class information on the .formDescription default value
+# and that causes checkFormArgs to complain.
+
+  library(XML)
+  library(EML)
+## A case that should validate
+  txt = saveXML(xmlParse("http://knb.ecoinformatics.org/emlparser/eml-sample.xml"))
+  o = eml_validate(txt)
+#hdoc = htmlParse(o)
+#xpathSApply(hdoc, "//h2")
+
+
+## A case that should fail validation
+  sampleDoc = xmlParse("http://knb.ecoinformatics.org/emlparser/eml-sample.xml")
+  r = xmlRoot(sampleDoc)
+  newXMLNode("garbage", parent = r)
+  newXMLNode("otherNonsense", parent = r)
+
+  o = eml_validate(saveXML(sampleDoc))
+
+## 
+
+#hdoc = htmlParse(o)
+#xpathSApply(hdoc, "//h2")
+
+})

### 2014-01-28 tests/testthat/test_write.R

@@ -1,0 +1,82 @@
+context("Serializing S4 to XML")
+
+## Test basic writing of S4 classes back into XML
+require(EML)
+
+## Okay, we know this works, so stick it in the global space for the next functions...
+
+## Rest of tests use data.set object
+
+dat = data.set(river = factor(c("SAC",  "SAC",   "AM")),
+               spp   = c("Oncorhynchus tshawytscha",  "Oncorhynchus tshawytscha", "Oncorhynchus kisutch"),
+               stg   = ordered(c("smolt", "parr", "smolt"), levels=c("parr", "smolt")), # levels indicates increasing level, eg. parr < smolt
+               ct    = c(293L,    410L,    210L),
+               day   = as.Date(c("2013-09-01", "2013-09-1", "2013-09-02")),
+               stringsAsFactors = FALSE, 
+               col.defs = c("River site used for collection",
+                            "Species scientific name",
+                            "Life Stage", 
+                            "count of live fish in traps",
+                            "day traps were sampled (usually in morning thereof)"),
+               unit.defs = list(c(SAC = "The Sacramento River",                         # Factor 
+                                  AM = "The American River"),
+                                "Scientific name",                                      # Character string (levels not defined)
+                                c(parr = "third life stage",                            # Ordered factor 
+                                  smolt = "fourth life stage"),
+                                c(unit = "number", precision = 1, bounds = c(0, Inf)),  # Integer
+                                c(format = "YYYY-MM-DD", precision = 1)))               # Date
+
+
+
+
+
+
+test_that("We can write S4 EML to XML and validate", {
+
+  eml_write(dat, contact = as("Carl Boettiger <cboettig@gmail.com>", "contact"), file="title.xml")
+
+  o <- eml_validate("title.xml")                
+  expect_true(all(o)) # all cases validate
+
+   unlink("title.xml")
+   unlink("title.csv")
+})
+
+
+
+# FIXME ADD ID 
+test_that("We can add coverage information and validate", {
+  contact <- as("Carl Boettiger <cboettig@ropensci.org>", "contact")
+  creator <- c(as(contact, "creator"))
+  uid <- EML:::EML_id()
+  eml_write(new("eml",
+                packageId = uid[["id"]], 
+                system = uid[["system"]],
+                scope = uid[["scope"]], 
+                dataset = 
+                new("dataset", 
+                  dataTable = c(eml_dataTable(dat, 
+                                              description = "description",
[…]

### 2014-01-28 inst/doc/pubs/one-simple-way.Rmd
writing on one simple way

--- 2014-01-21 inst/doc/one-simple-way.Rmd
@@
 # One simple way to share your data
+
+_Ideas in Ecology and Evolution_ recently published a paper by White _et
+al._ that should be considered required reading in any graduate training
+program Ecology or Evolution training program. Entitled _"Nine simple ways
+to make it easier to (re)use your data"_, the authors present a thorough
+and accessible overview of practices essential for any researcher but
+getting increasing intention in light of spreading requirements from
+primary publishers and funders of ecological and evolutionary research
+for the formal deposition and archiving of data upon publication. The
+spread of such requirements brings new urgency to the elements described
+by White and colleagues, which echo sentiments and recommendations that
+have appeared regularly in this literature over the previous years and
+decades.  The development of standards such as the Ecological Metadata
+Language, the emergence of suitable online data repositories and the
+widespread adoption of scripting languages such as R among students of
+Ecology and Evolution have all reduced barriers to sharing and reusing
+data efficiently and effectively.
+
+Here we introduce the `EML` package for the R statistical environment, 
+as one simple way to accomplish these nine elements and much more besides.
+
+<!--
+These developments, together with new semantic technologies currently on
+the horizon, and the billons of dollars invested in enormous monitoring
+and data collection efforts such as the National Ecological Observatory
+Network, the Ocean Observatory Intiative, and similar efforts offer
+unpresidented opportunities to tackle the rich, integrative and data-intensive 
+questions that challenge our globe today.  
+-->
 
@@
 
-Use `EML`.
+You have your data as a `data.frame` in R, convient for your analyses
+and graphs.  How best to archive share that data in a way that will
+maximize its utility to your lab group, your future self?  What will
+most impress your NSF data management review, or improve the chances
+that others will use and cite your data set?  Good metadata.
 
-You have your data as a `data.frame` in R, convient for your analyses and graphs.
-How best to archive share that data in a way that will maximize its utility to your lab group, your future self?
-What will most impress your NSF data management review, or improve the chances that others will use and cite your data set?
-Good metadata.
-
-White et al.
-outline "Nine simple ways to make it easier to (re)use your data".
-Here we present one simple way to meet these criteria.
 Before we worry about what the criteria are, let's just jump in and make our archival data shared:
@@
                  stg = c("smolt", "parr", "smolt"),
-                 ct =  c(293L, 410L, 210L))
+                 ct =  c(293, 410, 210L))
 
@@
 
-Define the units used in each column, or the abbreviations we use in factors.
+Define the units used in each column.  In the case of factors, we define
+the abbreviations in a named string.
 
@@
 
-Specify the basic file metadata of author/contact, title, and a description, and choose where to publish:
+```{r}
+dat_with_metadata <- data.set(dat, col.defs = col_metadata, unit.defs = unit_metadata)
+```
+
+Specify the basic file metadata of creator and contact person, give it
+a title, and even choose where to publish:
 
 ```{r}
-eml_write(dat, col_metadata, unit_metadata, 
-          .contact="Carl Boettiger <cboettig@gmail.com>",  
-          .title="A test EML file", 
-          file_description="Test data, only intended for testing",
-          destination="figshare")
+eml_write(dat_with_metadata,
+          contact = "Carl Boettiger <cboettig@ropensci.org>",  
+          title = "A test EML file", 
+          destination = "figshare")
 ```
 
+Just like that, we have accomplished all nine steps, and more besides.
+In the sections below, we'll unpack what this function has done, and how
+to take advantage of this metadata when (re)using this data in the future.
+
+<!--
+
+If we had used just the original `data.frame` as the first argument, the
+function would prompt us for the appropriate unit and column metadata. If
+no destination is given, a filename can be provided instead.
+
 We can provide much richer annotation than this example includes.
-We can also store information such as the contact name and address in a configuration option to avoid having to provide it repeatedly.
+We can also store information such as the contact name and address
+in a configuration option to avoid having to provide it repeatedly.
 Let's see how this example does on the Nine simple criteria.
+
+-->
 
@@
 
-Sharing is as easy as `publish_EML("my_eml_data.xml")`
+Sharing is as easy as `publish_eml("my_eml_data.xml")`
 
@@
 
-R has good null values in its use of `data.frames`, so we already have these.
-Better yet, we write out our choice of null value in the metadata.
-Reading in the data, we can explicitly pass this to `read.csv` rather than just assume the null value is appropriate.
-This is handled automatically by `eml_read`.
+R has good null values in its use of `data.frames`, so we already have
+these.  Better yet, we write out our choice of null value in the metadata.
+Reading in the data, we can explicitly pass this to `read.csv` rather than
+just assume the null value is appropriate.  This is handled automatically
+by `eml_read`.
 
@@
 
-We have quality control on the metadata as well as the data.  While user mistakes are still possible, automated validation of the metadata provided and
+We have quality control on the metadata as well as the data.  While user
+mistakes are still possible, automated validation of the metadata
+provided and
 
@@
 
-`EML` can automatically push our data to a variety of established repositories, as illustrated in this example.
+The `EML` package can automatically push our data to a variety of
+established scientific repositories, including some of those discussed by
+White _et al._.  In the introductory example we submitted the data file to
+figshare, an internationally recognized interdisciplinary data repository
+that is integrated with PLOS and several Nature Publishing Group journals.
+This command uploaded both the raw data in the recommended plain-text
+CSV format and the XML file containing the EML metadata description to
+the same figshare fileset object.  Like most repositories, figshare
+requires basic metadata of it's own to enable it users to search and
+discover content.  The `EML` package has automatically extracted the
+dataset creators and title, as well as any description or keywords already
+available in the metadata and copied them over to appropriate fields
+in the figshare metadata. In return, figshare has supplied us with a
+permanent identifier (DOI) that can be used to persistently locate and
+cite this data, even if figshare itself goes out of business.
+
+Other repositories may be even better suited for publishing of
+EML-annotated data. The Knowledge Network for Biocomplexity (KNB) is
+particularly appropriate for EML annotated data, as a member of the
+DataONE network (see @Reichman2011) and because the repository already
+"speaks" EML (@Jones2005).  This means that while figshare exposes only
+a few of the metadata fields for search and discovery, a researcher can
+search for particular attribute level information on the KNB.  Such as
+"all data sets that have a column containing dissolved oxygen levels".
+
+Publishing to the KNB is just as easy, after authentication credentials
+have been set up.
+
+```r
+eml_publish(eml, permanent = TRUE, identifier = "doi", destination="knb")
+```
+
+The KNB provides several different kinds of identifiers, including
+`uuid` and `doi`. A DOI (`doi`) should only be requested for finished,
+published, public dataset, not individual files within a dataset or
+in a dataset that is being regularly updated.  By default, the `EML`
+package uses `uuid` for unique ids of component files and EML packages.
+The option `permanent = TRUE` can be omitted for testing purposes.
+
+The KNB already hosts several decades worth of EML metadata files from
+around the world, including all data collected by Long Term Ecological
+Research (LTER) sites.  Researchers can make these powerful search
+queries directly from the R terminal, and use the `EML` package to parse
+and explore the resulting hits.
 
@@
 
-We see by default `EML` has provided us with the liberal CC0 license for the data, consistent with major repositories such as Dryad.  Other options can be specified if needed.
+By default, the `EML` package has provided us with the liberal CC0
+declaration for the data, consistent with major repositories such as
+Dryad and Figshare.
 
-Ethan White et al have recently published an important paper on "Nine simple ways to share your data", which has garnered considerable attention even as a preprint [^1]
+```{r}
+eml_get(eml, "rights")
+```
 
-[^1]: Disclaimer: I reviewed this paper, Karthik Ram served as handling editor.
+White _et al._ discuss the motivation for CC0 as the recommended default
+for scientific data, while also pointing to other license options.
+We could have provided many other possible rights when writing the EML
+by using the `intellectualRights` argument in an `eml` or `eml_write`
+function call:
+
+```{r}
+eml_write(dat_with_metadata,
+          contact = "Carl Boettiger <cboettig@ropensci.org>",  
+          title = "A test EML file",
+          intellectualRights = "Name of an extablished license or
+          copyright declaration, Text string describing the rights to
+          reuse this data, request for appropriate acknowledgment, etc")
+```
+
+[^1]: Disclaimer: CB served as a reviewer and KR as handling editor for White et al.

### 2014-01-29 inst/doc/pubs/one-simple-way.Rmd
still working on draft blog post

@@
 
-The data is publically available at the DOI returned.
-At this DOI we find two files: the raw data in `.csv` format, and the EML data file providing machine readable metadata.
-We'll unpack all the information provided in this metadata as we go along.
-Many other options for sharing are possible, including private sharing and sharing through a variety of data publication platforms.
+The data is publically available at the DOI returned. At this DOI we
+find two files: the raw data in `.csv` format, and the EML data file
+providing machine readable metadata.  We'll unpack all the information
+provided in this metadata as we go along.  Many other options for sharing
+are possible, including private sharing and sharing through a variety
+of data publication platforms.
 
@@
 
-We provide `n` lines of metadata in the EML file, one of the "metadata standards" mentioned by White et al.  This includes basic information like the title, author(s), and contact address.
+We provide `n` lines of metadata in the EML file, one of the "metadata
+standards" mentioned by White et al.  This includes basic information
+like the title, author(s), and contact address.
 
@@
 Methods sections in the file provide answers to these questions.
-Though our minimal example does not describe the collection itself, note that we nonetheless have generated a `methods` node, which describes the software that created the metadata file (after all, this is a step in the analysis).
+Though our minimal example does not describe the collection itself, note
+that we nonetheless have generated a `methods` node, which describes
+the software that created the metadata file (after all, this is a step
+in the analysis).
 
@@
 
-Rich access options are possible depending on the publication outlet, including specifying what users or groups have access to the data.
+Rich access options are possible depending on the publication outlet,
+including specifying what users or groups have access to the data.
 
@@
 
-We can see at a glance what taxa are covered by the data.  More detailed description can be provided in an abstract.
+We can see at a glance what taxa are covered by the data.  More detailed
+description can be provided in an abstract.
 
-> 4) warnings about known problems or inconsistencies in the data, and 
-> 5) information to check that the data are properly imported, such as the number of rows and columns in the dataset and the total sum of numerical columns
+> 4) warnings about known problems or inconsistencies in the data,
+ and 
+> 5) information to check that the data are properly imported, such
+as the number of rows and columns in the dataset and the total sum of
+numerical columns
 
-Checks have automatically been performed in generating the data, and are performed in reading in the data into R.
+Checks have automatically been performed in generating the data, and
+are performed in reading in the data into R.
 
@@
 
-The second file provides the unprocessed data as a `.csv` file.  Because `csv` file formats vary
+The second file provides the unprocessed data as a `.csv` file.
+Because `csv` file formats can vary in the details of data orientation,
+delimeters, and so forth, this information should also be provided as metadata
+to ensure reliable and machine-readable intepretation.  The metadata generated 
+using the EML package automatically adds this information based on R's handling
+of csv files:
+
+```xml
+<objectName>data.csv</objectName>
+<size>4984</size>
+<characterEncoding>en_US.UTF-8</characterEncoding>
+<dataFormat>
+<textFormat>
+  <numHeaderLines>1</numHeaderLines>
+  <numFooterLines>0</numFooterLines>
+  <recordDelimiter>\r\n</recordDelimiter>
+  <attributeOrientation>column</attributeOrientation>
+  <simpleDelimited>
+    <fieldDelimiter>,</fieldDelimiter>
+  </simpleDelimited>
+</textFormat>
+</dataFormat>
+```
 
@@
 
-Of the "nine ways", this one places the greatest burden on the creator and user.  Rather than coercing our data into prescribed formats, EML lets us simply describe whatever format we are using.
+Of the "nine ways", this one places the greatest burden on the creator
+and user.  Rather than coercing our data into prescribed formats, EML
+lets us simply describe whatever format we are using.
 
@@
 
-Our example generate standard .csv file format, and declares this explicitly so that users can be sure what character is used to seperate values, rows, etc:
+Our example generate standard .csv file format, and declares this
+explicitly so that users can be sure what character is used to seperate
+values, rows, etc:
 
@@
 - Standard table formats
+
 - Standard cell formats
 
-Rather than insist everyone use the ISO8601 format for dates, the user can declare whatever date format they prefer.
+White _et al._ recommend that all users learn and adopt certain standard
+formats for representing data, such as the ISO8601 standard for dates
+(YYYY-MM-DD).  While we support this recommendation, this practice can be
+not only burdensome to the data creator but also leaves any future user of
+the data guessing or hoping that the format is one that they recognize.
+Like most computer programs, R does not assume that any such pattern
+is a date in ISO8601 format, but rather uses a Date type that permits
+a variety of printing options.  Similarly, the `EML` package relies on
+the metadata to state which cells contain dates and in what format those
+dates are rendered.
+
+Likewise, the package detects the appropriate metadata based on the R 
+representation (such as factor, ordered factor, numeric, integer, character, or Date),
+and writes the appropriate metadata for this format.
 
@@
 
-R has good null values in its use of `data.frames`, so we already have
-these.  Better yet, we write out our choice of null value in the metadata.
+R has good null values in its use of `data.frames`, and permits 
+the user to identify other values for null or missing data.
+
+Better yet, we write out our choice of null value in the metadata.
+
 Reading in the data, we can explicitly pass this to `read.csv` rather than
@@
 
-Context nodes serve the pupose described in this section.   
+The detailed column and unit metadata serve the pupose described in this section of White _et al_.   
 Machine-readable metadata,
@@
 We have quality control on the metadata as well as the data.  While user
-mistakes are still possible, automated validation of the metadata
-provided and
+mistakes are still possible, automated validation of the metadata are
+provided reduce the chance of errors.
+
+The `EML` package can perform three levels of validation on EML files
+written from or read into R: validating the XML, validating against the
+EML Schema, and performing other quality control checks.  The `eml_validate`
+function lets a user confirm that the metadata meets the EML standard,
+helping to catch errors that could cause difficulty for another or user
+or computer in intepreting the metadata:
+
+```{r}
+eml_validate("example.xml")
+```
+
+The `eml_validate` function is also run automatically any time a
+user attempts to publish data to an external respository. MD5 checksums
+of the data are also computed at this time to ensure that only complete 
+and uncorrupted data is tranfered over the connection.
+
+Several mechanisms are in place to evaluate the quality of the `EML`
+package functions themselves.  In addition to the examples from the
+documentation and the automated package checks provided for the R system,
+the `EML` package includes a complete test suite using the testthat
+package [Wickham 2011].  This helps ensures that the software works
+as expected and maintains consistent behavior in any future versions
+of the software package.
+
+[Wickham 2011]: http://vita.had.co.nz/papers/testthat.html
 

### 2014-03-28 inst/doc/pubs/manuscript.Rmd
use explicit sandbox credentials for figshare auth in vignette too

--- 2014-01-28 inst/doc/pubs/reml.Rmd
@@
+---
+layout: preprint
+#layout: review, 11pt
+title: "RNeXML: Parsing and Serializing the Next Generation of Phyloinformatic Data in R"
+author: 
+  - name: Carl Boettiger
+    affiliation: cstar
+    email: cboettig@gmail.com
+    footnote: Corresponding author
+  - name: Karthik Ram 
+    affiliation: Berkeley 
+  - name: Class-Thido Pfaff
+    affiliation: 
+  - name: Matt Jones 
+    affiliation: NCEAS 
+address: 
+  - code: cstar
+    address: | 
+      Center for Stock Assessment Research, 
+      Department of Applied Math and Statistics, 
+      University of California, Mail Stop SOE-2,
+      Santa Cruz, CA 95064, USA
+
+abstract:
+
+---
+
+<!--
 The future belongs to those who can leverage existing data and provide reusable data.  
+-->
+
+1. **Write**. Easy to generate and publish data and metadata using state-of-the-art technology and standards. 
+2.
+**Read**.
+Easy to consume richly structured data and metadata from rapidly growing archives of repositories such as KNB and DataONE, including over 20,000 datasets deposited by Long Term Ecological Research (LTER) areas over the past 30 years
+3. **Manipulate** Easy to manipulate and add rich machine readable annotations describing heterogenous data sets. This in turn facilitates programmatic integration and synthesis of heterogenous data.
 

### 2014-03-28 inst/doc/pubs/manuscript.Rmd
scratching away at manuscript outline

--- 2014-01-28 inst/doc/pubs/reml.Rmd
@@
+---
+layout: preprint
+#layout: review, 11pt
+title: "Ecological Metadata Language interface for R: facilitating synthesis and integration of heterogenous data"
+author: 
+  - name: Carl Boettiger
+    affiliation: cstar
+    email: cboettig@gmail.com
+    footnote: Corresponding author
+  - name: Karthik Ram 
+    affiliation: Berkeley 
+  - name: Class-Thido Pfaff
+    affiliation: Leipzig
+  - name: Matt Jones 
+    affiliation: NCEAS 
+address: 
+  - code: cstar
+    address: | 
+      Center for Stock Assessment Research, 
+      Department of Applied Math and Statistics, 
+      University of California, Mail Stop SOE-2,
+      Santa Cruz, CA 95064, USA
+
+abstract: | 
+  1. Something about the importance of data, data publication, sharing
+     and annotation
+
+  2. Something about EML
+
+  3. The EML package provides automated and scriptable tools to make it
+     easy to generate and publish data and metadata using state-of-the-art
+     technology and publishing standards.
+
+  4. The EML package also makes it easy to consume richly structured data
+     and metadata from rapidly growing archives of repositories such as
+     KNB and DataONE, including over 20,000 datasets deposited by Long
+     Term Ecological Research (LTER) areas over the past 30 years
+
+  5. The EML package facilitates manipulation and integration of
+     heterogeneous data sources by providing simple functions to add rich,
+     machine-readable annotations to almost any data set.  These
+     descriptions can then be used to enable programmatic recombination and
+     synthesis of different data sets, allowing users to both take better
+     advantage of their own data collections and more easily integrate
+     them with others.
+
+---
+
+<!--
 The future belongs to those who can leverage existing data and provide reusable data.  
-
-1. Following metadata standards facilitates reuse
-2. data 
-3. EML provides an easy way to utilize the Ecological Metadata Language to generate, access, and publish data.
+-->
 
@@
 
-Not sure this counts as a section - but we can automatically detect at least the difference between numeric and character data, factors, ordered factors, etc.
+Not sure this counts as a section - but we can automatically detect
+at least the difference between numeric and character data, factors,
+ordered factors, etc.
 
@@
 
-- Examples of filling in contact information, etc using `eml$set`, using an existing EML file as a template, and using an external yaml file.
+- Examples of filling in contact information, etc using `eml$set`,
+using an existing EML file as a template, and using an external yaml file.
 
-- Similar example for `column_metadata` and `unit_metadata` based on an existing EML file (just using `eml_read`).
+- Similar example for `column_metadata` and `unit_metadata` based on an
+existing EML file (just using `eml_read`).
 
@@
 
-<!-- Not sure if we'll actually have/need to have all this in place for this paper?  -->
-Examples of validation against schema.  Examples checking md5sums on files or other corruption checks?  Examples checking species names (taxize?)
+<!-- Not sure if we'll actually have/need to have all this in place for
+this paper?  --> Examples of validation against schema.  Examples checking
+md5sums on files or other corruption checks?  Examples checking species
+names (taxize?)
 
@@
 
-Illustrate a basic example of synthesis.
-While arbitrary synthesis will require semantics, a lab could store their data in EML format and build up there own library of common scripts to extract the data they need.
-As a lab database grows, the same queries can continue to deliver more data.
-By scripting the data access, analyses can be automatically updated as the data set grows.
+Illustrate a basic example of synthesis.  While arbitrary synthesis will
+require semantics, a lab could store their data in EML format and build
+up there own library of common scripts to extract the data they need. As
+a lab database grows, the same queries can continue to deliver more data.
+By scripting the data access, analyses can be automatically updated as
+the data set grows.
 
-Consider simple example in regression between different morphological characters of species, or such.
+Consider simple example in regression between different morphological
+characters of species, or such.
 
@@
 
-Completely automated data integration will require rich semantic information. Consider the simple example of combining the two data tables shown in Figure 2.
+Completely automated data integration will require rich semantic
+information. Consider the simple example of combining the two data tables
+shown in Figure 2.
 
 <!-- Example illustrating the need for semantics to resolve differences -->
-
-### Implementation and architecture
-
-The `EML` package is built for the Ecological Markup Language schema, [version 2.1.1], using the R software environment and the XML package for parsing and generating XML [Temple Lang 2013, XML].
-
-<!-- Possibly cite all dependencies and suggests from the DESCRIPTION --> 
-<!-- Reference any installation and configuration instructions necessary (e.g. API credentials for figshare) -->
 
@@
 
-The `EML` package can perform three levels of validation on EML files written from or read into R: validating the XML, validating against the EML 2.1.1 Schema, and performing
+The `EML` package performs severak levels of validation on EML files
+written from or read into R: validating the XML, and validating against
+the EML schema (based on the declared version), and performing additional
+checks through the EML online validation tool for elements that cannot
+be enforced by schema validation alone.  Validation serves to improve
+consistency in data encoding which facilitates the use of EML across a
+wide variety of platforms and software.
 
-Several mechanisms are in place to evaluate the quality of the `EML` package functions themselves.
-In addition to the examples from the documentation and the automated package checks provided for the R system, the `EML` package includes a complete test suite using the testthat package [Wickham 2011].
-
-[Wickham 2011]: http://vita.had.co.nz/papers/testthat.html
+Several mechanisms are in place to evaluate the quality of the `EML`
+package functions themselves.  In addition to the examples from
+the documentation and the automated package checks provided for the
+R system, the `EML` package includes a complete test suite using the
+`testthat` package [Wickham 2011]. Package development on Github includes
+continuous integration through the `travis` platform, in which the full
+test suite is run every time the code is changed.  This not only gives
+users confidence that the development copy can be successfully installed,
+but should assist other developers seeking to contribute to the package
+to test that their proposed changes have not negatively impacted any of
+the existing code base.
 
@@
 
-The `EML` package is available through the CRAN repository under a CC0 license.
-At the time of writing, the current version is `r packageDescription("EML", fields="Version")`.
-The current development branch can be found on Github at [https://github.com/ropensci/EML].
-See the NEWS file for changes in recent versions.
-Users are encouraged to submit bug reports or feature requests in the issues log.
-Contact ropensci-discuss@googlegroups.com or the package maintainer for help.
+The `EML` package is available through the CRAN repository under
+a CC0 license. At the time of writing, the current version is `r
+packageDescription("EML", fields="Version")`. The current development
+branch can be found on Github at [https://github.com/ropensci/EML].
+See the NEWS file for changes in recent versions. Users are encouraged
+to submit bug reports or feature requests in the issues log.  Contact
+`ropensci-discuss@googlegroups.com` or the package maintainer for help.
 
-There are many promising directions for further development of the `EML` package.
-The project's Milestones page provides a list of steps currently planned or under consideration.
-In particular, these milestones include increased support for interactive mode of entering data (wizards), and support for integration of semantics defined through the Web Ontology language (OWL), as discussed in the Challenges section.
+There are many promising directions for further development of the `EML`
+package.  The project's Milestones page provides a list of steps currently
+planned or under consideration.  In particular, these milestones include
+increased support for interactive mode of entering data (wizards), and
+support for integration of semantics defined through the Web Ontology
+language (OWL), as discussed in the Challenges section.
 
@@
 
-This project was supported in part by a grant from the Alfred P Sloan Foundation (CB and KR), NSF Grant DBI-1306697 (CB), ... `EML` is part of the rOpenSci project, [http://ropensci.org].
-
-OLD OUTLINE
-----------------------------------------------------------------------------------------------
-
-It's all metadata.
-
-Share 
-Metadata
-raw data
-formats: file formats, table formats, cell formats 
-good null values
-easy to combine (context data)
-quality control
-repository
-license
-
-Abstract
-
-1. Recording good metadata is at the heart of effective data synthesis and reuse
-2.
-
-Vertically integrated repositories vs metadata driven repositories.
-
-Data synthesis
-
-## Why EML?
-
-But specifying metadata is typically onerous, time-consuming and error prone.
-
-Such definitions are also not machine readable.
-
-Or at least, it doesn't have to be that way.  We just need better tools.
-
-## Why REML?
-
-* The Ecological Metadata Language (EML) has been around for a while now, but has so far seen limited use.
-
-* R is a widely adopted among the community, many researchers already familiar with reading and writing data from R.
-
-* R provides a rich set of data types and constructs that can be mapped into EML (e.g. integers, floats, ordered/nominal factors, character strings)
-
-### Current state of EML
-
-* Generation of EML is limited.  e.g. to LTER centers with informatics support. 
-* Applications of existing EML more limited still (?)
-
-* Software to work with EML is limited to tools with low adoption rates across the community (e.g. Kepler).
-
- Researchers are isolated from the metadata encoding, instead providing plain-text descriptions in MS Word documents.  Allows for low barrier of entry but also makes a barrier to richer encoding.
-
-### Promise of REML
-
-### Getting Started: A minimal example.
-
-Imagine we are working with a `data.frame` in R analgous to Table 1.
-
-<!--Writing a data.frame into EML -->
-
-Writing this table into EML
-
-### Authoring tools
-
-### Parsing tools
-
-### Publishing tools
+This project was supported in part by a grant from the Alfred P Sloan
+Foundation (CB and KR), NSF Grant DBI-1306697 (CB), ... `EML` is part
+of the rOpenSci project, [http://ropensci.org].

### 2014-03-31 inst/doc/pubs/manuscript.Rmd
a bit of writing on manuscript

@@
 abstract: | 
-  1. Something about the importance of data, data publication, sharing
-     and annotation
+  1. _Write something about the importance of data, data publication, sharing
+     and annotation_
 
-  2. Something about EML
+  2. _Write something about EML_
 
@@
 -->
+
+Heterogeneous Data
+[Jones 2006], [Green 2005]
 
@@
 Biocomplexity (KNB) is one example of a metadata-based ecological data 
-repository [Berkley 2001], [KNB].  The metadata approach preserves the original data
+repository [Jones 2001], [KNB].  The metadata approach preserves the original data
 model from the collecting study. Rather than requiring that the raw data
@@
 *  It's not just a problem of sharing culture 
-* Sharing is not effective: close competitors can scoop you, but broader reach can't be realized without good metadata 
-* [Strasser survey showing lagging data management practices]
-* [Stodden survey showing effort in curating data is a barrier]
+* Sharing is not effective: close competitors can still scoop you, but broader reach can't be realized without good metadata 
+* [Strasser] survey showing lagging data management practices
+* [Stodden 2010] survey showing effort in curating data is a barrier
 
@@
 [KNB]: http://knb.ecoinformatics.org
-[Berkley 2001]: http://dx.doi.org/10.1109/4236.957896
+[Jones 2001]: http://dx.doi.org/10.1109/4236.957896
 [Michener 1997]: http://dx.doi.org/10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2

### 2014-04-22 inst/doc/pubs/one-simple-way.Rmd

@@
 
-## 3. Provide unprocsessed data
+## 3. Provide unprocessed data
 
@@
 
+Data tables are easiest to manipulate and combine when users consistently
+enter variables as columns and observations as rows (sometimes called "long" format).  Partly because this format
+may require duplicated entries, such as repeating the value of the "site" when
+multiple observations have been made on the same site, many researchers are 
+still tempted to present data in "wide" format, with observations for a single 
+site added as additional rows.  While potentially convenient for field entry,
+the long format is more desirable for most future analysis or reuse.
+
+The EML package addresses this by both working in the R an environment where
+it is easy and routine to transform between "long" and "wide" formats, while
+also giving the user the flexibility to simply document the data in whatever
+format they feel works best for their analysis.
+
 - Standard cell formats
@@
 
-Likewise, the package detects the appropriate metadata based on the R 
-representation (such as factor, ordered factor, numeric, integer, character, or Date),
-and writes the appropriate metadata for this format.
+Likewise, the package detects the appropriate metadata based on the
+R representation (such as factor, ordered factor, numeric, integer,
+character, or Date), and writes the appropriate metadata for this format.
 
@@
 
-The detailed column and unit metadata serve the pupose described in this section of White _et al_.   
-Machine-readable metadata,
+As White _et al._ discuss, providing the kind of detailed column and unit
+metadata we have already discussed, as well as the standard formats addressed
+above, already make data easy to combine.
+
+The EML package lets the user go beyond this in being able to automate
+the combining of datasets based on the metadata alone. The key to this process
+is to replace a typical programmatic script for subsetting or combining data sets 
+which is based on implicit knowledge of the data structure (e.g. "select value from column 5 whenever column 2 has value = AAA") into one that is based only 
+on the metadata description: (e.g. "select the species_population_counts from the Sacramento_River site").  The latter has the advantage of not only 
+being easier to interpret, but can also be independent of how the data is structured.
+
+This is already possible with R, but requires explicit use of the column name.  
+As the column name may already have a specific value given to it by the researcher
+that serves some other purpose, it would be against the objectives of EML to insist that it have a specific name (e.g.
+all columns containing species names must be called "spp", which among other issues would break some code requiring that
+the column instead be called "species".)  To avoid renaming the columns, we simply
+need to use a consistent description of the column in the metadata instead.
 

### 2014-04-22 inst/doc/pubs/manuscript.Rmd
update affiliations

@@
       Santa Cruz, CA 95064, USA
+  - code: Berkeley
+    address: |
+      Department of Environmental Science and Policy Management,
+      University of California, Berkeley, 130 Mulford Hall,
+      Berkeley, CA 94720. USA
+  - code: Leipzig
+    address: 
+  - code: NCEAS
+    address |
+      National Center for Ecological Analysis and Synthesis,
+      735 State Street, Suite 300, 
+      Santa Barbara, CA 93101
 

### 2014-05-10 manuscripts/components/pandoc.R
adopt template layout: vignettes in vignettes/, manuscripts in manuscripts

@@ -1,0 +1,17 @@
+#!/usr/bin/Rscript
+
+## NOTE: Arguments to pandoc are set in the config file, pandoc_config.txt
+## Also note that config file path is relative to manuscripts/ directory,
+## not relative to the components/ directory. 
+
+
+args <- commandArgs(TRUE)
+file <- args[1]
+library("knitr")
+
+## Call pandoc
+knitr::pandoc(file, format="latex", config="components/pandoc_config.txt")
+
+# Note that path is relative to where the Rscript is called
+# (manuscripts/), not where the source script lives (manuscripts/components)
+

### 2014-06-19 R/dataTable_methods.R
allow passing col.defs and unit.defs directly addresses #113

@@ -1,93 +1,0 @@
-
-#' constructor helper function for data table
-#' 
-#' constructor helper function for data table
-#' @param dat a data.set object, data.frame object
-#' @param meta a list of meta information organized by column.  
-#'  Elements of the list are another list, in which: 
-#'  - the first object is the column name, 
-#'  - then the column description, 
-#'  - then the unit definition
-#' @param title the the title of the dataTable 
-#' @param description a description of the dataTable
-#' @param filename the filename of the csv file created
-#' @param ... additional arguments (not yet implemented)
-#' @return a dataTable element
-#' @export
-#' @include dataTable_classes.R
-eml_dataTable <- function(dat, 
-                          meta = NULL, 
-                          title = "data table", 
-                          description = character(0), 
-                          filename = character(0), 
-                          ...){
-
-## FIXME title should be called entityName, or maybe just "name"
-  
-  id <- eml_id()
-
-  dataTable <- new("dataTable",
-                  id = id[["id"]],
-                  system = id[["system"]],
-                  scope = id[["scope"]],
-                  entityName = title,
-                  entityDescription = description,
-                  attributeList = eml_attributeList(dat, meta),  
-                  physical = eml_physical(dat, filename = filename),
-                  caseSensitive = 'yes',
-                  numberOfRecords = dim(dat)[1])
-}
-
-
-
-eml_attributeList <- function(dat, meta = NULL){
-
-  ## Handle other types 
-  if(is(dat, "data.set")) 
-    meta <- get_metadata(dat)
-  if(is.null(meta))
-    meta <- metadata_wizard(dat)
-  meta <- detect_class(dat, meta)
-
-#  as(meta, "attributeList")
-
-  uid <- eml_id()
-  new("attributeList",
-      id = uid[["id"]], 
-      attribute = new("ListOfattribute", 
-                      lapply(meta, eml_attribute, uid[["id"]])))
-
-}
-
-
-### Helper Method updates metadata list with class #######################
-#' Helper function to attach classes to metadata
-#' @param dat a data frame
-#' @param meta a list containing attribute-level metadata for each column.  
-#' each element of list corresponds to a consecutive column, and gives a list containing
-#' the column name (as appearing in the data-frame), the column description (e.g. from col.defs)
[…]

### 2014-06-19 R/eml_attributeList.R

@@ -1,0 +1,161 @@
+#' eml_attributeList
+#'
+#' eml_attributeList
+#' 
+#' @param dat a data.frame (or a data.set which has col.defs and unit.defs)
+#' @param meta a list of column and unit metadata, OR
+#' @param col.defs the column definitions (if not given in meta)
+#' @param unit.defs the unit definitions (if not given in meta)
+#' @param col.classes the column classes of dat, if dat not given (otherwise inferred from dat)
+#'
+#' @export
+eml_attributeList <- function(dat = NULL, 
+                              meta = NULL,
+                              col.classes = sapply(dat, class),
+                              col.defs = NULL,
+                              unit.defs = NULL){
+
+
+  if(!is.null(col.defs) && !is.null(unit.defs) && 
+     is.null(meta) && is(dat, "data.frame")){
+    column_names <- names(dat)
+    meta <- lapply(1:length(col.defs), 
+                   function(i) list(
+                     column_names[[i]], 
+                     col.defs[[i]], 
+                     unit.defs[[i]]))
+
+  } else if(is(dat, "data.set")) { ## Handle other types 
+    meta <- get_metadata(dat)
+
+  } else  if(is.null(meta) && is.null(unit.defs)) {
+    meta <- metadata_wizard(dat)
+  }
+  
+  ## Handle column classes
+  meta <- detect_class(meta, col.classes)
+
+#  as(meta, "attributeList")
+
+  uid <- eml_id()
+  new("attributeList",
+      id = uid[["id"]], 
+      attribute = new("ListOfattribute", 
+                      lapply(meta, eml_attribute, uid[["id"]])))
+}
+
+
+#' Helper function to attach classes to metadata
+#' @param meta a list containing attribute-level 
+#' metadata for each column.  see details. 
+#' @param col.classes a character vector of column classes from the data.  
+#' @details each element of meta list corresponds to a consecutive
+#' column, and gives a list containing the column name (as appearing in the
+#' data-frame), the column description (e.g. from col.defs) and the unit defs
+#' (e.g. from unit.defs) as  a (named) character vector.
+detect_class <- function(meta, col.classes){
+  for(i in 1:length(meta)){
+    if(length(col.classes) == 1)
+      cls <- col.classes
+    else if(length(col.classes) < 1)
+      cls <- "character"
+    else 
+      cls <- col.classes[[i]]
+
+    meta[[i]][[4]] <- map(cls) 
+  }
[…]

### 2014-06-19 R/eml_dataset.R

@@ -1,0 +1,80 @@
+
+eml_dataset <- 
+  function(dat = NULL,
+           meta = NULL, 
+           title = "EML_metadata",
+           creator = NULL, 
+           contact = NULL, 
+           coverage = eml_coverage(scientific_names = NULL,
+                                   dates = NULL,
+                                   geographic_description = NULL,
+                                   NSEWbox = NULL),
+           methods = new("methods"),
+           col.defs = NULL,
+           unit.defs = NULL,
+           ref_id = eml_id(),
+           ...){
+
+  if(is.null(creator))
+    creator <- get("defaultCreator", envir=EMLConfig)
+  if(is.null(contact))
+    contact <- get("defaultContact", envir=EMLConfig)
+
+  ## Coerce character string persons into EML representations
+  if(!is.null(dat)) # this is written only into dataset nodes 
+    who <- contact_creator(contact = contact, 
+                           creator = creator)
+
+  ## Format methods appropriately 
+  if(is(methods, "character")){
+    methods <- new("methods", 
+                   methodStep = c(new("methodStep", 
+                                      description = methods)))
+  }
+
+  if(is(dat, "dataset")){ # pre-built dataset object
+    dataset <- dat 
+
+  } else if(is(dat, "data.frame")){  # data.set class, 
+
+    dataset = new("dataset",
+                  id = ref_id[["id"]],
+                  title = title, # required 
+                  creator = who$creator,
+                  contact = who$contact,
+                  coverage = coverage,
+                  methods = methods, 
+                  dataTable = c(eml_dataTable(dat = dat,
+                                              title = title,
+                                              meta = meta,
+                                              col.defs = col.defs,
+                                              unit.defs = unit.defs)),
+                  ...)
+
+
+  } else if(is(dat, "dataTable")){
+
+    dataset = new("dataset",
+                  id = ref_id[["id"]],
+                  title = title, # required 
+                  creator = who$creator,
+                  contact = who$contact,
+                  coverage = coverage,
+                  methods = methods, 
+                  dataTable = c(dat),
+                  ...)
+  } else if(is(dat, "ListOfdataTable")) {
+    dataset = new("dataset",
+                  id = ref_id[["id"]],
+                  title = title, # required 
+                  creator = who$creator,
+                  contact = who$contact,
+                  coverage = coverage,
+                  methods = methods, 
+                  dataTable = dat,
+                  ...)
+  }
+  dataset
+}
+
+

### 2014-06-19 R/eml_define_unit.R

@@ -1,81 +1,4 @@
-
-
-eml_attribute <- function(from, ...){
-        new("attribute", 
-            attributeName = from[[1]],
-            attributeDefinition = from[[2]],
-            measurementScale = eml_measurementScale(from[[3]], from[[4]], ...))
-}
-
-
-eml_measurementScale <- function(unit.def, col.class, attrListID = NULL){
-
-  ## FIXME HACK.  Consider more explicit R-based col.class, 
-  if(col.class == "nominal" && length(unit.def) == 1)
-    new("measurementScale", 
-        nominal = new("nominal", 
-                      nonNumericDomain = new("nonNumericDomain",
-                                             textDomain = new("textDomain", 
-                                                              definition = unit.def))))
-  else 
-  switch(col.class,
-         nominal = new("measurementScale", 
-                       nominal = new("nominal", 
-                                     nonNumericDomain = eml_nonNumericDomain(unit.def))), 
-         ordinal = new("measurementScale", 
-                       ordinal = new("ordinal", 
-                                     nonNumericDomain = eml_nonNumericDomain(unit.def))), 
-         ratio = new("measurementScale", 
-                     ratio = new("ratio",
-                                 numericDomain = new("numericDomain", 
-                                                     numberType = "real"), #FIXME get numberType properly
-                                 unit = eml_unit(unit.def[[1]]))), 
-         # FIXME has precision and numericDomain@numberType, numericDomain@boundsGroup
-         dateTime = new("measurementScale", 
-                        dateTime = eml_dateTime(unit.def[[1]]) ), 
-         interval = new("measurementScale", 
-                        interval = new("interval",
-                                       numericDomain = new("numericDomain", 
-                                                           numberType = "real"), 
-                                       unit = eml_unit(unit.def[[1]])))
-         )
-}
-
-eml_dateTime <- function(formatstring){ # FIXME perform validation. Accept R notation (%Y -> YYYY)
-  new("dateTime", formatString = formatstring)
-}
-eml_nonNumericDomain <- function(from){
-new("nonNumericDomain", 
-    enumeratedDomain = 
-      new("enumeratedDomain", 
-          codeDefinition = 
-            new("ListOfcodeDefinition", 
-                lapply(names(from), 
[…]

### 2014-06-19 inst/examples/ex3.R

@@ -1,20 +1,44 @@
-# Add example from ropenfisheries showing how factor definitions can come from two columns, rather than long entry...
-# Date should support R format strings (%Y, %m, %d) as well as EML strings (YYYY, MM, DD)
+library(EML)
+dat <- data.frame(river = factor(c("SAC",  
+                                   "SAC",   
+                                   "AM")),
+                  spp   = c("Oncorhynchus tshawytscha",  
+                            "Oncorhynchus tshawytscha", 
+                            "Oncorhynchus kisutch"),
+                  stg   = ordered(c("smolt", 
+                                    "parr", 
+                                    "smolt"), 
+                                  levels=c("parr", 
+                                           "smolt")), # => parr < smolt
+                  ct    = c(293L,    
+                            410L,    
+                            210L),
+                  day   = as.Date(c("2013-09-01", 
+                                    "2013-09-1", 
+                                    "2013-09-02")),
+                  stringsAsFactors = FALSE)
 
 
-# The weird thing about this is that it defines a loose object for unit.defs.  perhaps we should define as an explicit class(?!) 
 
-# numeric: 
-#   required: unit
-#   optional: 
-#     - type (natural, integer, real). If not given will be inferred from the column
-#     - precision (numeric)
-#     - bounds 
-#     - interval? (logical)  Otherwwise assumed to be aratio
+col.defs <- c("River site used for collection",
+              "Species scientific name",
+              "Life Stage", 
+              "count of live fish in traps",
+              "day traps were sampled (usually in morning thereof)")
 
-# dateTime
-#   required: format format string
-#   optional:
-#     - precision
+
+unit.defs <- list(
+  c(SAC = "The Sacramento River",     # Factor 
+    AM = "The American River"),
+ "Scientific name",                   # Character string 
+  c(parr = "third life stage",        # Ordered factor 
+    smolt = "fourth life stage"),
+  c(unit = "number", 
+    precision = 1, 
+    bounds = c(0, Inf)),              # Integer
+  c(format = "YYYY-MM-DD",            # Date
+    precision = 1))
+
+
 
 

### 2014-06-19 tests/testthat/test_create.R

@@ -1,114 +1,0 @@
-context("generating S4 from R dataframe")
-## This file tests constructors for the S4 objects 
-## using the expected R input formats 
-
-
-## FIXME add expect_that to the tests!
-
-require(EML)
-eml_reset_config()
-dat = data.frame(river = c("SAC",  "SAC",   "AM"),
-                 spp   = c("king",  "king", "ccho"),
-                 stg   = c("smolt", "parr", "smolt"),
-                 ct    = c(293L,    410L,    210L))
-metadata <- 
-  list("river" = list("river",
-                      "River site used for collection",
-                      c(SAC = "The Sacramento River", 
-                        AM = "The American River")),
-       "spp" = list("spp",
-                    "Species common name", 
-                    c(king = "King Salmon", 
-                      ccho = "Coho Salmon")),
-       "stg" = list("stg",
-                    "Life Stage", 
-                    c(parr = "third life stage", 
-                      smolt = "fourth life stage")),
-       "ct"  = list("ct",
-                    "count", 
-                    "number"))
-
-
-
-
-
-test_that("We can coerce factor definitions to nominal", {
-  as(metadata$river[[3]], "nominal")
-})
-
-test_that("We can coerce list to to attribute", {
-  metadata <- EML:::detect_class(dat, metadata)
-  as(metadata$river, "attribute")
-})
-
-test_that("We can coerce metadata to attributeList", {
-  metadata <- EML:::detect_class(dat, metadata) ## Needs metadata with class info from data
-  as(metadata, "attributeList") 
-})
-
-test_that("Generator for dataset containing generator for dataTable", {
-  metadata <- EML:::detect_class(dat, metadata)  ## Needs metadata with class info from data
-  set <- new("dataset")
-  set@dataTable = new("ListOfdataTable", list(new("dataTable", attributeList = as(metadata, "attributeList"))))
-})
-
-
-test_that("Constructor function for physical works given data.frame", {
-  EML:::eml_physical(dat, filename="tmp.csv") 
-  unlink("tmp.csv")
-})
-
-## Constructor functions handle the detect_class internally:
-
-test_that("Constructor function for dataTable works given data.frame and metadata list", {
-  EML:::eml_dataTable(dat, metadata, filename="tmp.csv") 
-  unlink("tmp.csv")
-}) 
-
-
-## FIXME how about a dataset level test?
-
-
-
-
-
-
-
-
-
-
[…]

### 2014-06-19 tests/test-all.R
remove data.set focused examples and methods

@@ -1,5 +1,5 @@
 if (packageVersion("testthat") >= "0.7.1.99") {
-    library(testthat)
+  library(testthat)
   test_check("EML")
 } else {
   library(testthat)

### 2014-06-20 R/eml_attributeList.R
Lots of small updates based around avoiding data.set
- Updated a handful of tests to remove depricated methods.
- Replaced the extraction method used by eml_get(eml, "data.frame").
The new approach is built more around functions than S4 methods.
Most importantly, it handles the type casting of columns more
explicitly, which requires comparing different parts of the metadata
(e.g. dateTime@formatString, not only ratio vs interval but integer
vs numeric, etc.)
- Updated the top-level interface to support col.defs and unit.defs all
  the way down.

@@ -3,36 +3,45 @@
 #' eml_attributeList
 #' 
 #' @param dat a data.frame (or a data.set which has col.defs and unit.defs)
-#' @param meta a list of column and unit metadata, OR
 #' @param col.defs the column definitions (if not given in meta)
 #' @param unit.defs the unit definitions (if not given in meta)
-#' @param col.classes the column classes of dat, if dat not given (otherwise inferred from dat)
-#'
+#' @param col.classes column classes, primarily for use if dat is 
+#' a path to a csv rather than a native R object. Optional otherwise.    
+#' @param col.names names of the columns, if dat is a csv file path.  
 #' @export
-eml_attributeList <- function(dat = NULL, 
-                              meta = NULL,
-                              col.classes = sapply(dat, class),
-                              col.defs = NULL,
-                              unit.defs = NULL){
+eml_attributeList <- 
+function(dat = NULL, 
+         col.defs = NULL,
+         unit.defs = NULL,
+         col.classes = NULL,
+         col.names = NULL){
+
+  if(is(dat, "data.frame")){
+    if(is.null(col.classes))
+      col.classes <-  sapply(dat, class)
+    if(is.null(col.names))
+      col.names <-  names(dat)
+  } else if (is(dat, "character") && is.null(col.names)){
+    tmp <- read.csv(dat, nrows=3, header=TRUE)
+    col.names <- names(tmp)
+  }
 
 
-  if(!is.null(col.defs) && !is.null(unit.defs) && 
-     is.null(meta) && is(dat, "data.frame")){
-    column_names <- names(dat)
-    meta <- lapply(1:length(col.defs), 
-                   function(i) list(
-                     column_names[[i]], 
-                     col.defs[[i]], 
-                     unit.defs[[i]]))
-
-  } else if(is(dat, "data.set")) { ## Handle other types 
+  if(is(dat, "data.set")) 
     meta <- get_metadata(dat)
 
-  } else  if(is.null(meta) && is.null(unit.defs)) {
-    meta <- metadata_wizard(dat)
-  }
-  
-  ## Handle column classes
+  if(!is.null(col.defs) && !is.null(unit.defs) && 
+     !is.null(col.names)){
+      meta <- lapply(1:length(col.defs), 
+                     function(i) list(
+                       col.names[[i]], 
+                       col.defs[[i]], 
+                       unit.defs[[i]]))
+  } #else if(is.null(meta)) {
[…]

### 2014-06-20 R/eml_physical.R

@@ -1,27 +1,61 @@
 ####### Internal function to coerce data.frame to physical #######
 
-## Custom generator. Consider making int an S4 method?  
-## FIXME filename should follow https://github.com/ropensci/EML/issues/106
-eml_physical <- function(dat, filename=character(0), ...){
-  if(length(filename) == 0 | is.null(filename)){
-    longID <- gsub('.*:(.*)', '\\1', eml_id()[["id"]])
-    shortID <- gsub("(.{6}).+", "\\1", longID)
-    filename <- paste("table_", shortID, ".csv", sep = "")
+## Custom generator. 
+
+#' eml_physical constructor
+#' 
+#' eml_physical constructor
+#' @param dat a data.frame or path to existing csv file
+#' @param filename the filename to be used. Ignored if `dat` 
+#' is a path to a file.  Can include the full path (e.g. if
+#' file is to be written to a different directory).
+#' @param characterEncoding string defining the encoding used. will guess based on R's configuration (sensible for data.frames or csv files produced by R, but not otherwise).
+#' @param ... additional arguments passed to eml_dataFormat, eml_distribution or write.csv. 
+eml_physical <- 
+function(dat,
+         filename = character(), # Full path, file extension
+         characterEncoding = guess_encoding(),
+         ...){
+
+  # FIXME base this on entityName?  
+  if(length(filename) == 0 || is.null(filename)){
+
+    generate_filename("table")
   }
-  suppressWarnings(write.csv(dat, file = filename, row.names=FALSE, ...)) # don't care about warning in coercion of data.set to data.frame
+
+
+  if(gsub(".*(\\.csv)", "\\1", filename) != ".csv")
+    filename <- paste(filename, "csv", sep=".")
+
+  ## handle existing csv path
+  if(is.character(dat)){
+    if(file.exists(dat)){
+      filename <- dat
+      size <- file.info(dat)$size # in bytes
+    }
+  } else if(is(dat, "data.frame")){ 
+    suppressWarnings(# don't warn in coerce of data.set 
+      write.csv(dat, 
+                file = filename, 
+                row.names=FALSE)) 
+  }
+
+
+  dirname <- dirname(filename)
+  objectName <- basename(filename)
+
+
   new("physical", 
-      objectName = filename, 
+      objectName = objectName, 
       size = as(object.size(dat), "numeric"), 
-      characterEncoding = guess_encoding(),
-      dataFormat = new("dataFormat", 
-                       textFormat = new("textFormat", 
[…]

### 2014-06-20 R/physical_methods.R

@@ -1,6 +1,6 @@
 #' @include physical_classes.R
-
-########### extract methods for physical
+#' 
+NULL
 
 setMethod("extract", signature("physical", "character"),
           function(from, using=NA){
@@ -10,8 +10,10 @@ setMethod("extract", signature("physical", "character"),
             ## FIXME use the EML to determine read.csv options
             dat <- read.csv(filepath(from))  #colClasses=using)
           })
-# If col_classes(object) returns a list, means that colClasses are not simple types (e.g. c("ordered", "factor") or "NULL")
-# See read.csv for details.  Meanwhile, if we get a list, just read in ignoring the declared colClasses
+
+# If col.classes(object) returns a list, means that col.classes are not simple types (e.g. c("ordered", "factor") or "NULL")
+# See read.csv for details.  Meanwhile, if we get a list, just read in ignoring the declared col.classes
+
 setMethod("extract", signature("physical", "list"),
           function(from, using=NA){
             ## FIXME use the EML to determine read.csv options
@@ -39,7 +41,7 @@ setMethod("extract", signature("physical"),
 setGeneric("filepath", function(from) standardGeneric("filepath"))
 setMethod("filepath", signature("physical"),
           function(from){
-            path <- paste0(from@dirname, from@objectName)
+            path <- catpaths(from@dirname, from@objectName)
             if(file.exists(path))
               path
             else {
@@ -55,8 +57,7 @@ setMethod("filepath", signature("physical"),
 
 ### Coerce data.frame to physical (csv) using constructor function:
 
-setAs("data.frame", "physical", function(from)
-      eml_physical(from)) 
+setAs("data.frame", "physical", function(from) eml_physical(from)) 
 
 
 

### 2014-06-20 demo/ex1.R
add example to demo

@@ -1,0 +1,43 @@
+dat <- data.frame(river = factor(c("SAC",  
+                                   "SAC",   
+                                   "AM")),
+                  spp   = c("Oncorhynchus tshawytscha",  
+                            "Oncorhynchus tshawytscha", 
+                            "Oncorhynchus kisutch"),
+                  stg   = ordered(c("smolt", 
+                                    "parr", 
+                                    "smolt"), 
+                                  levels=c("parr", 
+                                           "smolt")), # => parr < smolt
+                  ct    = c(293L,    
+                            410L,    
+                            210L),
+                  day   = as.Date(c("2013-09-01", 
+                                    "2013-09-1", 
+                                    "2013-09-02")),
+                  stringsAsFactors = FALSE)
+
+
+
+col.defs <- c("River site used for collection",
+              "Species scientific name",
+              "Life Stage", 
+              "count of live fish in traps",
+              "day traps were sampled (usually in morning thereof)")
+
+
+unit.defs <- list(
+  c(SAC = "The Sacramento River",     # Factor 
+    AM = "The American River"),
+ "Scientific name",                   # Character string 
+  c(parr = "third life stage",        # Ordered factor 
+    smolt = "fourth life stage"),
+  c(unit = "number", 
+    precision = 1, 
+    bounds = c(0, Inf)),              # Integer
+  c(format = "YYYY-MM-DD",            # Date
+    precision = 1))
+
+
+
+

### 2015-11-06 R/eml_knb.R
avoid require() calls

@@ -17,21 +17,12 @@ eml_knb <- function(file,
                     permanent = FALSE,
                     public = TRUE,
                     mn_nodeid = "urn:node:KNB",
-                    cli = D1Client("PROD", mn_nodeid)){
+                    cli = dataone::D1Client("PROD", mn_nodeid)){
 
-  # require call leads to a warning when package is not on the SUGGESTS list.  A problem for TRAVIS, not for CRAN.  
-  success <- require("dataone", character.only = TRUE, quietly = TRUE)
-  if(!success){
-    message("dataone package not found. Attempting to install")
-    install.packages("dataone")
-    success <- require("dataone", character.only = TRUE, quietly = TRUE)
-     if(!success)  
-      stop("The dataone package must be installed to publish data to the KNB")
-  }
   # Declare the metadata format ## FIXME get namespace from the file 
   if(permanent == FALSE){
     mn_nodeid <- "urn:node:mnDemo5" # A Development server for testing
-    cli <- D1Client("DEV", mn_nodeid)
+    cli <- dataone::D1Client("DEV", mn_nodeid)
 
   }
 
@@ -56,15 +47,15 @@ eml_knb <- function(file,
 
 
   # Query the object to show its identifier
-  csv_pid <- getIdentifier(csv_object)
+  csv_pid <- dataone::getIdentifier(csv_object)
   message(paste("CSV file identifier is :", csv_pid))
-  eml_pid <- getIdentifier(eml_object)
+  eml_pid <- dataone::getIdentifier(eml_object)
   message(paste("EML file identifier is :", eml_pid))
 
   if(public){
   # Set access control on the data object to be public
-    setPublicAccess(csv_object)
-    setPublicAccess(eml_object)
+    dataone::setPublicAccess(csv_object)
+    dataone::setPublicAccess(eml_object)
     if (canRead(eml_object,"public")) {
       message("successfully set public access");
     } else {
@@ -75,12 +66,12 @@ eml_knb <- function(file,
   # Assemble our data package containing both metadata and data
   pkg.id <- paste0(package_id, "_package")
   data.package <- new("DataPackage", packageId=pkg.id)
-  addData(data.package, csv_object)
-  addData(data.package, eml_object)
-  insertRelationship(data.package, package_id, csv["id"])
+  dataone::addData(data.package, csv_object)
+  dataone::addData(data.package, eml_object)
[…]

### 2015-12-28 R/entity_classes.R
support yaml-style show/print method for all EML classes

@@ -8,7 +8,8 @@ setClass("entityGroup",
                    physical = "physical",
                    coverage = "coverage",
                    methods = "methods",
-                   additionalInfo = "character"))
+                   additionalInfo = "character"),
+         contains = "eml.2.1.1")
 
 setAs("XMLInternalElementNode", "entityGroup",  function(from) emlToS4(from))
 setAs("entityGroup", "XMLInternalElementNode",   function(from) S4Toeml(from))

### 2015-12-28 R/software_classes.R

@@ -13,7 +13,8 @@
 # See http://www.loc.gov/standards/iso639-2/ e.g. ISO639-2
 setClass("eml_language", 
          slots = c(languageValue = "character", 
-                   languageCodeStandard = "character")) 
+                   languageCodeStandard = "character"),
+         contains = "eml.2.1.1") 
 setAs("XMLInternalElementNode", "language",   function(from) emlToS4(from))
 setAs("language", "XMLInternalElementNode",   function(from) S4Toeml(from))
 
@@ -22,7 +23,8 @@ setAs("language", "XMLInternalElementNode",   function(from) S4Toeml(from))
 # Not sure this is correctly defined? Simple types 
 setClass("action", 
          slots = c(install = "character",
-                   assert = "character"))
+                   assert = "character"),
+         contains = "eml.2.1.1")
 setAs("XMLInternalElementNode", "action",   function(from) emlToS4(from))
 setAs("action", "XMLInternalElementNode",   function(from) S4Toeml(from))
 
@@ -31,7 +33,8 @@ setClass("software") # Dummy declarations are now required..
 
 setClass("dependency",
          slots = c("action" = "action",  # required
-                   "software" = "software")) # required 
+                   "software" = "software"),
+         contains = "eml.2.1.1") # required 
 setAs("XMLInternalElementNode", "dependency",   function(from) emlToS4(from))
 setAs("dependency", "XMLInternalElementNode",   function(from) S4Toeml(from))
 
@@ -56,7 +59,8 @@ setClass("implementation",
                    programmingLanguage = "character", # optional unbounded
                    checksum = "character"  # optional  
 #                   dependency = "ListOfdependency" # optional  unbounded
-                   ))
+                   ),
+          contains = "eml.2.1.1")
 
 setAs("XMLInternalElementNode", "implementation",   function(from) emlToS4(from))
 setAs("implementation", "XMLInternalElementNode",   function(from) S4Toeml(from))

### 2015-12-29 notebook/2015-12-28-eml-explore.Rmd
example notes inspired by @lwasser

---
output:
   md_document:
     variant: markdown_github
---

```{r eval=FALSE}
devtools::install_github(c("hadley/purrr", "ropensci/EML"))
```

```{r}
library("EML")
library("purrr")
library("dplyr")
```

```{r}
obj <- eml_read("http://harvardforest.fas.harvard.edu/data/eml/hf001.xml")
```

```{r}
stuff <- obj@dataset@dataTable[[4]]@attributeList@attribute

stuff[[1]]
```

R has summarized the output using the [YAML](https://en.wikipedia.org/wiki/YAML) format, which is a bit more human readable than the XML format that EML language uses.  YAML uses a `key: value` structure, where the `key` corresponds to the name of a tag/element (or occassionally an XML attribute instead of an XML tag, but attributes are used rarely in EML, mostly just for ids) in the EML language, followed by its value.  Some of these don't appear to have a value, like `attribute:`, but look closely and you'll see the next two keys are indented further.  This means that an `attribute` tag consists of two child tags, `attributeName` and `attributeDefinition`.  We can see this in the subsetting, as

```{r}
slotNames(stuff[1])

stuff[[1]]@attributeName
stuff[[1]]@attributeDefinition
```

So, we want to see all the `attributeName`s and `attributeDefinition`s for Table 4.  We could just print `stuff`, but that is much to verbose since it has all that stuff about `measurementScale` etc that we aren't interested in right now.  So, we use a little R scripting to extract just the information we want:

```{r}
purrr::map_df(stuff, function(x) data.frame(column.name = x@attributeName, definition = x@attributeDefinition))

# Or identically with dplyr alone:
#dplyr::bind_rows(lapply(stuff, function(x) data.frame(column.name = x@attributeName, definition = x@attributeDefinition)))

do.call(rbind, lapply(stuff, function(x) data.frame(column.name = x@attributeName, definition = x@attributeDefinition)))

```

This creates a `data.frame` with two columns: the name of every column in Table 4, and a description of what each of those 29 columns contains.  Similarly we can create a `data.frame` that summarizes all of the tables in the dataset:

```{r}
purrr::map_df(obj@dataset@dataTable, function(dt) 
  data.frame(table.id = dt@id,
             table.description = dt@entityDescription))
```

and we can also combine this information directly with the `data.frame` describing the attributes of each table:

```{r}
df <-
purrr::map_df(obj@dataset@dataTable, function(dt) 
  data.frame(table.id = dt@id,
             table.description = dt@entityDescription, 
             purrr::map_df(dt@attributeList@attribute, function(x) 
               data.frame(column.name = x@attributeName, column.definition = x@attributeDefinition))))
df
```

This permits us to use typical R manipulations to quickly search this metadata, for instance, to identify which tables contain information about, say, the extreme maximum air temperature over monthly intervals:

```{r}
df %>% filter(grepl("extreme maximum air temperature", column.definition), grepl("monthly", table.description))

```

### 2015-12-29 notebook/2015-12-28-eml-explore.Rmd
a few improvements

@@
 
-```{r}
+```{r message=FALSE}
 library("EML")
@@
 library("dplyr")
+knitr::opts_chunk$set(comment=NA, warning=FALSE)
+
+# Wide tables, since we can scroll left/right
+options("dplyr.width" = Inf, "width" = 1000)
 ```
@@
 
-stuff[[1]]
+stuff[1]
 ```
@@
 ```{r}
-slotNames(stuff[1])
+slotNames(stuff[[1]])
 
@@
 purrr::map_df(stuff, function(x) data.frame(column.name = x@attributeName, definition = x@attributeDefinition))
-
-# Or identically with dplyr alone:
-#dplyr::bind_rows(lapply(stuff, function(x) data.frame(column.name = x@attributeName, definition = x@attributeDefinition)))
-
-do.call(rbind, lapply(stuff, function(x) data.frame(column.name = x@attributeName, definition = x@attributeDefinition)))
-
 ```
 
-This creates a `data.frame` with two columns: the name of every column in Table 4, and a description of what each of those 29 columns contains.
+(Or if you don't like `purrr` and the hadley magic, the same thing with base R:)
+
+```{r eval=FALSE}
+do.call(rbind, lapply(stuff, function(x) data.frame(column.name = x@attributeName, definition = x@attributeDefinition)))
+```
+
+Either way, this creates a `data.frame` with two columns: the name of every column in Table 4, and a description of what each of those 29 columns contains.
 Similarly we can create a `data.frame` that summarizes all of the tables in the dataset:

### 2016-02-04 README.Rmd
add some more package structure

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## eml2

## Developer notes

### Creating eml2 class definitions

Class definitions (`classes.R`) and methods (`methods.R`) are created programmatically.  From the root of the package, run: `source("inst/create-package/create_package.R")`.

### 2016-02-04 tests/testthat/test-party.R

@@ -1,15 +1,17 @@
+testthat::context("eml-party")
 
-library("XML")
-eml_file <- "inst/xsd/test/eml-party.xml"
-root <- XML::xmlRoot(XML::xmlParse(eml_file))
-removeAttributes(root)
-p <- as(root, "party")
-roundtrip <- as(p, "XMLInternalElementNode")
+test_that("we can roundtrip parse and serialze test/eml-party.xml", {
+  library("XML")
+  eml_file <- "inst/xsd/test/eml-party.xml"
+  root <- XML::xmlRoot(XML::xmlParse(eml_file))
+  removeAttributes(root)
+  p <- as(root, "party")
+  roundtrip <- as(p, "XMLInternalElementNode")
 
-## identical except for namespace information in top node
-testthat::expect_identical(
-  saveXML(newXMLNode("party", xmlChildren(root))),
-  saveXML(newXMLNode("party", xmlChildren(roundtrip)))
-  )
-#library("xml2")
-#node <- read_xml(eml_file)
+  ## identical except for namespace information in top node
+  testthat::expect_identical(
+    saveXML(newXMLNode("party", xmlChildren(root))),
+    saveXML(newXMLNode("party", xmlChildren(roundtrip)))
+    )
+
+})

### 2016-02-04 R/classes.R
pass check

@@ -77,6 +77,13 @@ setClass('unit', slots = c('description' = 'character', 'annotation' = 'characte
 #####  inst/xsd/eml-text.xsd  ####
 
 
+setClass('ListOfsection', contains = 'list')
+setClass('ListOfpara', contains = 'list')
+setClass('ListOfvalue', contains = 'list')
+setClass('ListOfvalue', contains = 'list')
+setClass('ListOfvalue', contains = 'list')
+setClass('ListOflistitem', contains = 'list')
+setClass('ListOfvalue', contains = 'list')
 setClass('emphasis', slots = c('value' = 'ListOfvalue'), contains = c('eml-2.1.1'))
 setClass('literalLayout', slots = c('value' = 'ListOfvalue'), contains = c('eml-2.1.1'))
 setClass('ulink', slots = c('citetitle' = 'i18nString', 'url' = 'xml_attribute'), contains = c('eml-2.1.1'))
@@ -107,13 +114,6 @@ setClass('orderedlist', contains = 'ListType')
 setClass('value', contains = 'i18nString')
 setClass('subscript', contains = 'SubSuperScriptType')
 setClass('superscript', contains = 'SubSuperScriptType')
-setClass('ListOfsection', contains = 'list')
-setClass('ListOfpara', contains = 'list')
-setClass('ListOfvalue', contains = 'list')
-setClass('ListOfvalue', contains = 'list')
-setClass('ListOfvalue', contains = 'list')
-setClass('ListOflistitem', contains = 'list')
-setClass('ListOfvalue', contains = 'list')
 
 
 #####  inst/xsd/eml-documentation.xsd  ####
@@ -146,6 +146,14 @@ setClass('angleUnitType', contains = 'character')
 
 
 setClass('RoleType', contains = 'character')
+setClass('ListOfaddress', contains = 'list')
+setClass('ListOfphone', contains = 'list')
+setClass('ListOfelectronicMailAddress', contains = 'list')
+setClass('ListOfonlineUrl', contains = 'list')
+setClass('ListOfuserId', contains = 'list')
+setClass('ListOfsalutation', contains = 'list')
+setClass('ListOfgivenName', contains = 'list')
+setClass('ListOfdeliveryPoint', contains = 'list')
 setClass('phone', slots = c('phonetype' = 'xml_attribute'), contains = c('xs:string', 'eml-2.1.1'))
 setClass('userId', slots = c('directory' = 'xml_attribute'), contains = c('xs:string', 'eml-2.1.1'))
[…]

### 2016-02-04 inst/create-package/xsd_classes.R

@@ -209,6 +209,14 @@ create_classes <- function(xsd_file,
   xml2::xml_find_all(xsd, "//xs:simpleType[@name]", ns = ns) %>%
     purrr::map(set_class_simpletype, file = classes_file)
 
+  ## Create additional ListOf classes for any element that can appear multiple times
+  named_elements <- xml2::xml_find_all(xsd, "//xs:element[@name]", ns = ns)
+  if(length(named_elements) > 0){
+    element <- element_attrs_table(named_elements)
+    element$slot[grepl("ListOf", element$slot)] %>%
+      purrr::map(set_class_list, file = classes_file)
+  }
+
   ## Now get those <xs:element> nodes without a type
   ## For those with one-or-more complexType, define setClass(element-name, slots = <children-of-complex-type)
   untyped_elements <- xml2::xml_find_all(xsd, "//xs:element[@name and not(@type)]", ns = ns)
@@ -229,14 +237,6 @@ create_classes <- function(xsd_file,
   typed_elements %>% purrr::map(set_class_element, file = classes_file)
 
 
-  ## Create additional ListOf classes for any element that can appear multiple times
-  named_elements <- xml2::xml_find_all(xsd, "//xs:element[@name]", ns = ns)
-
-  if(length(named_elements) > 0){
-    element <- element_attrs_table(named_elements)
-    element$slot[grepl("ListOf", element$slot)] %>%
-      purrr::map(set_class_list, file = classes_file)
-  }
 
   ## Define coercions
   xml2::xml_find_all(xsd, "//xs:element[@name] | //xs:simpleType[@name] | //xs:complexType[@name]", ns = ns) %>%

### 2016-02-17 README.Rmd
pass check
fix warnings, update readme

@@
 
-## eml2
+[![Travis-CI Build Status](https://travis-ci.org/cboettig/eml2.svg?branch=master)](https://travis-ci.org/cboettig/eml2)
+
+# eml2
+
+*Note*: `eml2` is work in progress. Please see the Issues tracker in this repository for details about current issues and development milestonds.
+
+## Installation
+
+`eml2` has not yet been released.  Please install from GitHub after installing the `devtools` package (from CRAN):
+
+```{r eval=FALSE}
+devtools::install_github("cboettig/eml2")
+```
+
+## Quickstart
+
+Load the package and read in an EML file:
+
+```{r}
+library("eml2")
+f <- system.file("xsd/test/eml.xml", package = "eml2")
+eml <- read_eml(f)
+```
+
+This creates a native R object called `eml`.  Although this is an "S4 object" type, it uses a "show"
+method which displays in an XML-like layout for convenience:
+
+```{r}
+eml
+```
+
+Validate EML against the official schema
+
+```{r}
+eml_validate(eml)
+```
+
+Write out as EML:
+
+```{r}
+write_eml(eml, "example.xml")
+```
+
+```{r include=FALSE, echo=FALSE}
+unlink("example.xml")
+```
+
+## Manipulating EML objects
+
+Eventually `eml2` will provide constructor and extract methods to create and extract common metadata sections from convenient R stuctures (e.g.
+`data.frames` with unit metadata, common R classes like `Person`, `bibtype`).
+For now, the only method to access and modify EML is to use the standard S4 subsetting and constructor methods.
+
+Our current example does not have a publication date.  Let's add one:
+
+```{r}
+eml@dataset@ResourceGroup@pubDate <- new("pubDate", "2016")
+```
+
+Note that we use the constructor method `new()` to create an object.
+
+### Working with repeating elements
 

### 2016-02-17 R/S4Toeml.R

@@ -32,7 +32,8 @@ S4Toeml <- function(obj,
     if(is.null(node))
       node <- newXMLNode(node_name, namespaceDefinitions = ns)
 
-    base_attributes = c("lang", "schemaLocation")
+    base_attributes = c("lang")
+    schema_attributes = c("schemaLocation")
 
 
 
@@ -49,6 +50,8 @@ S4Toeml <- function(obj,
             attrs <- as.character(slot(obj,s))
             if(s %in% base_attributes)
               s <- paste0("xml:", s)
+            if(s %in% schema_attributes)
+              s <- paste0("xsi:", s)
             names(attrs) <- s
             suppressWarnings(addAttributes(node, .attrs = attrs))
           }

### 2016-02-17 R/classes.R

@@ -1,6 +1,56 @@
+setClass('ListOfvalue')
+setClass('ReferencesGroup')
+setClass('Coverage')
+setClass('OfflineType')
+setClass('OnlineType')
+setClass('UrlType')
+setClass('ConnectionType')
+setClass('horizCoordSysType')
+setClass('CardinalityChildOccurancesType')
+setClass('SingleDateTimeType')
+setClass('MethodsType')
+setClass('ConstraintBaseGroup')
+setClass('ForeignKeyGroup')
+setClass('GRingType')
+setClass('PhysicalOnlineType')
+setClass('Action')
+setClass('NonNumericDomainType')
+setClass('UnitType')
+setClass('PrecisionType')
+setClass('NumericDomainType')
+setClass('EntityGroup')
+setClass('DateTimeDomainType')
+setClass('Accuracy')
+setClass('NumberType')
+setClass('BoundsDateGroup')
+setClass('BoundsGroup')
+setClass('CitationType')
+setClass('PhysicalType')
+setClass('DatasetType')
+setClass('DataQuality')
+setClass('TopologyLevel')
+setClass('ImagingConditionCode')
+setClass('CellGeometryType')
+setClass('ImagingConditionCode')
+setClass('MaintenanceType')
+setClass('MaintUpFreqType')
+setClass('metadata')
+setClass('ConnectionDefinitionType')
+setClass('Article')
+setClass('Chapter')
+setClass('Book')
+setClass('Manuscript')
+setClass('Report')
+setClass('Thesis')
+setClass('ConferenceProceedings')
+setClass('PersonalCommunication')
+setClass('Map')
+setClass('Generic')
+setClass('AudioVisual')
+setClass('Presentation')
 
 setClass('xml_attribute', contains = 'character')
-setClass('eml-2.1.1', slots = c('schemaLocation' = 'xml_attribute', lang = 'xml_attribute'))
+setClass('eml-2.1.1', slots = c('schemaLocation' = 'xml_attribute', 'lang' = 'xml_attribute'))
 setClass('i18nNonEmptyStringType', slots = c('value' = 'ListOfvalue', 'lang' = 'xml_attribute'), contains = c('eml-2.1.1', 'character'))
 setClass('InlineType', contains=c('list'))
 setClass('xs:float', contains = 'numeric')
@@ -37,7 +87,7 @@ setClass('stmml', contains = c('eml-2.1.1', 'character')) ## B
[…]

### 2016-02-17 tests/testthat/test-eml.R

@@ -6,7 +6,6 @@ testthat::test_that("We can parse the sample EML file correctly", {
   f <- system.file("xsd/test", "eml.xml", package = "eml2")
   eml <- read_eml(f)
 
-
   ## FIXME: even basic schema validation needs network connection for w3.org schema checks
   check <- eml_validate(eml)
   testthat::expect_equal(check$status, 0)
@@ -16,32 +15,6 @@ testthat::test_that("We can parse the sample EML file correctly", {
   testthat::expect_equal(check2$status, 0)
 
   unlink("test.xml")
-
-
-  ## Sub-tests
-  node <- xmlRoot(xmlParse(f))
-  removeAttributes(node, .attrs = "xsi:schemaLocation")
-  eml <- as(node, "eml")
-  out <- as(eml, "XMLInternalElementNode")
-
-
-
-  testthat::expect_identical(sort(names(xmlChildren(node[["dataset"]]))), sort(names(xmlChildren(out[["dataset"]]))))
-
-  ## Nodes without repeated elements are identical (as text)
-  sink("test.log")
-  testthat::expect_identical(show(node[["dataset"]][["purpose"]]), show(out[["dataset"]][["purpose"]]))
-  sink()
-  unlink("test.log")
-
-  ## Ordering differs
-  #testthat::expect_identical(show(node[["dataset"]][["distribution"]]), showL(out[["dataset"]][["distribution"]]))
-
-  #out[["dataset"]][["distribution"]]
-  #node[["dataset"]][["distribution"]]
-
-
-  ##testthat::expect_true()
 
 })
 

### 2016-02-26 README.Rmd
rename from eml2 to EML

@@
 
-[![Travis-CI Build Status](https://travis-ci.org/cboettig/eml2.svg?branch=master)](https://travis-ci.org/cboettig/eml2)
+[![Travis-CI Build Status](https://travis-ci.org/cboettig/EML.svg?branch=master)](https://travis-ci.org/cboettig/EML)
 
-# eml2
+# EML
 
-*Note*: `eml2` is work in progress. Please see the Issues tracker in this repository for details about current issues and development milestonds.
+*Note*: `EML` is work in progress. Please see the Issues tracker in this repository for details about current issues and development milestonds.
 
@@
 
-`eml2` has not yet been released.  Please install from GitHub after installing the `devtools` package (from CRAN):
+`EML` has not yet been released.  Please install from GitHub after installing the `devtools` package (from CRAN):
 
 ```{r eval=FALSE}
-devtools::install_github("cboettig/eml2")
+devtools::install_github("cboettig/EML")
 ```
@@
 ```{r}
-library("eml2")
-f <- system.file("xsd/test/eml.xml", package = "eml2")
+library("EML")
+f <- system.file("xsd/test/eml.xml", package = "EML")
 eml <- read_eml(f)
@@
 
-Eventually `eml2` will provide constructor and extract methods to create and extract common metadata sections from convenient R stuctures (e.g.
+Eventually `EML` will provide constructor and extract methods to create and extract common metadata sections from convenient R stuctures (e.g.
 `data.frames` with unit metadata, common R classes like `Person`, `bibtype`).
@@
 
-### Creating eml2 class definitions
+### Creating EML class definitions
 

### 2016-02-26 R/eml.R

@@ -9,7 +9,7 @@
 # @importFrom XML xmlRoot xmlParse
 #' @import methods XML
 #' @examples
-#' f <- system.file("xsd/test", "eml.xml", package = "eml2")
+#' f <- system.file("xsd/test", "eml.xml", package = "EML")
 #' eml <- read_eml(f)
 read_eml <- function(file, ...){
   node <- XML::xmlRoot(XML::xmlParse(file, ...))
@@ -29,7 +29,7 @@ read_eml <- function(file, ...){
 #' @import methods XML
 # @importFrom XML saveXML
 #' @examples
-#' f <- system.file("xsd/test", "eml.xml", package = "eml2")
+#' f <- system.file("xsd/test", "eml.xml", package = "EML")
 #' eml <- read_eml(f)
 #' write_eml(eml)
 write_eml <- function(eml, file = NULL, ...){
@@ -46,7 +46,7 @@ write_eml <- function(eml, file = NULL, ...){
 #' @export
 eml_validate <- function(eml){
 
-  schema <- system.file("xsd/eml.xsd", package = "eml2") #"http://ropensci.github.io/EML/eml.xsd"
+  schema <- system.file("xsd/eml.xsd", package = "EML") #"http://ropensci.github.io/EML/eml.xsd"
 
   if(is(eml, "eml"))
     eml <- write_eml(eml)

### 2016-02-26 inst/examples/hf-205.R

@@ -1,4 +1,4 @@
-library("eml2")
+library("EML")
 
 # Define the metadata shown in "https://raw.githubusercontent.com/ropensci/EML/master/inst/examples/hf205.xml"
 
@@ -37,10 +37,10 @@ keys <-
 description <- "The Geographic region of the kelp bed data extends along the California coast, down through the coast of Baja, Mexico: Central California (Halfmoon Bay to Purisima Point), Southern California (Point Arguello to the United States/Mexico border including the Channel Islands) and Baja California (points south of the United States/Mexico border including several offshore islands)"
 box <- c(-122.44, -117.15, 37.38, 30.00)
 
-eml2:::set_geographicCoverage(box, description)
+EML:::set_geographicCoverage(box, description)
 
 
-eml2:::set_taxonomicCoverage("Sarracenia purpurea")
+EML:::set_taxonomicCoverage("Sarracenia purpurea")
 
 
 

### 2016-02-26 tests/testthat/test-eml.R

@@ -3,7 +3,7 @@ testthat::context("xsd/test/eml.xml")
 testthat::test_that("We can parse the sample EML file correctly", {
   library("XML")
 
-  f <- system.file("xsd/test", "eml.xml", package = "eml2")
+  f <- system.file("xsd/test", "eml.xml", package = "EML")
   eml <- read_eml(f)
 
   ## FIXME: even basic schema validation needs network connection for w3.org schema checks

### 2016-02-29 R/coverage.R
Vignette & helper methods in progress
Vignette illustrates set_EntityGroup and set_coverage now, also avoids dependencies in example setting an attributes table. tweaks to documentation, passes check.

@@ -26,15 +26,15 @@
 #'coverage <- 
 #'  set_coverage(begin = '2012-06-01', end = '2013-12-31',
 #'               sci_names = "Sarracenia purpurea",
-#'               geographicDescription = "the kelp bed data extends along the California coast, down through the coast of Baja, Mexico",
+#'               geographicDescription = "California coast, down through Baja, Mexico",
 #'               west = -122.44, east = -117.15, 
 #'               north = 37.38, south = 30.00)
 #' 
 set_coverage <- function(beginDate = character(), endDate = character(),
                          date = character(), sci_names = character(),
-                         geographicDescription = character(), westBoundingCoordinate = character(), 
-                         eastBoundingCoordinate = character(), northBoundingCoordinate = character(), 
-                         southBoundingCoordinate = character(), altitudeMinimum = numeric(), 
+                         geographicDescription = character(), westBoundingCoordinate = numeric(), 
+                         eastBoundingCoordinate = numeric(), northBoundingCoordinate = numeric(), 
+                         southBoundingCoordinate = numeric(), altitudeMinimum = numeric(), 
                          altitudeMaximum = numeric(), altitudeUnits = character()
                          ){
   new("coverage", 
@@ -52,9 +52,9 @@ set_coverage <- function(beginDate = character(), endDate = character(),
 
 ######## Geographic Coverage ####################
 ## Fixme just rewrite constructor?
-set_geographicCoverage <- function(geographicDescription = character(), westBoundingCoordinate = character(), 
-                                   eastBoundingCoordinate = character(), northBoundingCoordinate = character(), 
-                                   southBoundingCoordinate = character(), altitudeMinimum = numeric(), 
+set_geographicCoverage <- function(geographicDescription = character(), westBoundingCoordinate = numeric(), 
+                                   eastBoundingCoordinate = numeric(), northBoundingCoordinate = numeric(), 
+                                   southBoundingCoordinate = numeric(), altitudeMinimum = numeric(), 
                                    altitudeMaximum = numeric(), altitudeUnits = character()){
   ## Should permit G-Polygon definitions
   new(
@@ -62,10 +62,10 @@ set_geographicCoverage <- function(geographicDescription = character(), westBoun
     geographicDescription = geographicDescription,
     boundingCoordinates = new(
       "boundingCoordinates",
-      westBoundingCoordinate = westBoundingCoordinate,
[…]

### 2016-03-04 inst/create-package/create_package.R
finish up set_section

@@ -1,3 +1,5 @@
+library("EML") ## ick, create_methods actually uses true slots to get slots and contains slots
+
 library("XML")
 library("xml2")
 library("purrr")
@@ -21,7 +23,7 @@ xs_base_classes <- function(file = "R/classes.R", methods_file = "R/methods.R"){
 
   write("
 setClass('xml_attribute', contains = 'character')
-setClass('eml-2.1.1', slots = c('schemaLocation' = 'xml_attribute', 'lang' = 'xml_attribute'))
+setClass('eml-2.1.1', slots = c('schemaLocation' = 'xml_attribute', 'lang' = 'xml_attribute', slot_order = 'character'))
 setClass('i18nNonEmptyStringType', slots = c('value' = 'ListOfvalue', 'lang' = 'xml_attribute'), contains = c('eml-2.1.1', 'character'))
 setClass('InlineType', contains=c('list'))",
         file, append = TRUE)

### 2016-03-09 R/coercions.R
classes using ResponsibleParty get slots directly

@@ -16,34 +16,64 @@ c_as <- function(x, cls){
 
 ## Coercions from "person" type
 setOldClass("person")
-setAs("person", "ResponsibleParty", function(from){
-  new("ResponsibleParty",
-      individualName = c(new("individualName",
-                             "givenName" = as.character(from$given),
-                             "surName" = as(as.character(from$family), "surName"))),
-      electronicMailAddress = as.character(from$email))
-})
 
-setAs("person", "associatedParty", function(from){
-  new("associatedParty",
-      ResponsibleParty =  as(from, "ResponsibleParty"),
-      role = new("role", from$role))
-})
+from_a_person <- function(to){
+  
+  if(to %in% c("associatedParty", "personnel"))
+  
+  function(from){
+  new(to,
+      individualName = new("individualName",
+                           "givenName" = as.character(from$given),
+                           "surName" = as(as.character(from$family), "surName")),
+      electronicMailAddress = as.character(from$email), 
+      role = as.character(from$role))
+  }
+  
+  else
+    
+    function(from){
+      new(to,
+          individualName = new("individualName",
+                               "givenName" = as.character(from$given),
+                               "surName" = as(as.character(from$family), "surName")),
+          electronicMailAddress = as.character(from$email))
+    }
+  
+  
+  }
 
-setAs("person", "creator", function(from){
-  as(as(from, "ResponsibleParty"), "creator")
-})
+from_person <- function(to){
+  function(from){
+  if(length(from) > 1){
+    do.call(c, lapply(from, from_a_person(to)))
+  } else {
+    from_a_person(to)(from)
+  }
+}}
+
+setAs("person", "ResponsibleParty", from_person("ResponsibleParty"))
+setAs("person", "associatedParty", from_person("associatedParty"))
+setAs("person", "personnel", from_person("personnel"))
+setAs("person", "creator", from_person("creator"))
+setAs("person", "contact", from_a_person("contact"))
+setAs("person", "ListOfcreator", from_person("creator"))
 
 ## Coercions to "person" type
 
-setAs("ListOfcreator", "person", function(from){
-  do.call(c, lapply(from, function(x) as(as(x, "ResponsibleParty"), "person")))
+
+
+setAs("creator", "person", function(from){
+  do.call(c, unname(lapply(from@individualName, function(x)
+  person(given = unlist(x@givenName),
+         family = x@surName))))
 })
 
[…]

### 2016-03-09 tests/testthat/test-coercions.R

@@ -1,11 +1,8 @@
-context("coercions")
+testthat::context("coercions")
 
 
 from <- as.person("Carl Boettiger <cboettig@gmail.com>")
-to <- as(from, "ResponsibleParty")
-testthat::expect_is(to, "ResponsibleParty")
-
-to <- as(as(from, "ResponsibleParty"), "contact")
+to <- as(from, "contact")
 testthat::expect_is(to, "contact")
 
 x = as.person("Carl Boettiger <cboettig@gmail.com> [ctb]")

### 2016-03-09 inst/create-package/create_package.R
finish up first draft of vignette

@@ -166,6 +166,7 @@ entity_group_slots <- "'alternateIdentifier' = 'ListOfalternateIdentifier', 'ent
 R <- gsub("'EntityGroup' = 'EntityGroup'", entity_group_slots, R)
 responsible_party_slots = "'individualName' = 'ListOfindividualName', 'organizationName' = 'ListOforganizationName', 'positionName' = 'ListOfpositionName', 'address' = 'ListOfaddress', 'phone' = 'ListOfphone', 'electronicMailAddress' = 'ListOfelectronicMailAddress', 'onlineUrl' = 'ListOfonlineUrl', 'userId' = 'ListOfuserId', 'references' = 'references', 'id' = 'xml_attribute', 'system' = 'xml_attribute', 'scope' = 'xml_attribute'"
 R <- gsub("'ResponsibleParty' = 'ResponsibleParty'", responsible_party_slots, R)
+R <- gsub("'ProcedureStepType' = 'ProcedureStepType'", "'description' = 'TextType', 'citation' = 'ListOfcitation', 'protocol' = 'ListOfprotocol', 'instrumentation' = 'ListOfinstrumentation', 'software' = 'ListOfsoftware', 'subStep' = 'ListOfsubStep'", R)
 
 R <- unique(R)
 write(R, classes_file)

### 2016-03-18 inst/examples/metadata-repositories.Rmd
Draft vignette testing out metadata repositories.

---
title: "Metadata Repositories"
author: "Carl Boettiger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

One of the principle advantages in creating EML is to make it easier to find the data you need, without having to standardize all your data files themselves into one giant database.  Instead, the data files can be whatever you want, provided the relevant information you might search on to discover data of interest is listed in the metadata.

To do this, we will use the `dataone` package to upload a private copy of our data file to the central data repository.

```{r eval = FALSE}
install.packages("dataone", repos= c("https://cran.rstudio.com", "http://nceas.github.io/drat"))
```

```{r}
library("dataone")
library("datapackage")
```

Imagine we have the file paths to our `.csv` data file and an `.xml` EML file providing metadata for it:

```{r}
sampleData <- system.file("extdata/sample.csv", package="dataone")
sampleEML <- system.file("extdata/sample-eml.xml", package="dataone")
```

To upload these to a metadata repository such as [the KNB](http://knb.ecoinformatics.org), we simply create `DataObject`s for both files:

```{r}
dataObj <- new("DataObject", format="text/csv", file=sampleData)
metadataObj <- new("DataObject", format="eml://ecoinformatics.org/eml-2.1.1", file=sampleEML)
```

Note that optionally, `new("DataObject")` could have been given an `id` argument, which could be a (namespaced) UUID from `UUIDgenerate`, or a DOI from a member node (see `generateIdentifier()`).  Since no `id` has been given, a UUID is automatically generated for each.

```{r}
accessRules <- data.frame(subject="CN=Noam Ross A45991,O=Google,C=US,DC=cilogon,DC=org", permission="write")
dataObj <- addAccessRule(dataObj, accessRules)
metadataObj <- addAccessRule(metadataObj, accessRules)
```

We now want to bundle these two objects (data and metadata) into a single "data package" to be uploaded.  To do so, we just create a new `DataPackage` object and then add the data and metadata using the `addData` file:

```{r}
dp <- new("DataPackage")
addData(dp, dataObj, metadataObj)
```

This both adds the files and registers that the metadata object describes the data object.

```{r}
d1c <- D1Client("STAGING", "urn:node:mnStageUCSB2")
packageId <- uploadDataPackage(d1c, dp)
```

Let's see if the ID returned for our package now appears in the DataONE index:

```{r}
query(CNode("STAGING"), searchTerms = list(id = packageId))
```

---------------

Gimme DOI! Note that the example would fail if run here since only the Production (`PROD`) environment can provide DOIs (the `STAGING` environment is only for tests and training examples), and only then on member nodes that offer DOIs, such as the KNB (`urn:node:KNB`).

```{r eval=FLASE}
cn <- CNode("STAGING")
mn <- getMNode(cn, "urn:node:mnStageUCSB2")
newid <- generateIdentifier(mn, "DOI")

## now use the DOI in packaging the DataObject:
new("DataObject", id = newid, format = "eml://ecoinformatics.org/eml-2.1.1", file=sampleEML)
```

### 2016-03-23 R/classes-stmml.R
re-organize functions, implement eml_get as in #153

@@ -1,3 +1,5 @@
+#' @include classes.R
+
 
 setClass('ListOfunitType', contains = 'list')
 setClass('ListOfunit', contains = 'list')

### 2016-03-23 R/eml_find.R

@@ -1,0 +1,50 @@
+#eml_get <- function(eml, element, ...){
+#  eml_get(eml_find(eml, element), ...)
+#  
+#}
+
+
+
+#' eml_find
+#' 
+#' find all occurrances of an EML element in at a given node at any depth
+#' @param node an EML S4 object; e.g. `eml` or any of it's child/descendant slots.
+#' @param element the name of an EML element to find
+#' @return the requested element as an EML S4 object
+#' @export
+#' 
+eml_find <- function(node, element){
+  
+  if(is.list(node) && length(node) > 0){
+    nonempty_lapply(node, eml_find, element)
+  } else {
+    slots <- getSlots(class(node))
+    children <- names(slots)
+    children <- children[!(slots %in% c("character", "xml_attribute", "XMLInternalNode"))]
+ 
+    if(element %in% children){
+      slot(node, element)
+  
+    } else { 
+      nonempty_lapply(children, function(x) eml_find(slot(node, x), element))
+    }
+  }
+}
+
+
+nonempty_lapply <- function(x, f, ...){
+  found <- list()
+  for(i in x){
+    out <- f(i, ...)
+    if(isS4(out) || length(out) > 0) 
+      found <- c(found, list(out))
+  }
+  if(length(found) == 1)
+    found[[1]]
+  else if(length(found) == 0)
+    NULL
+  else 
+    found
+} 
+
+

### 2016-03-23 R/read_eml.R

@@ -1,0 +1,25 @@
+
+#' read_eml
+#'
+#' read_eml
+#' @param file path to an EML file
+#' @param ... additional arguments to \code{\link{xmlParse}}
+#' @return an eml object (S4 object)
+#' @export
+# @importFrom XML xmlRoot xmlParse
+#' @import methods XML
+#' @examples
+#' f <- system.file("xsd/test", "eml.xml", package = "EML")
+#' eml <- read_eml(f)
+read_eml <- function(file, ...){
+  node <- XML::xmlRoot(XML::xmlParse(file, ...))
+  emlToS4(node)
+}
+
+
+
+## Creates a 'show' method so that eml S4 elements display in XML format instead of the
+## impossible-to-read S4 format
+setMethod("show", signature("eml-2.1.1"), function(object) show(S4Toeml(object)))
+
+# cat(yaml::as.yaml(XML::xmlToList(S4Toeml(object))))

### 2016-05-06 inst/examples/metadata-repositories.Rmd
update to reflect changes in dataone & datapack

@@
 ```{r eval = FALSE}
-install.packages("dataone", repos= c("https://cran.rstudio.com", "http://nceas.github.io/drat"))
+#install.packages("dataone", repos= c("https://cran.rstudio.com", "http://nceas.github.io/drat"))
+
+## or
+devtools::install_github(c("ropensci/datapack", "DataONEorg/rdataone"))
 ```
@@
 library("dataone")
-library("datapackage")
+library("datapack")
 ```
@@
 dp <- new("DataPackage")
-addData(dp, dataObj, metadataObj)
+dp <- addData(dp, dataObj, metadataObj)
 ```

### 2016-06-17 R/set_section.R
Address #158 by introducing set_TextType()
set_TextType() replaces set_section as a more general way to handle creating TextType objects from external files, both as para and section formats.

@@ -1,20 +1,11 @@
 
-#' set_section
+#' set_TextType
 #' 
-#' For any EML element that takes a "section" argument, this function can be used to generate the appropriate EML from a markdown-formatted file.
+#' For any EML element of class TextType, this function can be used to generate the appropriate EML from a markdown-formatted file.
+#' @param text a plain text character string which will be used directly as the content of the node if no file is given
 #' @param file path to a file providing formatted input text, see details.
 #' @import XML
-#' @return a ListOfsection object that can be used in constructing any element with section slots, see examples
-#' @examples 
-#' f <- system.file("examples/hf205-methods.md", package = "EML")
-#' set_section(f)
-#' new("description", section = set_section(f))
-#' 
-#' ## Can also import from methods written in a .docx MS Word file. 
-#' f <- system.file("examples/hf205-methods.docx", package = "EML")
-#' set_section(f)
-#' new("description", section = set_section(f))
-#' 
+#' @return a TextType object that can be coerced into any element inheriting from TextType, see examples
 #' @importFrom tools file_ext
 #' @details If the `rmarkdown` package is installed, then the input file can 
 #' be a Microsoft Word (.docx) file, a markdown file, or other file
@@ -23,15 +14,65 @@
 #' .xml or .dbk extension).  Note that pandoc comes pre-installed in RStudio and is 
 #' required for the rmarkdown package.  
 #' @export
-set_section <- function(file = NULL){
-  
-  if(!tools::file_ext(file) %in% c("xml", "dbk")){
+#' @examples
+#' \dontrun{
+#' ## using a simple character string
+#' a <- set_TextType(text = "This is the abstract")
+#' as(a, "abstract")
+#'
+#' ## Using an external markdown file
+#' f <- system.file("examples/hf205-abstract.md", package = "EML")
+#' a <- set_TextType(f)
+#' as(a, "abstract")
+#' 
+#' ## Can also import from methods written in a .docx MS Word file. 
+#' f <- system.file("examples/hf205-abstract.docx", package = "EML")
+#' a <- set_TextType(f)
+#' as(a, "abstract")
+#' 
+#' ## Documents with title headings use `section` instead of `para` notation 
[…]

### 2016-06-17 tests/testthat/test-TextType.R

@@ -1,0 +1,19 @@
+## using a simple character string
+a <- set_TextType(text = "This is the abstract")
+as(a, "abstract")
+
+## Using an external markdown file
+f <- system.file("examples/hf205-abstract.md", package = "EML")
+a <- set_TextType(f)
+as(a, "abstract")
+
+## Can also import from methods written in a .docx MS Word file.
+f <- system.file("examples/hf205-abstract.docx", package = "EML")
+a <- set_TextType(f)
+as(a, "abstract")
+
+## Documents can have richer formatting.  Documents with title headings use `section` instead of `para` notation
+f <- system.file("examples/hf205-methods.docx", package = "EML")
+d <- set_TextType(f)
+as(d, "description")
+ 

### 2016-06-17 README.Rmd
minor readme updates

@@
 
-[![Travis-CI Build Status](https://travis-ci.org/cboettig/EML.svg?branch=master)](https://travis-ci.org/cboettig/EML)
+[![Travis-CI Build Status](https://travis-ci.org/ropensci/EML.svg?branch=master)](https://travis-ci.org/ropensci/EML)
 
@@
 
-`EML` has not yet been released.  Please install from GitHub after installing the `devtools` package (from CRAN):
+`EML` has not yet been released.  You can install the current development version from rOpenSci using:
+
+```{r eval=FALSE}
+install.packages("EML", repos = c("http://packages.ropensci.org", "https://cran.rstudio.com"))
+```
+
+or install directly from GitHub (after installing the `devtools` package from CRAN):
 
@@
 
-```{r}
+```{r, message=TRUE}
+# An EML document with no validation errors
 eml_validate(eml)
+
+# An EML document with validation errors
+invalid_eml <- system.file("tests/testthat/", "example-eml-invalid.xml", package = "EML")
+
+eml_validate(invalid_eml)
 ```
@@
 ```{r}
-eml@dataset@ResourceGroup@pubDate <- new("pubDate", "2016")
+eml@dataset@pubDate <- new("pubDate", "2016")
 ```

### 2016-06-21 README.Rmd
Add link to EML schema, more background in README
closes #164

@@
 
-# EML
+# EML: The Ecological Metadata Language Standard
 
-*Note*: `EML` is work in progress. Please see the Issues tracker in this repository for details about current issues and development milestonds.
+EML is a widely used metadata standard in the ecological and environmental sciences.
+We strongly recommend that interested users visit the [EML Homepage](https://knb.ecoinformatics.org/#external//emlparser/docs/index.html) for an introduction and thorough documentation of the standard.
+Additionally, the scientific article []() provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly hetergenous data that cannot be easily reduced to a traditional vertically integrated database.
+At this time, the `EML` R package provides support for the serializing and parsing of all low-level EML concepts, but still assumes some familiarity with the EML standard, particularly for users seeking to create their own EML files.
+We hope to add more higher-level functions which will make such familiarity less essential in future development.
+
+*Note*: The `EML` R package is work in progress. Please see the Issues tracker in this repository for details about current issues and development milestonds.
 
@@
 
-## Manipulating EML objects
+## Reading EML files
 
-Eventually `EML` will provide constructor and extract methods to create and extract common metadata sections from convenient R stuctures (e.g.
-`data.frames` with unit metadata, common R classes like `Person`, `bibtype`).
-For now, the only method to access and modify EML is to use the standard S4 subsetting and constructor methods.
+Please see the vignette, [Parsing EML](https://github.com/ropensci/EML/blob/master/vignettes/parsing-EML.Rmd) for a more thorough introduction to reading EML files and extracting relevant metadata.
 
-Our current example does not have a publication date.  Let's add one:
+## Creating EML files
 
-```{r}
-eml@dataset@pubDate <- new("pubDate", "2016")
-```
+Please see the vignette, [Creating EML](https://github.com/ropensci/EML/blob/master/vignettes/creating-EML.Rmd) for a proper introduction to creating your own EML files.
 
-Note that we use the constructor method `new()` to create an object.
+## Working with custom units
 
-### Working with repeating elements
+Users may occassionally need to define custom unit types that EML does not know about.
+Please see the vignette, [Working with units](https://github.com/ropensci/EML/blob/master/vignettes/working-with-units.Rmd) for an overview of how units work in EML, how to view the standard unit library, and how to access and define custom units and custom unit types.
+
+## Archiving, Sharing, and Publishing EML
+
+One of the chief advantages of using EML to manage your own data is the improved ability to search all of your data files, e.g.
+for the ones that have the particular measurements you need on the particular location or species of interest, and be able to share this data either publicly or securely with select collaborators.
+To do so, you will need to upload your EML file and data to a scientific metadata repository that understands EML.
+The vignette, [metadata repositories](https://github.com/ropensci/EML/blob/master/vignettes/metadata-repositories.Rmd) describes how to do this.
 

### 2016-06-23 README.Rmd

@@
 ```{r eval=FALSE}
-devtools::install_github("cboettig/EML")
+devtools::install_github("ropensci/EML")
 ```
@@
 # An EML document with validation errors
-invalid_eml <- system.file("tests/testthat/", "example-eml-invalid.xml", package = "EML")
+invalid_eml <- system.file("xsd/test/example-eml-invalid.xml", package = "EML")
 
@@
 
-```{r}
+```{r, eval=FALSE}
 write_eml(eml, "example.xml")
@@
 Class definitions (`classes.R`) and methods (`methods.R`) are created programmatically.  From the root of the package, run: `source("inst/create-package/create_package.R")`.
+
+### Meta
+Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2016-06-24 README.Rmd

@@
 [![Travis-CI Build Status](https://travis-ci.org/ropensci/EML.svg?branch=master)](https://travis-ci.org/ropensci/EML)
+[![codecov.io](https://codecov.io/github/ropensci/EML/coverage.svg?branch=master)](https://codecov.io/github/ropensci/EML?branch=master)
 

### 2016-06-24 README.Rmd
fix missing link, closes #171

--- 2016-06-23 README.Rmd
@@
 We strongly recommend that interested users visit the [EML Homepage](https://knb.ecoinformatics.org/#external//emlparser/docs/index.html) for an introduction and thorough documentation of the standard.
-Additionally, the scientific article []() provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly hetergenous data that cannot be easily reduced to a traditional vertically integrated database.
+Additionally, the scientific article _[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](http://doi.org/10.1146/annurev.ecolsys.37.091305.110031)_ provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly hetergenous data that cannot be easily reduced to a traditional vertically integrated database.
 At this time, the `EML` R package provides support for the serializing and parsing of all low-level EML concepts, but still assumes some familiarity with the EML standard, particularly for users seeking to create their own EML files.
@@
 ### Meta
+
 Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2016-10-13 README.Rmd
add appveyor #175

--- 2016-06-24 README.Rmd
@@
 [![Travis-CI Build Status](https://travis-ci.org/ropensci/EML.svg?branch=master)](https://travis-ci.org/ropensci/EML)
+[![codecov.io](https://codecov.io/github/ropensci/EML/coverage.svg?branch=master)](https://codecov.io/github/ropensci/EML?branch=master)
+[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/EML?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/EML)
 

### 2017-03-02 README.Rmd
update README
include gh-pages version of dcos from pkgdown
- Also modifies Description with URL for EML
- and updates Readme not to say no CRAN version is available.

--- 2016-06-24 README.Rmd
@@
 [![Travis-CI Build Status](https://travis-ci.org/ropensci/EML.svg?branch=master)](https://travis-ci.org/ropensci/EML)
+[![codecov.io](https://codecov.io/github/ropensci/EML/coverage.svg?branch=master)](https://codecov.io/github/ropensci/EML?branch=master)
+[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/EML?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/EML)
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/EML)](http://cran.r-project.org/package=EML)
+[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/EML)](http://cran.r-project.org/web/packages/EML/index.html)
+[![Github Stars](https://img.shields.io/github/stars/ropensci/EML.svg?style=social&label=Github)](https://github.com/ropensci/EML)
+[![DOI](https://zenodo.org/badge/10894022.svg)](https://zenodo.org/badge/latestdoi/10894022)
 
@@
 
-`EML` has not yet been released.  You can install the current development version from rOpenSci using:
+You can install the current development version from rOpenSci using:
 
@@
 
-Validate EML against the official schema
+Validation helps ensure that an EML file can be accurately interpreted by any computer program designed to read EML.  We can validate any EML against the official schema:
 

### 2017-04-20 README.Rmd
use canonical form link to CRAN in README

--- 2016-06-24 README.Rmd
@@
 [![codecov.io](https://codecov.io/github/ropensci/EML/coverage.svg?branch=master)](https://codecov.io/github/ropensci/EML?branch=master)
+[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/EML?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/EML)
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/EML)](http://cran.r-project.org/package=EML)
+[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/EML)](https://CRAN.R-project.org/package=EML)
+[![Github Stars](https://img.shields.io/github/stars/ropensci/EML.svg?style=social&label=Github)](https://github.com/ropensci/EML)
+[![DOI](https://zenodo.org/badge/10894022.svg)](https://zenodo.org/badge/latestdoi/10894022)
 

### 2017-04-25 README.Rmd
canonical cran link should be https

--- 2016-10-13 README.Rmd
@@
 [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/EML?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/EML)
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/EML)](https://cran.r-project.org/package=EML)
+[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/EML)](https://CRAN.R-project.org/package=EML)
+[![Github Stars](https://img.shields.io/github/stars/ropensci/EML.svg?style=social&label=Github)](https://github.com/ropensci/EML)
+[![DOI](https://zenodo.org/badge/10894022.svg)](https://zenodo.org/badge/latestdoi/10894022)
 

### 2017-07-21 README.Rmd

--- 2017-04-20 README.Rmd
@@
 [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/EML?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/EML)
-[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/EML)](http://cran.r-project.org/package=EML)
+[![](https://ropensci.org/badges/80_status.svg)](https://github.com/ropensci/onboarding/issues/80)
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/EML)](https://cran.r-project.org/package=EML)
 [![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/EML)](https://CRAN.R-project.org/package=EML)
+[![DOI](https://zenodo.org/badge/10894022.svg)](https://zenodo.org/badge/latestdoi/10894022)
+
+<!--
 [![Github Stars](https://img.shields.io/github/stars/ropensci/EML.svg?style=social&label=Github)](https://github.com/ropensci/EML)
-[![DOI](https://zenodo.org/badge/10894022.svg)](https://zenodo.org/badge/latestdoi/10894022)
+-->
 

### 2017-07-21 README.Rmd
add onboarding badge

--- 2017-04-20 README.Rmd
@@
 [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/EML?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/EML)
-[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/EML)](http://cran.r-project.org/package=EML)
+[![Onboarding](https://ropensci.org/badges/80_status.svg)](https://github.com/ropensci/onboarding/issues/80)
+[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/EML)](https://cran.r-project.org/package=EML)
 [![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/EML)](https://CRAN.R-project.org/package=EML)
-[![Github Stars](https://img.shields.io/github/stars/ropensci/EML.svg?style=social&label=Github)](https://github.com/ropensci/EML)
 [![DOI](https://zenodo.org/badge/10894022.svg)](https://zenodo.org/badge/latestdoi/10894022)
+
+<!-- [![Github Stars](https://img.shields.io/github/stars/ropensci/EML.svg?style=social&label=Github)](https://github.com/ropensci/EML)-->
 

### 2017-08-15 inst/onboarding/review-reply.md
onboarding review reply draft

We believe we have now addressed the very helpful issues and suggestions raised by both reviewers.  While EML is still under active development that may address some of the more big-picture issues about being more user friendly, we feel this addresses the essential issues highlighted by reviewers as listed below.  Apologies for the long delay.

# cgries

---

> For the average user (me included) it would be nice to have a little more extensive documentation. I.e., in the description field of functions it would help to see more than just the function name again.  The examples are all very helpful and work great. The vignettes are great as well and really help to get started writing code to generate EML. The explanation for the ‘new’ method is, however, very short. I realize that it mainly takes understanding XML and EML specifically to get the hang of that and EML is complex with many ‘new’ constructs for using the more obscure elements.

We agree.  We're working on making easier constructors, as suggested by @gmbecker, basically this is just waiting on his s4Coffee implementation which should be able to automate the generation of the 100s of methods needed to do this.  With help from the user community, we're also flushing out documentation aimed at users less familiar with EML and/or R internals, though of course it was necessary to get the fundamentals in place first, which is what we feel the current release now does.

> One improvement that could still be considered for the first release is dealing with people information. I found myself very quickly writing some helper functions which I then also discovered in the arctic repository that was mentioned by @amoeba in [https://github.com/ropensci/EML/issues/183](url). Given the fact that we both seemed to need such and the basic `as.person` is not sufficient, some thought might go into improving that before everyone starts writing those helper functions. I would like to see the option to provide a csv file/data frame with the people information following the pattern of attribute and taxonomic coverage information. I have had datasets with almost 60 creators and their sequence changed several times in the process of publishing the paper. That would be very tedious to work out in code.
Which leads me to responding to @gmbecker that I find it very convenient to provide attribute and taxonomic information as a data frame. It allows for easy retrieval of the information from either a database or more frequently from a spreadsheet that data providers supply with this information. Most data providers don’t know about the EML subtleties of measurementScale or how to correctly spell the units, but it takes very little editing on my part to prepare a spreadsheet from the information that I do receive. Taxonomic coverage can be extensive and again frequently comes out of a database, i.e., can be easily provided as csv file and data frame.

We agree that being able to maintain metadata such as taxonomic information or author details in in simple .csv files that can quickly be read into EML is a big win, and have been building methods around these patterns.  Understanding the use cases such as @cgries outlines here helps us do that better.  This is all currently possible with the `new` constructors, but requires a bit more coding.  Once we've auto

----

# gmbecker

## Package Review

### Constructors

> Generally, all classes which any end users are expected to create themselves should have exported constructors, rather than requiring users to call `new` directly. This preserves the abstraction if the underlying class structure changes, and usually increases the readability of the users' code. ....

We agree, this will be done once we can automate the creation of these methods with your awesome S4Coffee package!  See https://github.com/gmbecker/S4Coffee/issues/1.

### show method

> The current show method for `eml` objects converts the object to XML before printing. This seems cumbersome, and, I suspect, is likely to cause problems for real-world-sized EML documents (in the same way the XML package does when printing a large XML blob).  It seems that EML is structured enough that a meaningful summary could be displayed by `show`, while  the emlToXML function would still allow users to easily display the entire XML blob if desired.

Yup, display issues now addressed in #182 and #189.

### duplicated classes

> I understand that the classes are automatically generated, but there seem to be some minor issues with how this is done. My copy of the package has two `value` class definitions, one right after the other, and they are different to boot. The first one, of course, will do nothing.

> Also, there are the, e.g., `author` `Author` style pairs.  These are troublesome as well, from a design perspective, because the classes are the same, other than a different ordering of the class slots.  If the actual order of the class slots is relied upon in the code, I strongly  encourage refactoring away from that. Slots should, in my opinion, be checked for and used exclusively by name or accessor method.

This is a consequence of the underlying XSD schema.  `author` and `Author` are different elements in EML, so they are different here.  (Any class beginning with a captial is a generic XSD type, while those starting with a lower-case are actual valid complex elements in the XML.)  S4 classes are also simpler/more restricted than XSD classes, so the mapping is imperfect (XSD permits repeated elements, ordered or unordered child elements, etc).  An ideal implementation would probably avoid S4 entirely and focus on a more 1:1 correspondence to the XSD classes, but this would require creating a unique class system which is clearly beyond the scope here.  Note that regarding validation, this already happens on the XSD schema validation level, so is not re-implemented in S4 validation.  Some higher-level constructor functions do now have validation support though.

## Other

### eml_get

The eml_get function's return value seems unintuitive to me. When run on the example hf205 EML content for "physical" elements, it returns a list, of length 3, with 3 length-1 ListOfPhysical objects. I don't follow why this is not a single length 3 ListOfPhysical object.

### class(blah) == "thing"

> This is not a safe way to test the class of an object, as it will miss/fail for both S3 (vector of classes) and S4 (formal) inheritance. E.g., a tibble will fail a `class(obj) == "data.frame"` check, as would an S4 class that contains data.frame. All such checks should be refactored to use `is`

fixed

### set_coverage data.frame column order requirement

> This seems like something that can be checked and enforced programatically rather than relying on good user behavior. The possible columns are fixed and the order is known.

> Another approach here is to create an S4 object with constructor that formally models the taxonomic hierarchy, rather than relying on a powerful but semantically poor structure like a data.frame. Particularly since from what I can tell it will only ever have one row, making the structure not a great fit anyway.

> It's not clear that data.frames are the correct model for _constructing_ the components of eml. It seems that an `eml_attribute` constructor, which can be called repeatedly and constructs a single attribute (row in the data.frame as it is currently factored) is more useful friendly With respect, it seems like it would be at least as easy to write a csv file directly than constructing and rbinding together these data.frames. I feel more/deeper api-design work is called for on the EML construction side of things.

> As @cgries points out, there are often many rows (for data working with many different species), which make this format convenient.

### auto-generated code

Files with autogenerated code should contain a header comment to that effect and telling possible future contributors not to edit the file manually. This should be emitted automatically by the code-generation apparatus.

### importing packages

> The entire XML packages is imported. I actually feel this is correct in this case (and not terrible generally imho) but it is technically against the ropensci package guidelines. I leave it up to @noamross as the editor to make the final call on this. I recommend it not be an issue for final acceptance.

Dependency is now on `xml2`.

### Automated submission to EML repository?

> This is somewhat conjecture on my part, given the caveat I started these review comments with, but is there a single (or small set of), major repository for EML? If so, does it have an API for submission? If so, bindings to that so that users can submit their newly created EML from within R would magnify the utility of the creation aspect of this package substantially, and should be added either to this package (or via the creating of a sister convenience package).

Indeed!  This is already implemented in the `dataone` R package and docuemnted there.  We have also added a vignette to the EML package documenting it on our end!

### have the S4 version of the object carry around it's source XML

> XML objects are pointers, so duplication/memory will not be an issue. This could get hairy/infeasible if users are modifying the S4 representation of the eml manually and then re-exporting, but I don't get the sense that this is an intended use-pattern. This would make a couple of parts of the code much cleaner/more efficient when they switch between representations.

Users can create S4 objects from scratch that don't have a source XML.  In general things have been much more stable if we avoid too much manipulation through pointers to the XML and work primarily with in-memory, native S4 R objects.

### Does everything need to inherit from eml-2.1.1?

> I can see some benefits to this, but it also complicates things. A lot of the low-level/internal classes would inherit directly from character if not for also implimenting eml-2.1.1. Are those extra slots ever used at the low level, or only, e.g., once per larger document?

Yes, this lets us distinguish between some required behavior of all the EML objects when creating the serialization methods, etc.

# Further reviewer comment

> @cboettig has started a discussion regarding the time-scale of implementing improvements vs putting out a stable release here ropensci/EML#183. In my capacity as a reviewer I feel it is most appropriate that I place my thoughts on this subject here as a secondary part of my review. Package authors, please find my recommendations on this question below.

> The most important aspect of a stable (read: CRAN) release is that the API is stable upon release for at least the near and preferably the medium term. As such, the most important issues to fix before release are ones that would actually break API compatibility. These are
> - changing the output class of eml_get, and
> - changing the names of existing constructors to remove the set_ prefix which I argue is misleading/a misnomer

These are excellent points, and actually the reason we ended up doing a CRAN release: the package already had signficant use in some important pipelines for the NSF Arctic database, and we felt having an archived, stable version for this was important.  Future modifications will provide a more friendly higher-level interface without breaking code that is built on lower-level constructors.

> Beyond that, I feel the class checking is the next most important. People in the CRAN sphere are almost sure (in my opinion) to want to pass tbl_df objects to things that take data.frames. Tests to ensure that this works would not be out-of-place, and they should pass before CRAN release.

Yup, we've added these tests on `tbl_df`.

> The other changes, while more important from a package design improvement perspective, can probably wait if they must, because they will likely be either backwards compatible or large enough that they warrant a new major version of the software.

Yup, that's the plan.

### 2018-09-14 README.Rmd
Import eml2

--- 2017-04-25 README.Rmd
@@
   md_document:
-    variant: markdown_github
+    variant: gfm
 ---
+
+[![Build Status](https://travis-ci.org/cboettig/eml2.svg?branch=master)](https://travis-ci.org/cboettig/eml2)
+[![Coverage Status](https://img.shields.io/codecov/c/github/cboettig/eml2/master.svg)](https://codecov.io/github/cboettig/eml2?branch=master)
+[![stability-experimental](https://img.shields.io/badge/stability-experimental-orange.svg)](https://github.com/joethorley/stability-badges#experimental)
 
@@
 
-[![Travis-CI Build Status](https://travis-ci.org/ropensci/EML.svg?branch=master)](https://travis-ci.org/ropensci/EML)
-[![codecov.io](https://codecov.io/github/ropensci/EML/coverage.svg?branch=master)](https://codecov.io/github/ropensci/EML?branch=master)
-[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/EML?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/EML)
-[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/EML)](https://cran.r-project.org/package=EML)
-[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/EML)](https://CRAN.R-project.org/package=EML)
-[![Github Stars](https://img.shields.io/github/stars/ropensci/EML.svg?style=social&label=Github)](https://github.com/ropensci/EML)
-[![DOI](https://zenodo.org/badge/10894022.svg)](https://zenodo.org/badge/latestdoi/10894022)
-
-# EML: The Ecological Metadata Language Standard
-
-EML is a widely used metadata standard in the ecological and environmental sciences.
-We strongly recommend that interested users visit the [EML Homepage](https://knb.ecoinformatics.org/#external//emlparser/docs/index.html) for an introduction and thorough documentation of the standard.
-Additionally, the scientific article _[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](http://doi.org/10.1146/annurev.ecolsys.37.091305.110031)_ provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly hetergenous data that cannot be easily reduced to a traditional vertically integrated database.
-At this time, the `EML` R package provides support for the serializing and parsing of all low-level EML concepts, but still assumes some familiarity with the EML standard, particularly for users seeking to create their own EML files.
-We hope to add more higher-level functions which will make such familiarity less essential in future development.
-
-*Note*: The `EML` R package is work in progress. Please see the Issues tracker in this repository for details about current issues and development milestonds.
-
-## Installation
-
-You can install the current development version from rOpenSci using:
-
-```{r eval=FALSE}
-install.packages("EML", repos = c("http://packages.ropensci.org", "https://cran.rstudio.com"))
+```{r include=FALSE}
+has_pandoc <-  rmarkdown::pandoc_available()
 ```
 
-or install directly from GitHub (after installing the `devtools` package from CRAN):
+# eml2
 
-```{r eval=FALSE}
-devtools::install_github("ropensci/EML")
+The goal of `eml2` is to provide both a drop-in replacement for the higher-level functions of the existing EML package while also providing additional functionality.
+
+`eml2` uses only simple and familiar list structures (S3 classes) instead of the more cumbersome use of S4 found in the original EML.
+While the higher-level functions are identical, this makes it easier to for most users and developers to work with `eml` objects and also to write their own functions for creating and manipulating EML objects.
+Under the hood, `eml2` relies on the [emld](https://github.com/cboettig/emld) package, which uses a Linked Data representation for EML.
+It is this approach which lets us combine the simplicity of lists with the specificy required by the XML schema.
+
+# Creating EML
+
+```{r message=FALSE, warning=FALSE}
+library(eml2)
 ```
 
-## Quickstart
-
-Load the package and read in an EML file:
+## A minimal valid EML document:
 
 ```{r}
-library("EML")
-f <- system.file("xsd/test/eml.xml", package = "EML")
-eml <- read_eml(f)
+me <- list(individualName = list(givenName = "Carl", surName = "Boettiger"), id="cboettig")
+my_eml <- list(dataset = list(
+              title = "A Mimimal Valid EML Dataset",
+              creator = me,
+              contact = me)
+            )
+
+write_eml(my_eml, "ex.xml")
+eml_validate("ex.xml")
 ```
 
-This creates a native R object called `eml`.  Although this is an "S4 object" type, it uses a "show"
-method which displays in an XML-like layout for convenience:
+## A Richer Example
+
+Here we show a the creation of a relatively complete EML document using `eml2`.
+This closely parallels the function calls shown in the original EML [R-package vignette](https://ropensci.github.io/EML/articles/creating-EML.html).
+
+## `set_*` methods
+
+The original EML R package defines a set of higher-level `set_*` methods to facilitate the creation of complex metadata structures.
+`eml2` provides these same methods, taking the same arguments for `set_coverage`, `set_attributes`, `set_physical`, `set_methods` and `set_textType`, as illustrated here:
+
+### Coverage metadata
 
 ```{r}
-eml
+geographicDescription <- "Harvard Forest Greenhouse, Tom Swamp Tract (Harvard Forest)"
+coverage <- 
+  set_coverage(begin = '2012-06-01', end = '2013-12-31',
+               sci_names = "Sarracenia purpurea",
+               geographicDescription = geographicDescription,
+               west = -122.44, east = -117.15, 
+               north = 37.38, south = 30.00,
+               altitudeMin = 160, altitudeMaximum = 330,
+               altitudeUnits = "meter")
 ```
 
-Validation helps ensure that an EML file can be accurately interpreted by any computer program designed to read EML.  We can validate any EML against the official schema:
+### Reading in text from Word and Markdown
 
-```{r, message=TRUE}
-# An EML document with no validation errors
-eml_validate(eml)
+We read in detailed methods written in a Word doc.
+This uses EML's docbook-style markup to preserve formatting of paragraphs, lists, titles, and so forth.
+(This is a drop-in replacement for EML `set_method()`)
 
-# An EML document with validation errors
-invalid_eml <- system.file("xsd/test/example-eml-invalid.xml", package = "EML")
-
-eml_validate(invalid_eml)
+```{r eval=has_pandoc}
+methods_file <- system.file("examples/hf205-methods.docx", package = "EML")
+methods <- set_methods(methods_file)
 ```
 
-Write out as EML:
+We can also read in text that uses Markdown for markup elements:
 
-```{r, eval=FALSE}
-write_eml(eml, "example.xml")
+```{r eval=has_pandoc}
+abstract_file <-  system.file("examples/hf205-abstract.md", package = "EML")
+abstract <- set_TextType(abstract_file)
 ```
 
-```{r include=FALSE, echo=FALSE}
-unlink("example.xml")
+### Attribute Metadata from Tables
+
+Attribute metadata can be verbose, and is often defined in separate tables (e.g. separate Excel sheets or `.csv` files).
+Here we use attribute metadata and factor definitions as given from `.csv` files.
+
+```{r}
+attributes <- read.table(system.file("extdata/hf205_attributes.csv", package = "eml2"))
+factors <- read.table(system.file("extdata/hf205_factors.csv", package = "eml2"))
+attributeList <- 
+  set_attributes(attributes, 
+                 factors, 
+                 col_classes = c("character", 
+                                 "Date",
+                                 "Date",
+                                 "Date",
+                                 "factor",
+                                 "factor",
+                                 "factor",
+                                 "numeric"))
 ```
 
-## Reading EML files
+### Data file format
 
-Please see the vignette, [Parsing EML](https://github.com/ropensci/EML/blob/master/vignettes/parsing-EML.Rmd) for a more thorough introduction to reading EML files and extracting relevant metadata.
+Though the `physical` metadata specifying the file format is extremely flexible, the `set_physical` function provides defaults appropriate for `.csv` files. 
+DEVELOPER NOTE: ideally the `set_physical` method should guess the appropriate metadata structure based on the file extension.
 
-## Creating EML files
+```{r}
+physical <- set_physical("hf205-01-TPexp1.csv")
+```
 
-Please see the vignette, [Creating EML](https://github.com/ropensci/EML/blob/master/vignettes/creating-EML.Rmd) for a proper introduction to creating your own EML files.
+## Generic construction
 
-## Working with custom units
+In the `EML` R package, objects for which there is no `set_` method are constructed using the `new()` S4 constructor.
+This provided an easy way to see the list of available slots.
+In `eml2`, all objects are just lists, and so there is no need for special methods.
+We can create any object directly by nesting lists with names corresponding to the EML elements.
+Here we create a `keywordSet` from scratch:
 
-Users may occassionally need to define custom unit types that EML does not know about.
-Please see the vignette, [Working with units](https://github.com/ropensci/EML/blob/master/vignettes/working-with-units.Rmd) for an overview of how units work in EML, how to view the standard unit library, and how to access and define custom units and custom unit types.
+```{r}
+keywordSet <- list(
+    list(
+        keywordThesaurus = "LTER controlled vocabulary",
+        keyword = list("bacteria",
+                    "carnivorous plants",
+                    "genetics",
+                    "thresholds")
+        ),
+    list(
+        keywordThesaurus = "LTER core area",
+        keyword =  list("populations", "inorganic nutrients", "disturbance")
+        ),
+    list(
+        keywordThesaurus = "HFR default",
+        keyword = list("Harvard Forest", "HFR", "LTER", "USA")
+        ))
+```
 
-## Archiving, Sharing, and Publishing EML
+Of course, this assumes that we have some knowledge of what the possible terms permitted in an EML keywordSet are!
+Not so useful for novices.
+We can get a preview of the elements that any object can take using the `emld::template()` option, but this involves a two-part workflow.
+Instead, `eml2` provides generic `construct` methods for all objects.
 
-One of the chief advantages of using EML to manage your own data is the improved ability to search all of your data files, e.g.
-for the ones that have the particular measurements you need on the particular location or species of interest, and be able to share this data either publicly or securely with select collaborators.
-To do so, you will need to upload your EML file and data to a scientific metadata repository that understands EML.
-The vignette, [metadata repositories](https://github.com/ropensci/EML/blob/master/vignettes/metadata-repositories.Rmd) describes how to do this.
+## Constructor methods
 
-## Developer notes
+For instance, the function `eml$creator()` has function arguments corresponding to each possible slot for a creator.
+This means we can rely on *tab completion* (and/or autocomplete previews in RStudio) to see what the possible options are.
+`eml$` functions exist for all complex types.
+If `eml$` does not exist for an argument (e.g.
+there is no `eml$givenName`), then the field takes a simple string argument.
 
-### Creating EML class definitions
+### Creating parties (creator, contact, publisher)
 
-Class definitions (`classes.R`) and methods (`methods.R`) are created programmatically.  From the root of the package, run: `source("inst/create-package/create_package.R")`.
+```{r}
+aaron <- eml$creator(
+  individualName = eml$individualName(
+    givenName = "Aaron", 
+    surName = "Ellison"),
+  electronicMailAddress = "fakeaddress@email.com")
+```
 
-### Meta
+```{r}
+HF_address <- eml$address(
+                  deliveryPoint = "324 North Main Street",
+                  city = "Petersham",
+                  administrativeArea = "MA",
+                  postalCode = "01366",
+                  country = "USA")
+```
 
-Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
+```{r}
+publisher <- eml$publisher(
+                 organizationName = "Harvard Forest",
+                 address = HF_address)
+```
+
+```{r}              
+contact <- 
+  list(
+    individualName = aaron$individualName,
+    electronicMailAddress = aaron$electronicMailAddress,
+    address = HF_address,
+    organizationName = "Harvard Forest",
+    phone = "000-000-0000")
+
+```
+
+### Putting it all together
+
+```{r}
+my_eml <- eml$eml(
+           packageId = uuid::UUIDgenerate(),  
+           system = "uuid",
+           dataset = eml$dataset(
+               title = "Thresholds and Tipping Points in a Sarracenia",
+               creator = aaron,
+               pubDate = "2012",
+               intellectualRights = "http://www.lternet.edu/data/netpolicy.html.",
+               abstract = abstract,
+               keywordSet = keywordSet,
+               coverage = coverage,
+               contact = contact,
+               methods = methods,
+               dataTable = eml$dataTable(
+                 entityName = "hf205-01-TPexp1.csv",
+                 entityDescription = "tipping point experiment 1",
+                 physical = physical,
+                 attributeList = attributeList)
+               ))
+
+```
+
+## Serialize and validate
+
+We can also validate first and then serialize:
+
+```{r}
+eml_validate(my_eml)
+write_eml(my_eml, "eml.xml")
+```
+
+```{r include = FALSE}
+unlink("eml.xml")
+unlink("ex.xml")
+codemetar::write_codemeta()
+```

### 2018-09-14 paper/paper.Rmd
Import eml2

---
title: Working with Ecological Metadata Language Files from R
author:
  - name: Carl Boettiger
    email: cboettig@berkeley.edu
    affiliation: ucb
    footnote: Corresponding Author
address:
  - code: ucb
    address: Department of Environmental Science, Policy, and Management, University of California, Berkeley, Berkeley, CA
abstract: |
  This is the abstract.

  It consists of two paragraphs.

journal: "Methods in Ecology and Evolution"
date: "`r Sys.Date()`"
bibliography: references.bib
output: rticles::elsevier_article
---

---
title: "Project Description"
subtitle: "Realizing the Promise of Machine-Readable Metadata: Ecoinformatics for the Rest of Us"
# author: "Carl Boettiger"
bibliography: "references.bib"
output: pdf_document
header-includes:
   - \usepackage{multicol}
   - \usepackage{bold-extra}

---

Despite a sea change in the availability of public data collected in ecological research, broad synthesis of ecological data remains a rare and largely manual endeavor. Realizing the potential of these unprecedented new data sources requires thorough and accurate machine-readable data and metadata files, and the tools to make use of them.  Here, I describe user friendly R tools that will enable more ecologists to both take advantage of and contribute to growth of well-annotated public ecological data.

Many factors drive the rapid expansion of data availability in ecological research: from drones to satellite images, new technology for gene sequencing, large-scale simulations of climate and other processes [@Overpeck2011], to cultural changes such as the adoption of public data archiving [@Fairbairn2010; @Whitlock2010] and the significant investment in a national network of ecological [NEON, @Keller2008] and ocean [OOI, @Witze2013] observatories.  These factors also reflect the range and heterogeneity in ecological data, from the microscopic to the planetary.  How can researchers combine laboratory measurements of species environmental tolerances, aerial estimates of species occurrence, and simulated projections of changing climate?  How do we even find if the relevant data exist in these rapidly expanding archives?

Rich, machine-readable metadata greatly facilitates discovery, integration and synthesis of the kind of heterogeneous data inherent to ecological research. Metadata provides all of the context required to understand raw data files -- without it data may be little more than meaningless columns of numbers.  Metadata tells us what column labels mean, what units were used. Metadata tells us which species, geographic areas, and time periods are covered by the study, who performed the research, and how.  Data search and discovery requires good metadata: we are looking for measurements of body mass on species found in this region.  Data integration and synthesis require good metadata: we want to combine nine different data sets on different species found in the same region into a single data table, so we must know the species involved, the units of measurement, and so forth.

This metadata is most effective in a format that can be parsed and interpreted by a computer program.  While conventions have primarily relied on scientific papers or the occasional README file to provide this information, machine-readable metadata formats can help ensure these records are complete and accurate and can automate the often tedious process of recording metadata. A machine-readable format can be indexed by specialized search engines, allowing a user to identify all data sets that include measurements of a specific variable on a specific species or from a certain geographic region.  A machine-readable format can even help automate the process of data integration and synthesis, such as standardizing units when combining data columns from different studies.  Such automation is helpful even in small studies, it is crucial if ecological research is to keep up with our rapidly expanding data.

### What is EML?

The Ecological Metadata Language [EML, @Jones2006], originally designed around previous work from the Ecological Society of America [@Michener1997], seeks to tackle this challenge.  EML is an is an XML-based schema which provides a rich, extensible, machine-readable format to describe a wide range of data types, including tabular data, spatial raster and vector data, methods, protocols, software or citations.  EML files can be *validated* against the schema, ensuring this information conforms to the predictable structure and can thus be read by any computer software implementing the EML specification.  Recognizing the importance of thorough, machine-readable metadata, key providers of ecological data such as the NSF-supported Long Term Ecological Research (LTER) sites [@LTER_EML], and the NSF's recently launched National Ecological Observatory Network [NEON, @Keller2008; @NEON] now provide rich EML metadata accompanying their data products.

More than XX individual datasets have been descibed in EML over the past two decades.  EML metadata are required for all data produced by LTER projects (representing about about $12 million NSF support annually), by all NEON sites (about $469 million NSF support) and all projects supported by the NSF Arctic programs (about $40 million annually) (https://www.nsf.gov/pubs/2016/nsf16055/nsf16055.jsp).

Unfortunately, this XML-based format remains ignored or inaccessible to many ecologists, who lack tools and training to capitalize on these advances in computation and data management.  Few are familiar with the skills required to create EML descriptions of their own data, or to fully benefit from the rich information provided in EML [@Hernandez2012].  NEON relies on dedicated professionals to generate EML metadata, while ecologists working with LTER sites benefit from the support of informatics staff in creating EML files describing the data they contribute to LTER data repositories.  Consequently, LTER sites contribute the majority of EML files found in the DataONE Network (which indexes most major repositories of ecological and environmental data), Fig 1.  This highlights the need for tools for ecologists to create EML without assistance from experts or prior expertise in working with XML documents and schema.

![Number of EML metadata files by member nodes for the DataONE Network of the major ecological and environmental data repositories.  EML is used across a wide range of data repositories shown here, all though repositories in which ecoinformatics staff assist in preparation of metadata, such as LTER, provide significantly more metadata than those which primarly host direct community contributions, such as KNB.](img/repo_counts.pdf)

### Relation to existing software tools

<!--
3. How the proposed software compares to alternative or existing elements (including other commercial and research solutions) and what are the limitations of these existing elements.
-->

No extensible tool or utility for reading and writing EML files exists that integrates into the workflow of most ecologists.  Researchers can use the Java application `Morpho`, which provides a graphical user interface (GUI) for creating EML files describing their own data -- this provides an excellent introduction to creating EML, but can be tedious and the interface prohibits automation and templating possible with a programmatic tool. Moreover, `Morpho` can only write new EML files, but not help users parse and analyze existing ones.  The open-source software platform `MetaCat` is available for data repositories which wish to serve and index data files described in EML [@metacat].  This allows data hosts to present users with a nice web-based search interface for discovering data (`MetaCat-UI`) which is already in use by repositories such as KNB and DataONE (see Fig 2).  However, this web-based platform is ultimately software for data repositories to use and deploy, not software for end users.  The environment cannot be easily deployed or extended by research ecologists looking to to work with a generic collection of EML files, cannot perform arbitrary queries or assist in data integration and synthesis tasks.  @Weitz2016 identify these factors (a focus on servers, limited extensibility, and a software stack that is foreign to most researchers in the field of application) as the leading reasons why cyberinfrastructure ends up under-utilized by communities they seek to serve.  Most importantly, `Morpho` and `MetaCat` do little to empower the researcher to take advantage of machine-readable metadata in their own work: instead, these tools work in concert to allow other researchers to take advantage of the user's data -- a noble objective that is often seen as self-sacrificing and a disincentive to data sharing [@VanNoorden2013].  Researchers need an extensible tool that can easily be incorporated into their existing workflow and facilitate their own research.

![A powerful EML metadata-driven search interface from DataONE, based on `MetaCat` platform.  The web interface identifies data files containing measurements of mass of *Cyprinidae* fishes in the geographical regions indicated by the grid. Such a search requires metadata that is thorough, granular, and machine readable. An R package for EML would both make it easier for more researchers to create this EML metadata, perform search queries like that shown here and extend both search and integration tasks through a programmable interface](img/dataone_search.png)

## An R package for EML

This proposal seeks to bridge this gap between the skills and workflows of most research ecologists and the powerful expressiveness of EML-annotated data being generated by NSF's major investments in informatics-supported data generation such as NEON and LTER through the development of an R package for reading, writing and manipulating EML files.  The R language is widely used throughout the ecological research community and has proven an effective platform to engage both users and developers [@Mair2015; @ropensci].  This proposal would support the development and maintenance of the EML R package.  A note on semantics: following the convention for naming R packages dedicated to parsing and serializing a given format, the proposed package name and the data format are both referred to by the acronym EML.  Throughout, I will refer to the data format as simply "EML", and the software package specifically as "the EML R package."

EML parsing in R would enable users to analyze metadata in the same programmatic environment many already use for data analysis.  This kind of programmable environment is essential for users to truly leverage the power of machine-readable metadata in large scale analyses. For instance, this would make it possible for a user to create scripts which can combine data frames coming from different studies which may use different column names or different units for the same variable.  Currently such data integration tasks require manual inspection of the metadata for each data file to verify things like column names and units, a process which does not easily scale to arbitrarily large numbers of separate data files. Or a user might search for a particular phrase being used in the protocol section of a corpus of EML files, or automatically extract citations from metadata files into format recognized by a reference manager.  Once metadata is available in a predictably structured R objects, users can take advantage of the familiar, programmable environment provided by R to automate these tasks.

The ability to create EML from within the R environment is equally powerful.  Again, this is a natural setting to generate metadata since many users are already working with their data in R.  Metadata creation and curation is infamously tedious, making a scripted approach that can automate repetitive tasks a natural choice. Reuse of common metadata elements can avoid tedious data entry (including individual and institutional information, citations, and common data formats), automatically detect metadata from existing data files (such as species coverage or geographic range).  Helper routines within the package can also improve the quality of metadata created. For instance, EML includes a description of the taxonomic coverage of the data being described.  This allows search engines such as MetaCat or the EML R package search functions to identify all data on a given species.  This metadata is most useful when it includes a complete taxonomic description, e.g. taxonomic ranks involved, such that a search for *Aves* returns all bird species.  The EML R package can leverage existing functions in the R package `taxize` to automatically generate the rank classification metadata [@Chamberlain2013].   Package users have the greatest incentive to generate careful and thorough metadata when they can benefit directly from these annotations.  By generating EML files for their own data, users can later leverage the EML parsing functions to help search and synthesize old data sets.  The EML package also integrates with `dataone` R package to automate uploading data and metadata files to a public or private repository archive in the DataONE network, such as the KNB.  This gives users a convenient way to back-up, archive, and share their data files using configurable permissions options and authentication provided by DataONE, and to release data publicly under a permanent DOI, facilitating discovery and citation by other researchers.  Once users can realize these benefits for themselves, opting in to public data sharing can be only a click away, with data already annotated. In this approach, the public benefits of data annotation and data sharing come as a by-product of a tool that puts the user's own needs first. 
These use cases are summarized in Box 1.  This proposal outlines the project plan to implement this functionality into a robust, user-friendly, sustainable software tool.

# Software Architecture and Engineering Process #

The software design follows the format of an R package, the definitive community standard for packaging software to R users. This provides a widely recognized layout for the project, cross platform support, and well defined norms for documentation, testing, distribution, release and evaluation described here.

**Documentation**: The project uses the widely adopted `roxygen2` framework for composing and maintaining all documentation in the software files, consistent with literate programming practice. Keeping code and documentation together helps ensure documentation is up-to-date and consistent with the software.  The automated testing suite also checks the documentation for consistency with the code.  R generates both HTML and PDF forms of documentation from these files.  A package README, Code of Conduct, and Contributing Guide provide further documentation relevant to other developers.  The project also uses the `pkgdown` package to generate a GitHub-hosted website from the package documentation.

**Testing and validation**: The package uses `testthat` [@testthat], a widely used unit-testing framework for R packages, and `covr` [@covr] to measure test coverage.  The test suite is fully automated both locally and through the Travis Continuous Integration (CI) service.

**Security and Trustworthiness**: Open source development practices ensure users and developers can inspect all aspects of the code and tests.  EML files conform to open standards that permit validation and encryption of data.

**Provenance**: Semantic versioning, a NEWS log summarizing changes, and the automatic generation of archival DOIs for each version help researchers both access and cite specific version of the EML package that they use.  Additionally, the EML software itself helps researchers track and document provenance of their data and methods in the machine-readable EML format.

**Reproducibilty**: Our development process is committed to making sure analyses using the EML package remain as reproducible as possible in the future. All effort is made to maintain a backwards-compatible function interface by avoiding changes to the function signature that could break existing code whenever possible.  An increase of the major version number indicates when a new version may break backwards compatibility, and older versions of the software are archived by CRAN and in the Zenodo data repository at the DOI indicated for the version.

\pagebreak

# References

### 2018-09-14 R/eml_validate.R

@@ -1,0 +1,30 @@
+
+
+#' eml_validate
+#'
+#' eml_validate processes an EML document using the XSD schema for the
+#' appropriate version of EML and determines if the document is schema-valid
+#' as defined by the XSD specification
+#' @param eml file path, xml_document,
+#' @param encoding optional encoding for files, default UTF-8.
+#' @param ... additional arguments to eml_write, such as namespaces
+#' @param schema path to schema
+#' @return Whether the document is valid (logical)
+#' @note this function is simply an alias to `eml_validate` in `emld` package
+#' @examples \donttest{
+#'
+#'  f <- system.file("extdata", "example.xml", package = "emld")
+#'
+#'  ## validate file directly from disk:
+#'  eml_validate(f)
+#'
+#'  ## validate an eml object:
+#'  eml <- read_eml(f)
+#'  eml_validate(eml)
+#'
+#' }
+#'
+#' @export
+#' @importFrom emld eml_validate
+eml_validate <- emld::eml_validate
+

### 2018-09-14 R/read_eml.R

@@ -1,0 +1,20 @@
+#' read_eml
+#'
+#' Read an EML file into R as an emld object.
+#' @param x path to an EML file
+#' @param from explicit type for the input format. Possible values:
+#' "xml", "json", "list", or "guess" with "xml" as the default.
+#'
+#' @return an emld object (list / S3 object)
+#' @export
+#' @examples
+#' f <- system.file("extdata", "example.xml", package = "emld")
+#' eml <- read_eml(f)
+#' @export
+#' @importFrom emld as_emld
+
+read_eml <- function(x, from = "xml"){
+  emld::as_emld(x, from = from)
+}
+
+

### 2018-09-14 R/set_unitList.R

@@ -1,0 +1,140 @@
+
+
+#' set_unitList
+#'
+#' Define custom units, including new unitTypes.  Note that it is not necessary to define
+#' most common units.
+#'
+#' @param units a data.frame describing the custom units, see details.
+#' @param unitTypes optional, a data.frame defining any additional unitTypes not already defined
+#' @param as_metadata logical, default FALSE. If true, returns an `additionalMetadata` element, see below.
+#' @return unitList list object
+#'
+#'
+#' @export
+#' @details  The units data.frame must have the following columns:
+#'  - id: the referenced name of unit (singular). e.g. 'meter', 'second'
+#'  - unitType: the base type of unit, e.g. 'length'.  If not from a standard type, a new unitType must be provided
+#'  - multiplierToSI: the multiplicative constant to convert to the SI unit.
+#'  - parentSI: the name of the parent SI unit, e.g. second.
+#'  - description: a text string describing the unit of measure.
+
+#'  The following columns are optional:
+#'  - name: usually the same as the id of the unit, e.g. second
+#'  - abbreviation: common abbreviation, e.g. s
+#'  - constantToSI: an additive constant to convert to the equivalent SI unit. If not given, default is "0"
+#'
+#' In practice, researchers may save these tables of custom units they frequently use in an external .csv
+#' or other format and read them in to R for ready re-use.
+#'
+#' The unitType table must have the following columns:
+#'  - id: the name by which the unitType is referred to.
+#'  - name: optional, default is same as the id
+#'  - dimension: name of a base dimension of the unit
+#'  - power: the power to which the dimension is raised (NA implies power of 1)
+#'
+#' @examples
+#'  ## create the "unitType" table for custom unit
+#'  id = c("speed", "speed", "acceleration", "acceleration", "frequency")
+#'  dimension = c("length", "time", "length", "time", "time")
+#'  power = c(NA, "-1", NA, "-2", "-1")
+#'  unitTypes <- data.frame(id = id, dimension = dimension,
+#'                          power = power, stringsAsFactors = FALSE)
+#'
+#'  ## Create the units table
+#'  id = c("minute", "centimeter")
+#'  unitType = c("time", "length")
+#'  parentSI = c("second", "meter")
+#'  multiplierToSI = c("0.0166", "1")
[…]

### 2018-09-14 R/shiny_attributes.R

@@ -1,0 +1,413 @@
+#' Create/Edit EML attributes
+#'
+#' @description Create/edit EML attributes, custom units, and factors in a shiny environment.
+#'
+#' @details Attributes can be created from scratch using \code{shiny_attributes()}.
+#' Or an existing attribute table can be edited using \code{shiny_attributes(NULL, attributes)}.
+#' Or new attributes can be created from a data table using \code{shiny_attributes(data, NULL)}.
+#' If attributes are created from a data table, fields such as `attributeName` and `numberType` will be automatically
+#' completed based on the attributes within the data table.
+#' If both existing attributes and data table are entered (i.e. \code{shiny_attributes(data, attributes)}),
+#' any automatically generated fields based attributes within the data table **will not** override any non-empty fields in the
+#' entered attributes
+#'
+#' @param data (data.frame) the data.frame of data that needs an attribute table
+#' @param attributes (data.frame) an existing attributes table
+#' @importFrom utils write.csv
+#'
+#' @examples
+#' \dontrun{
+#' # from scratch
+#' out <- shiny_attributes(NULL, NULL)
+#'
+#' # from data
+#' data <- iris
+#' out <- shiny_attributes(data, NULL)
+#'
+#' # from exisiting attributes
+#' file <- system.file("xsd/test/eml-datasetWithAttributelevelMethods.xml", package = "EML")
+#' eml <- read_eml(file)
+#' x <- eml$dataset$dataTable$attributeList
+#' df <- get_attributes(x, eml)
+#' out <- shiny_attributes(NULL, df$attributes)
+#'
+#' # from attributes and data
+#' out <- shiny_attributes(data, df$attributes)
+#'}
+#' @export
+shiny_attributes <- function(data = NULL, attributes = NULL) {
+
+  if (!requireNamespace("shiny", quietly = TRUE)) {
+    stop(call. = FALSE,
+         "shiny_attributes requires the
+         'shiny' package to be installed. Install shiny package for this functionality.")
+  }
+
+  if (!requireNamespace("shinyjs", quietly = TRUE)) {
+    stop(call. = FALSE,
+         "shiny_attributes requires the
+         'shinyjs' package to be installed. Install shinyjs package for this functionality.")
+  }
+
+  if (!requireNamespace("htmlwidgets", quietly = TRUE)) {
+    stop(call. = FALSE,
+         "shiny_attributes requires the
+         'htmlwidgets' package to be installed. Install htmlwidgets package for this functionality.")
+  }
+
[…]

### 2018-09-14 R/template_constructor.R

@@ -1,0 +1,24 @@
+
+# eml_creator <- template_constructor("creator")
+template_constructor <- function(object){
+  args <- emld::template(object)
+  f <- paste0("function(",
+         paste(names(args), "=", args, collapse = ", "),
+         "){ list(",
+         paste(names(args), "=", names(args), collapse = ","),
+         ")}")
+  eval(parse(text = f))
+}
+
+
+#who <- readLines(system.file("extdata/complexTypes.txt", package="eml2"))
+#names(who) <- who
+#construct <- lapply(who, template_constructor)
+
+#' eml
+#'
+#' @docType data
+#' @keywords datasets
+#' @name eml
+#' @format A list with constructor functions
+NULL

### 2018-09-14 tests/testthat/test-additional-tests.R

@@ -1,0 +1,43 @@
+library(testthat)
+library(eml2)
+
+testthat::context("Additional tests")
+
+test_that("template constructor",{
+
+  creator <- eml2:::template_constructor("creator")
+  expect_is(creator, "function")
+})
+
+test_that("we can validate an xml_document", {
+  f <- system.file("extdata", "example.xml", package = "emld")
+  x <- xml2::read_xml(f)
+  expect_true(eml_validate(x))
+})
+
+
+test_that("set_physical can compute file size automatically", {
+  skip_if_not(file.exists("hf205-01-TPexp1.csv"))
+  expect_message(physical <-
+                   set_physical("hf205-01-TPexp1.csv"),
+                "calculated file size")
+
+})
+
+test_that("set_method with additional input files", {
+  coverage <-
+    set_coverage(date = c('2012-06-01','2013-12-31'),
+                 sci_names = "Sarracenia purpurea",
+                 geographicDescription = "text",
+                 west = -122.44, east = -117.15,
+                 north = 37.38, south = 30.00,
+                 altitudeMin = 160, altitudeMaximum = 330,
+                 altitudeUnits = "meter")
+
+  file <- system.file("examples/hf205-methods.docx", package = "EML")
+  methods <- set_methods(file,
+                         qualityControl_file = file,
+                         sampling_file = file,
+                         sampling_coverage = coverage)
+
+})

### 2018-09-14 tests/testthat/test-set_attributes.R

@@ -1,0 +1,617 @@
+context("set_attributes")
+
+
+test_that("we can have numeric data with bounds
+          where some bounds are missing",
+          {
+            df <- data.frame(
+              attributeName = "svl",
+              attributeDefinition = "Snout-vent length",
+              unit = "meter",
+              numberType = "real",
+              minimum = "0",
+              maximum = NA,
+              stringsAsFactors = FALSE
+            )
+            attributeList <-
+              set_attributes(df, col_classes = "numeric")
+
+            dataTable <- list(
+              attributeList = attributeList,
+              entityName = "file.csv",
+              entityDescription = "csv file containing the data",
+              physical = set_physical("file.csv")
+            )
+
+            me <-
+              list(individualName =
+                     list(givenName = "Carl",
+                          surName = "Boettiger"))
+
+            dataset <-
+              list(
+                title = "Example EML",
+                creator = me,
+                contact = me,
+                dataTable = dataTable
+              )
+            eml <-
+              list(packageId = "123",
+                   system = "uuid",
+                   dataset = dataset)
+
+            write_eml(eml, "eml.xml")
+            expect_true(eml_validate("eml.xml"))
+
+          })
+
+test_that("The set_attributes function
+          works for the vignette example", {
+            attributes <-
+              data.frame(
+                attributeName = c(
+                  "run.num",
+                  "year",
+                  "day",
+                  "hour.min",
+                  "i.flag",
+                  "variable",
+                  "value.i",
+                  "length"
+                ),
+                formatString = c(NA,
+                                 "YYYY",
+                                 "DDD",
+                                 "hhmm",
+                                 NA,
+                                 NA,
+                                 NA,
+                                 NA),
+                definition = c("which run number",
+                               NA,
+                               NA,
+                               NA,
+                               NA,
+                               NA,
+                               NA,
+                               NA),
+                unit = c(NA,
+                         NA,
+                         NA,
+                         NA,
+                         NA,
+                         NA,
+                         NA,
+                         "meter"),
+                attributeDefinition = c(
+          "which run number (=block). Range: 1 - 6. (integer)",
+          "year, 2012",
+          "Julian day. Range: 170 - 209.",
+          "hour and minute of observation. Range 1 - 2400 (integer)",
[…]

### 2018-09-14 old-vignettes/parsing-EML.Rmd
import and nativize eml2  as EML

---
title: "Reading and parsing EML"
author: "Carl Boettiger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parsing EML}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# listviewer

If the `listviewer` package is installed it can provide a convenient way to view and edit EML:

```{r, eval = FALSE}
f <- system.file("tests/eml.xml", package = "emld")
eml <- read_eml(f)
listviewer::jsonedit(eml)
```

# Parsing an EML file

```{r}
f <- system.file("xsd/test/eml-i18n.xml", package = "EML")
eml <- read_eml(f)
```

Here we request all `temporalCoverage` elements occurring in the anywhere in the `eml` document:

```{r}
temporalCoverage <- eml_get(eml, "temporalCoverage")
temporalCoverage
```

Any EML element can be extracted in this way. Let's try an example metadata file for a `dataset` that documents 11 seperate `dataTables`:

```{r}
hf001 <- system.file("examples/hf001.xml", package="EML") 
eml_HARV <- read_eml(hf001)
```

How many `dataTable` entities are there in this dataset?

```{r}
dt <- eml_get(eml_HARV, "dataTable")
length(dt)
```

We can iterate over our list of `dataTable` elements to extract relevant metadata, such as the `entityName` or the download `url`:

```{r}
entities <- eml_get(eml_HARV, "dataTable.entityName")
urls <- sapply(dt, eml_get, "url")
```

Note that the latter example is the same as providing the more verbose arbument that specificies exactly where the `url` of interest is located:

```{r}
urls <- sapply(dt, function(x) x@physical[[1]]@distribution[[1]]@online@url)
```

this verbose syntax can be useful if there are multiple `url` elements in each `dataTable` metadata, and we are trying to get only certain ones and not others. Specifying the exact path in this way can also improve the speed of the command.  For these reasons, programmatic use should consider this format, while the much simpler `eml_get` example shown above is practical for most interactive applications.

Although the default return type for `eml_get` is just the S4 object (whose `print` method displays the corresponding XML structure used to represent that metadata), for a few commonly accessed complex elements, `eml_get` returns a more convenient `data.frame`.  For instance, the `attributeList` describing the metadata for every column in an EML document is returned as a pair of `data.frame`s, one for all the attributes, and an second optional `data.frame` defnining the levels for the factors, if any are used.  Let's take a look:

Here we get the `attributeList` for each `dataTable` in the dataset. We check the length to confirm we get one `attributeList` for each `dataTable`

```{r}
attrs <- eml_get(dt, "attributeList") 
length(attrs)
attrs[[1]]
```

(Note, we could have passed this argument the original `eml_HARV` instead of `dt` here, since we know all `attributeList` elements are inside `dataTable` elements, but this is a bit more explicit and a bit faster.)

This returned `data.frame` object containing the attribute metadata for the first table (hence the `[[1]]`, though `attrs` contains this metadata for all 11 tables now.)  This is the same result we would have gotten using the more explicit call to the helper function `get_attributes()`:

```{r}
get_attributes(eml_HARV@dataset@dataTable[[1]]@attributeList)
```

### 2018-09-14 R/get_unitList.R

@@ -13,7 +13,8 @@
 #'
 #' # Read in additional units defined in a EML file
 #' \donttest{
-#' f <- system.file("xsd/test/eml-datasetWithUnits.xml", package = "EML")
+#' f <- system.file("tests", options("emld_db"),
+#' "eml-datasetWithUnits.xml", package = "emld")
 #' eml <- read_eml(f)
 #' unitList <- get_unitList(eml)
 #'
@@ -25,8 +26,10 @@ get_unitList <-
   function(x = NULL) {
 
     if(is.null(x)){
-     unitList <- read_eml(system.file("xsd/eml-2.1.1/eml-unitDictionary.xml",
-                         package = "EML"))
+     unitList <- read_eml(system.file("tests",
+                                      options("emld_db"),
+                                      "eml-unitDictionary.xml",
+                         package = "emld"))
     } else {
       unitList <- eml_get(x, "unitList")
     }

### 2018-09-14 data-raw/construct_complexTypes.R

@@ -13,9 +13,9 @@ who <- names(db)
 writeLines(who, "inst/extdata/complexTypes.txt")
 
 
-#who <- readLines(system.file("extdata/complexTypes.txt", package="eml2"))
+#who <- readLines(system.file("extdata/complexTypes.txt", package="EML"))
 names(who) <- who
-eml <- lapply(who, eml2:::template_constructor)
+eml <- lapply(who, EML:::template_constructor)
 
 
 devtools::use_data(eml, overwrite=TRUE)

### 2018-09-14 R/eml_get.R
restyle via styler

@@ -8,12 +8,13 @@
 #' "xml", "json", "list", or "guess" with "list" as the default.
 #' @param ... additional arguments
 #'
-#' @examples \donttest{
+#' @examples
+#' \donttest{
 #' f <- system.file("tests", options("emld_db"), "eml-datasetWithUnits.xml", package = "emld")
 #' eml <- read_eml(f)
 #' eml_get(eml, "physical")
 #' eml_get(eml, "attributeList")
-#'
+#' 
 #' ## The first argument need not be an "eml" class, it could be a child element; e.g.
 #' eml_get(eml$dataset$dataTable, "physical")
 #' }
@@ -21,7 +22,7 @@
 #' @importFrom jqr jq combine
 #' @importFrom emld as_json as_emld
 #' @importFrom jsonlite fromJSON
-eml_get <- function(x, element, from = "list", ...){
+eml_get <- function(x, element, from = "list", ...) {
   doc <- as.character(emld::as_json(emld::as_emld(x, from = from)))
   out <- jqr::jq(doc, paste0("..|.", element, "? // empty"))
   json <- jqr::combine(out)

### 2018-09-14 R/get_unitList.R

@@ -10,36 +10,40 @@
 #' @export
 #'
 #' @examples
-#'
+#' 
 #' # Read in additional units defined in a EML file
 #' \donttest{
 #' f <- system.file("tests", options("emld_db"),
-#' "eml-datasetWithUnits.xml", package = "emld")
+#'   "eml-datasetWithUnits.xml",
+#'   package = "emld"
+#' )
 #' eml <- read_eml(f)
 #' unitList <- get_unitList(eml)
-#'
+#' 
 #' ## Read in the definitions of standard units:
 #' get_unitList()
 #' }
-#'
+#' 
 get_unitList <-
   function(x = NULL) {
-
-    if(is.null(x)){
-     unitList <- read_eml(system.file("tests",
-                                      options("emld_db"),
-                                      "eml-unitDictionary.xml",
-                         package = "emld"))
+    if (is.null(x)) {
+      unitList <- read_eml(system.file("tests",
+        options("emld_db"),
+        "eml-unitDictionary.xml",
+        package = "emld"
+      ))
     } else {
       unitList <- eml_get(x, "unitList")
     }
-    list(units = get_unit(unitList$unit),
-         unitTypes = get_unitType(unitList$unitType))
+    list(
+      units = get_unit(unitList$unit),
+      unitTypes = get_unitType(unitList$unitType)
+    )
   }
 
 
 
-get_unit <- function(unit){
+get_unit <- function(unit) {
   ## Unnested structure, easy to rectangle
   fromJSON(toJSON(unit))
 }
@@ -47,13 +51,8 @@ get_unit <- function(unit){
 #' @importFrom jsonlite toJSON fromJSON
 get_unitType <- function(unitType) {
   ## Nested data structure, rectangle via jq
-    y <- toJSON(unitType)
+  y <- toJSON(unitType)
 
-    tmp <- jqr::jq(as.character(y), '.[] | {
-            id,
-            name,
-            dimension: .dimension[].name?,
-            power: .dimension[].power?
-            }')
-    fromJSON(jqr::combine(tmp))
+  tmp <- jqr::jq(as.character(y), ".[] | {\n            id,\n            name,\n            dimension: .dimension[].name?,\n            power: .dimension[].power?\n            }")
+  fromJSON(jqr::combine(tmp))
 }

### 2018-09-14 R/set_coverage.R

@@ -28,38 +28,42 @@
 #' @export
 #'
 #' @examples
-#'coverage <-
-#'  set_coverage(begin = '2012-06-01', end = '2013-12-31',
-#'               sci_names = "Sarracenia purpurea",
-#'               geographicDescription = "California coast, down through Baja, Mexico",
-#'               west = -122.44, east = -117.15,
-#'               north = 37.38, south = 30.00)
+#' coverage <-
+#'   set_coverage(
+#'     begin = "2012-06-01", end = "2013-12-31",
+#'     sci_names = "Sarracenia purpurea",
+#'     geographicDescription = "California coast, down through Baja, Mexico",
+#'     west = -122.44, east = -117.15,
+#'     north = 37.38, south = 30.00
+#'   )
 set_coverage <-
   function(beginDate = character(),
-           endDate = character(),
-           date = character(),
-           sci_names = character(),
-           geographicDescription = character(),
-           westBoundingCoordinate = numeric(),
-           eastBoundingCoordinate = numeric(),
-           northBoundingCoordinate = numeric(),
-           southBoundingCoordinate = numeric(),
-           altitudeMinimum = numeric(),
-           altitudeMaximum = numeric(),
-           altitudeUnits = character()) {
-
+             endDate = character(),
+             date = character(),
+             sci_names = character(),
+             geographicDescription = character(),
+             westBoundingCoordinate = numeric(),
+             eastBoundingCoordinate = numeric(),
+             northBoundingCoordinate = numeric(),
+             southBoundingCoordinate = numeric(),
+             altitudeMinimum = numeric(),
+             altitudeMaximum = numeric(),
+             altitudeUnits = character()) {
     out <-
-      list(geographicCoverage = set_geographicCoverage(
-              geographicDescription,
-              westBoundingCoordinate,
-              eastBoundingCoordinate,
-              northBoundingCoordinate,
-              southBoundingCoordinate,
-              altitudeMinimum,
-              altitudeMaximum,
-              altitudeUnits),
-           temporalCoverage = set_temporalCoverage(beginDate, endDate, date),
-           taxonomicCoverage = set_taxonomicCoverage(sci_names))
+      list(
[…]

### 2018-09-14 R/set_responsibleParty.R

@@ -19,17 +19,17 @@
 #' carl <- set_responsibleParty(as.person("Carl Boettiger <cboettig@ropensci.org>"))
 #' matt <- set_responsibleParty("Matthew", "Jones", email = "mbjones@@nceas.ucsb.edu")
 set_responsibleParty <-
-  function (givenName = NULL,
-            surName = NULL,
-            organizationName = NULL,
-            positionName = NULL,
-            address = NULL,
-            phone = NULL,
-            electronicMailAddress = NULL,
-            onlineUrl = NULL,
-            userId = NULL,
-            id = NULL,
-            email = NULL) {
+  function(givenName = NULL,
+             surName = NULL,
+             organizationName = NULL,
+             positionName = NULL,
+             address = NULL,
+             phone = NULL,
+             electronicMailAddress = NULL,
+             onlineUrl = NULL,
+             userId = NULL,
+             id = NULL,
+             email = NULL) {
     UseMethod("set_responsibleParty", givenName)
   }
 
@@ -41,17 +41,17 @@ set_responsibleParty.person <- function(givenName, ...) {
 
 #' @export
 set_responsibleParty.character <-
-  function (givenName = NULL,
-            surName = NULL,
-            organizationName = NULL,
-            positionName = NULL,
-            address = NULL,
-            phone = NULL,
-            electronicMailAddress = NULL,
-            onlineUrl = NULL,
-            userId = NULL,
-            id = NULL,
-            email = NULL) {
+  function(givenName = NULL,
+             surName = NULL,
+             organizationName = NULL,
+             positionName = NULL,
+             address = NULL,
+             phone = NULL,
+             electronicMailAddress = NULL,
+             onlineUrl = NULL,
+             userId = NULL,
+             id = NULL,
+             email = NULL) {
     emld::as_emld(list(
       individualName = list(givenName = givenName, surName = surName),
       organizationName = organizationName,
@@ -59,14 +59,13 @@ set_responsibleParty.character <-
       address = address,
       phone = phone,
       electronicMailAddress =
-        if(is.null(electronicMailAddress)){
+        if (is.null(electronicMailAddress)) {
           email
         } else {
           electronicMailAddress
         },
       onlineUrl = onlineUrl,
       userId = userId,
-      "@id" = id)
-    )
-
+      "@id" = id
+    ))
   }

### 2018-09-14 R/template_constructor.R

@@ -1,19 +1,21 @@
 
 # eml_creator <- template_constructor("creator")
-template_constructor <- function(object){
+template_constructor <- function(object) {
   args <- emld::template(object)
-  f <- paste0("function(",
-         paste(names(args), "=", args, collapse = ", "),
-         "){ list(",
-         paste(names(args), "=", names(args), collapse = ","),
-         ")}")
+  f <- paste0(
+    "function(",
+    paste(names(args), "=", args, collapse = ", "),
+    "){ list(",
+    paste(names(args), "=", names(args), collapse = ","),
+    ")}"
+  )
   eval(parse(text = f))
 }
 
 
-#who <- readLines(system.file("extdata/complexTypes.txt", package="EML"))
-#names(who) <- who
-#construct <- lapply(who, template_constructor)
+# who <- readLines(system.file("extdata/complexTypes.txt", package="EML"))
+# names(who) <- who
+# construct <- lapply(who, template_constructor)
 
 #' eml
 #'

### 2018-09-14 R/write_eml.R

@@ -25,7 +25,7 @@ write_eml <- function(eml,
                       ns = "eml",
                       ...) {
   ## FIXME, insist on a packageId, or consider validation which can ignore id?
-  if(is.null(eml$packageId)){
+  if (is.null(eml$packageId)) {
     eml$packageId <- uuid::UUIDgenerate()
     eml$system <- "uuid"
   }

### 2018-09-14 data-raw/construct_complexTypes.R

@@ -1,21 +1,22 @@
 ## Small test case
-#who <- list("creator", "individualName", "contact", "dataTable")
-#names(who) <- who
-#construct <- lapply(who, template_constructor)
+# who <- list("creator", "individualName", "contact", "dataTable")
+# names(who) <- who
+# construct <- lapply(who, template_constructor)
 
 options("emld_db" = "eml-2.1.1")
 ## Full database
 db <- purrr::compact(
-  jsonlite::read_json("https://raw.githubusercontent.com/cboettig/emld/master/data-raw/eml_db.json"))
+  jsonlite::read_json("https://raw.githubusercontent.com/cboettig/emld/master/data-raw/eml_db.json")
+)
 db["UrlType"] <- NULL # 'function' as argument, not needed
 db["url"] <- NULL
 who <- names(db)
 writeLines(who, "inst/extdata/complexTypes.txt")
 
 
-#who <- readLines(system.file("extdata/complexTypes.txt", package="EML"))
+# who <- readLines(system.file("extdata/complexTypes.txt", package="EML"))
 names(who) <- who
 eml <- lapply(who, EML:::template_constructor)
 
 
-devtools::use_data(eml, overwrite=TRUE)
+devtools::use_data(eml, overwrite = TRUE)

### 2018-09-14 data-raw/standardUnits.R

@@ -1,2 +1,2 @@
 standardUnits <- EML::get_unitList()
-devtools::use_data(standardUnits, internal=TRUE)
+devtools::use_data(standardUnits, internal = TRUE)

### 2018-09-14 tests/testthat/test-eml2.R

@@ -4,13 +4,15 @@ library(emld)
 geographicDescription <-
   "Harvard Forest Greenhouse, Tom Swamp Tract (Harvard Forest)"
 coverage <-
-  set_coverage(begin = '2012-06-01', end = '2013-12-31',
-               sci_names = "Sarracenia purpurea",
-               geographicDescription = geographicDescription,
-               west = -122.44, east = -117.15,
-               north = 37.38, south = 30.00,
-               altitudeMin = 160, altitudeMaximum = 330,
-               altitudeUnits = "meter")
+  set_coverage(
+    begin = "2012-06-01", end = "2013-12-31",
+    sci_names = "Sarracenia purpurea",
+    geographicDescription = geographicDescription,
+    west = -122.44, east = -117.15,
+    north = 37.38, south = 30.00,
+    altitudeMin = 160, altitudeMaximum = 330,
+    altitudeUnits = "meter"
+  )
 
 
 methods_file <- system.file("examples/hf205-methods.docx", package = "emld")
@@ -18,8 +20,10 @@ methods <- set_methods(methods_file)
 
 
 
-R_person <- person("Aaron", "Ellison", ,"fakeaddress@email.com", "cre",
-                   c(ORCID = "0000-0003-4151-6081"))
+R_person <- person(
+  "Aaron", "Ellison", , "fakeaddress@email.com", "cre",
+  c(ORCID = "0000-0003-4151-6081")
+)
 aaron <- as_emld(R_person)
 
 ## ------------------------------------------------------------------------
@@ -34,12 +38,14 @@ HF_address <- list(
   city = "Petersham",
   administrativeArea = "MA",
   postalCode = "01366",
-  country = "USA")
+  country = "USA"
+)
 
 ## ------------------------------------------------------------------------
 publisher <- list(
   organizationName = "Harvard Forest",
-  address = HF_address)
+  address = HF_address
+)
 
 ## ------------------------------------------------------------------------
 contact <-
@@ -48,29 +54,35 @@ contact <-
     electronicMailAddress = aaron$electronicMailAddress,
     address = HF_address,
     organizationName = "Harvard Forest",
-    phone = "000-000-0000")
+    phone = "000-000-0000"
+  )
 
 
 ## ------------------------------------------------------------------------
 attributes <- read.table(system.file("extdata/hf205_attributes.csv",
-                                     package = "EML"))
+  package = "EML"
+))
 factors <- read.table(system.file("extdata/hf205_factors.csv",
[…]

### 2018-09-14 tests/testthat/test-eml_get.R

@@ -8,12 +8,11 @@ test_that("eml_get works on trivial calls", {
   x <- eml_get(eml, "physical")
   expect_s3_class(x, "emld")
   y <- eml_get(eml, "attributeList")
-  expect_length(y,3)
+  expect_length(y, 3)
   expect_s3_class(y, "emld")
 })
 
 test_that("eml_get works on more calls", {
-
   my_eml <- read_eml(f)
   y <- eml_get(my_eml, "attributeName")
   expect_s3_class(y, "emld")

### 2018-11-12 README.Rmd
link to rendered vignettes #248

--- 2017-07-21 README.Rmd
@@
 
-Please see the vignette, [Parsing EML](https://github.com/ropensci/EML/blob/master/vignettes/parsing-EML.Rmd) for a more thorough introduction to reading EML files and extracting relevant metadata.
+Please see the vignette, [Parsing EML](https://ropensci.github.io/EML/articles/parsing-EML.html) for a more thorough introduction to reading EML files and extracting relevant metadata.
 
@@
 
-Please see the vignette, [Creating EML](https://github.com/ropensci/EML/blob/master/vignettes/creating-EML.Rmd) for a proper introduction to creating your own EML files.
+Please see the vignette, [Creating EML](https://ropensci.github.io/EML/articles/creating-EML.html) for a proper introduction to creating your own EML files.
 
@@
 Users may occassionally need to define custom unit types that EML does not know about.
-Please see the vignette, [Working with units](https://github.com/ropensci/EML/blob/master/vignettes/working-with-units.Rmd) for an overview of how units work in EML, how to view the standard unit library, and how to access and define custom units and custom unit types.
+Please see the vignette, [Working with units](https://ropensci.github.io/EML/articles/working-with-units.html) for an overview of how units work in EML, how to view the standard unit library, and how to access and define custom units and custom unit types.
 
@@
 To do so, you will need to upload your EML file and data to a scientific metadata repository that understands EML.
-The vignette, [metadata repositories](https://github.com/ropensci/EML/blob/master/vignettes/metadata-repositories.Rmd) describes how to do this.
+The vignette, [metadata repositories](https://ropensci.github.io/EML/articles/metadata-repositories.html) describes how to do this.
 

### 2018-11-27 R/eml_view.R
EML 2: switching over to S3 system with JSON-LD powered parser, emld (#251)
* clean

* Import eml2

* import and nativize eml2  as EML

* add inst/examples files

* restyle via styler

* tweaking style

* whitespace

* update proj file

* fix tests

* Rproj file confuses appveyor wtf

* update tests and data-raw scripts

* appveyor

* appveyor

* update ignore list

* skip certaint tests on old pandoc

* Goodness gracious appveyor.

Note that these tests pass just fine on a local Windows machine with recent Pandoc / RStudio

@@ -1,55 +1,0 @@
-#' View an EML
-#'
-#' @param eml_path Path to the EML one wants to view
-#' @param ... additional arguments to \code{\link{read_xml}}
-#' @return Nothing, the EML is viewed in the Viewer.
-#' @export
-#' @importFrom xml2 as_list
-#' @details The function needs the \code{listviewer} package which you can install from CRAN
-#' via \code{install.packages("listviewer"))}. The function uses \code{listviewer::jsonedit}
-#' which gives a crude but useful viewing of the EML
-#'
-#' @examples
-#' f <- system.file("xsd/test", "eml-datasetWithUnits.xml", package = "EML")
-#' eml_view(f)
-#'  f <- system.file("xsd/test", "eml.xml", package = "EML")
-#'  eml_view(f)
-#'  f <- system.file("xsd/test", "eml-sample.xml", package = "EML")
-#'  eml_view(f)
-#'  f <- system.file("xsd/test", "eml-datasetWithCitation.xml", package = "EML")
-#'  eml_view(f)
-#'  f <- system.file("examples", "example-eml-2.1.0.xml", package = "EML")
-#'  eml_view(f)
-eml_view <- function(eml_path, ...){
-  if (!requireNamespace("listviewer", quietly = TRUE)) {
-      message("eml_view needs the listviewer package, \n
-              Install it via install.packages('listviewer')")
-    return(NULL)
-    }
-
-  doc <- read_xml(eml_path, ...)
-  eml_list <- xml2::as_list(doc)
-
-  # no unnamed elements please
-  if (any(names(eml_list)=="")){
-    names(eml_list)[names(eml_list) == ""] <- "noname"
-  }
-
-  eml_list <- name_all(eml_list)
-
-  ## shouldn't call function from a Suggested package, see get_TextType for how we use rmarkdown fns
-  jsonedit <- getExportedValue("listviewer", "jsonedit")
-  jsonedit(eml_list)
-}
-
-# code adapted from https://stackoverflow.com/questions/29818918/looping-nested-lists-in-r
-name_all <- function(l){
-  lapply(l, function(x){
-    if(is.list(x) && any(names(x)=="")){
-      names(x)[names(x) == ""] <- "noname"
-      x
-    }  else{
-      if(is.list(x)) name_all(x) else x
-    }
-  } )
-}

### 2018-11-27 R/get_TextType.R

@@ -1,82 +1,0 @@
-
-
-
-#' get_TextType
-#'
-#' Render a TextType node into HTML or some other format
-#'
-#' @param node any TextType node
-#' @param to desired format, default is html, but can be any type supported by pandoc (docx, md, etc)
-#' @param output name of the desired output file
-#' @param view if HTML, do we want to open result in browser?
-#' @importFrom utils browseURL
-#' @importFrom xml2 xml_dtd xml_add_child xml_new_root write_xml xml_children
-#' @return creates a file requested.
-#' @export
-#'
-#' @examples
-#' \donttest{
-#'
-#' ## Convert an EML abstract to markdown
-#' f <- system.file("examples/hf205.xml", package = "EML")
-#' eml <- read_eml(f)
-#' abstract <- eml_get(eml, "abstract")
-#' get_TextType(abstract[[1]], "markdown", "abstract.markdown")
-#' readLines("abstract.markdown")
-#' unlink("abstract.markdown") # tidy up
-#'
-#' ## Turn a docx file into EML abstract and preview at HTML
-#' f <- system.file("examples/hf205-abstract.docx", package = "EML")
-#' a <- as(set_TextType(f), "abstract")
-#' get_TextType(a)
-#' }
-#'
-get_TextType <-
-  function(node,
-           to = "html",
-           output = tempfile(class(node), fileext = paste0(".", to)),
-           view = TRUE) {
-
-    ## eml_get returns a list of nodes always.
-    if(!isS4(node) && length(node) == 1)
-      node <- node[[1]]
-
-    # serialize sections in ListOfsection or paras from ListOfpara into XML document, save, rmarkdown into desired format
-    x <- xml2::xml_contents(EML:::s4_to_xml(node, root = xml2::xml_new_root("root")))
-
-    if (!requireNamespace("rmarkdown", quietly = TRUE)) {
-      stop("rmarkdown package required to convert to Docbook format",
-           call. = FALSE)
-    }
-
-    pandoc_convert <-
-      getExportedValue("rmarkdown", "pandoc_convert")
-
-    wd <- getwd()
-    dir <- tempdir()
-    file.copy(output, file.path(dir, basename(output)), overwrite = TRUE)
-    setwd(dir)
-    docbook_file <- tempfile(pattern = "docbook", tmpdir = ".", fileext = ".db")
-
-    doctype <-
-      xml2::xml_new_root(xml2::xml_dtd("section",
[…]

### 2018-11-27 R/set_physical.R

@@ -1,7 +1,7 @@
 #' set_physical
 #'
-#' Will calculate the file size, checksum, and checksum algorithm automatically
-#' if the argument \code{objectName} is a file that exists.
+#' Will calculate the file size, checksum, and checksum authentication method
+#' algorithm automatically if the argument \code{objectName} is a file that exists.
 #'
 #' @param objectName name for the object, usually a filename like "hf205-1.csv"
 #' @param id optional, an id value for the <physical> element in EML, for use in referencing
@@ -26,36 +26,33 @@
 #'
 #' @examples
 #' set_physical("hf205-01-TPexp1.csv")
-# FIXME set recordDelimiter based on user's system?
-# FIXME richer distribution options? use set_distribution at top level?
+#' # FIXME set recordDelimiter based on user's system?
+#' # FIXME richer distribution options? use set_distribution at top level?
 set_physical <- function(objectName,
                          id = character(),
                          numHeaderLines = character(),
                          numFooterLines = character(),
-                         recordDelimiter = "\\r\\n",
+                         recordDelimiter = "\\n\\r",
                          fieldDelimiter = ",",
                          collapseDelimiters = logical(),
                          literalCharacter = character(),
                          quoteCharacter = character(),
                          attributeOrientation = "column",
-                         size = character(),
+                         size = NULL,
                          sizeUnit = "bytes",
-                         authentication = character(),
-                         authMethod = character(),
+                         authentication = NULL,
+                         authMethod = NULL,
                          characterEncoding = character(),
                          encodingMethod = character(),
                          compressionMethod = character(),
                          url = character()) {
-  dataFormat <- new(
-    "dataFormat",
-    textFormat = new(
-      "textFormat",
+  dataFormat <- list(
+    textFormat = list(
       numHeaderLines = numHeaderLines,
       numFooterLines = numFooterLines,
       recordDelimiter = recordDelimiter,
       attributeOrientation = attributeOrientation,
-      simpleDelimited = new(
-        "simpleDelimited",
+      simpleDelimited = list(
         fieldDelimiter = fieldDelimiter,
         collapseDelimiters = as.character(collapseDelimiters),
         quoteCharacter = quoteCharacter,
@@ -70,26 +67,36 @@ set_physical <- function(objectName,
 
   if (file.exists(objectName)) {
     if (length(size) == 0) {
[…]

### 2018-11-27 R/set_software.R

@@ -1,0 +1,138 @@
+#' set_software
+#'
+#' @param codemeta codemeta object, see examples
+#' @return an eml software element
+#' @export
+#' @examples
+#' cm <- jsonlite::read_json(system.file("extdata/codemeta.json", package = "EML"))
+#' software <- set_software(cm)
+#' my_eml <- eml$eml(packageId = "eml-1.2", system = "knb", software = software)
+#'
+#' # write_eml(my_eml, "test.xml")
+set_software <- function(codemeta) {
+  cm <- codemeta
+  eml$software(
+    id = cm[["id"]],
+    alternateIdentifier = cm$identifier,
+    title = cm$name,
+    creator = schema_to_eml(cm$author),
+    associatedParty = schema_to_eml(cm$contributor),
+    pubDate = cm$datePublished,
+    language = cm$language,
+    abstract = cm$description,
+    keywordSet = if (!is.null(cm$keywords)) {
+      eml$keywordSet(
+        keyword = cm$keywords,
+        keywordThesaurus = "codemeta"
+      )
+    },
+    licenseURL = cm$license,
+    implementation =
+      eml$implementation(
+        distribution = list(online = list(url = guess_url(cm))),
+        size = cm$fileSize,
+        operatingSystem = cm$operatingSystem,
+        machineProcessor = cm$processorRequirements,
+        diskUsage = cm$storageRequirements,
+        runtimeMemoryUsage = cm$memoryRequirements,
+        programmingLanguage = parse_programmingLanguage(cm$programmingLanguage)
+      ),
+    dependency = get_dependencies(cm)
+  )
+}
+
+
+## internal helper routines
+set_implementation <- function(cm) {
+  if (is.na(guess_url(cm))) {
+    return(NULL)
+  }
+
+  eml$implementation(
+    distribution = list(online = list(url = guess_url(cm))),
+    size = cm$fileSize,
+    operatingSystem = cm$operatingSystem,
+    machineProcessor = cm$processorRequirements,
+    diskUsage = cm$storageRequirements,
+    runtimeMemoryUsage = cm$memoryRequirements,
+    programmingLanguage = parse_programmingLanguage(cm$programmingLanguage)
+  )
+}
+
+
+get_dependencies <- function(cm) {
+  ## Drop dependencies we cannot document
+  out <- lapply(cm$softwareRequirements, set_software)
[…]

### 2018-11-27 R/template_constructor.R

@@ -1,0 +1,26 @@
+
+# eml_creator <- template_constructor("creator")
+template_constructor <- function(object) {
+  args <- emld::template(object)
+  f <- paste0(
+    "function(",
+    paste(names(args), "=", args, collapse = ", "),
+    "){ list(",
+    paste(names(args), "=", names(args), collapse = ","),
+    ")}"
+  )
+  eval(parse(text = f))
+}
+
+
+# who <- readLines(system.file("extdata/complexTypes.txt", package="EML"))
+# names(who) <- who
+# construct <- lapply(who, template_constructor)
+
+#' eml
+#'
+#' @docType data
+#' @keywords datasets
+#' @name eml
+#' @format A list with constructor functions
+NULL

### 2018-11-27 R/write_eml.R

@@ -1,7 +1,9 @@
+
+
 #' write_eml
 #'
 #' write_eml
-#' @param eml an eml class object
+#' @param eml an emld class object
 #' @param file file name to write XML.
 #' @param namespaces named character vector of additional XML namespaces to use.
 #' @param ns root namespace abbreviation
@@ -9,83 +11,24 @@
 #' @return If file is not specified, the result is a character string containing
 #'    the resulting XML content. Otherwise return silently.
 #' @export
-#' @import methods
-#' @importFrom xml2 write_xml xml_set_namespace xml_name xml_ns
-#'  xml_find_all xml_remove xml_root
 #' @importFrom uuid UUIDgenerate
+#' @importFrom xml2 write_xml xml_set_namespace
 #' @examples
-#' f <- system.file("examples", "example-eml-valid.xml", package = "EML")
+#' f <- system.file("extdata", "example.xml", package = "emld")
 #' eml <- read_eml(f)
 #' write_eml(eml, "test.xml")
+#' eml_validate("test.xml")
 #' unlink("test.xml") # clean up
 write_eml <- function(eml,
                       file,
                       namespaces = NULL,
                       ns = "eml",
                       ...) {
-
-  ## Make sure `eml` node has a schemaLocation
-  if(is(eml, "eml") && is_blank(eml@schemaLocation))
-    eml@schemaLocation@.Data <- "eml://ecoinformatics.org/eml-2.1.1 eml.xsd"
-
-  ## By default, we use UUID system to generate packageId
-  if("system" %in% slotNames(eml) && is_blank(eml@system))
-    slot(eml,"system") <- as("uuid", "xml_attribute")
-
-  ## By default, a packageId will be generated if one is not available (required by schema)
-  if(is(eml, "eml") && is_blank(eml@packageId))
-    slot(eml,"packageId") <- as(uuid::UUIDgenerate(), "xml_attribute")
-
-  # id <- basename(tempfile("eml"))
-
-  ## use default namespaces if not provided
-  if(is.null(namespaces))
-    namespaces <- eml_namespaces
-
-  ## Convert to xml
-  node <- s4_to_xml(eml, ns = namespaces)
-  root <- xml2::xml_root(node)
-
-  prune_empty(root)
-
-  ## setting root element ns doesn't appear to do anything:
-  #xml2::xml_set_namespace(tmp, ns,
[…]

### 2018-11-27 R/xml_to_s4.R

@@ -1,150 +1,0 @@
-## Used as the basis for all coercion methods from xml_node
-## to EML's S4 objects.
-
-## Assumes slot is named by the name of the repeated element
-## but has type "ListOf<element>" which contains a list containing
-## the repeated elements.
-
-#' @importFrom xml2 xml_attrs xml_name xml_children xml_text
-#' xml_set_text xml_set_attr xml_set_name xml_remove
-xml_to_s4 <- function(node){
-  node_name <- fix_protected_names( xml2::xml_name(node) )
-  attrs <- xml2::xml_attrs(node)
-  children <-  xml2::xml_children(node)
-  xml_names <- as.character(xml2::xml_name(children))
-  xml_attr_names <- names(attrs)
-
-  s4 <- new(node_name)
-  slot_classes <- get_slots(node_name)
-  s4_names <- sapply(names(slot_classes), strip_prefix)
-  names(slot_classes) <- s4_names
-
-  ## sub-classes/attrs are in the xml but do not have an s4 slot, such as: xmlns attributes (additional namespaces)
-  subclasses <- xml_names[!xml_names %in% s4_names]
-  not_sub <- xml_names[xml_names %in% s4_names]
-  subattrs <- xml_attr_names[!xml_attr_names %in% s4_names]
-  not_subattrs <- xml_attr_names[xml_attr_names %in% s4_names]
-  subclasses <- c(subclasses, subattrs)
-
-  ## metanames are s4 slots referring to a collection of xml elements, such as ResourceGroup, BoundsGroup, etc.
-  ## These always start with a capital (but are not in all caps, e.g. ISSN is not a metaclass), and would be an inherited class but that doesn't preserve slot order, so we use them a slot
-  metanames <- s4_names[grepl("^[A-Z][a-z]", s4_names)]
-  metaclasses <- lapply(metanames, get_slots)
-  names(metaclasses) <- metanames
-
-  ## fill in attributes for which we have a slot
-  for (child in not_subattrs) {
-    slot(s4, child) <- new("xml_attribute", attrs[[child]])
-  }
-
-
-
-  ## InlineType nodes are left as xml
-  if (inherits(s4, "InlineType")) {
-
-    kids <- xml_children(node)
-    if(length(kids) > 0){
-      s4@.Data <- list(kids)
-    } else {
-      s4@.Data <- list(node)
-    }
-  } else {
-
-    ## Simple data node
[…]

### 2018-11-27 data-raw/construct_complexTypes.R

@@ -1,0 +1,23 @@
+## Small test case
+# who <- list("creator", "individualName", "contact", "dataTable")
+# names(who) <- who
+# construct <- lapply(who, template_constructor)
+
+options("emld_db" = "eml-2.1.1")
+## Full database
+db <- purrr::compact(
+  jsonlite::read_json(paste0("https://raw.githubusercontent.com/cboettig/emld/",
+                             "master/data-raw/", getOption("emld_db"), ".json"))
+)
+db["UrlType"] <- NULL # 'function' as argument, not needed
+db["url"] <- NULL
+who <- names(db)
+writeLines(who, "inst/extdata/complexTypes.txt")
+
+
+# who <- readLines(system.file("extdata/complexTypes.txt", package="EML"))
+names(who) <- who
+eml <- lapply(who, EML:::template_constructor)
+
+
+devtools::use_data(eml, overwrite = TRUE)

### 2018-11-27 inst/create-package/post_process.R

@@ -1,118 +1,0 @@
-set_dummy_class <- function(class, file = "R/classes.R") {
-  write(sprintf("setClass('%s')", class), file, append = TRUE)
-}
-
-
-replace_class <- function(class, new, file = "R/classes.R") {
-  R <- readLines(file)
-  R <- gsub(sprintf("^setClass\\('%s'.*", class), new, R)
-  write(R, file)
-}
-
-
-move_up_n <- function(class, file = "R/classes.R", n = 2) {
-  R <- readLines(file)
-  N <- length(R)
-  i <- grep(sprintf("^setClass\\('%s'.*", class), R)[[1]]
-  write(c(R[1:(i - n)], R[i], R[(i - n + 1):(i - 1)], R[(i + 1):N]), file)
-}
-
-
-move_down_n <- function(class, file = "R/classes.R", n = 2) {
-  R <- readLines(file)
-  N <- length(R)
-  i <- grep(sprintf("^setClass\\('%s'.*", class), R)[[1]]
-  write(c(R[1:(i - 1)], R[(i + 1):(i + n)], R[i], R[(i + n + 1):N]), file)
-}
-
-
-move_to_top <- function(class, file = "R/classes.R") {
-  R <- readLines(file)
-  i <- grep(sprintf("^setClass\\('%s'.*", class), R)
-  write(c(R[i], R[-i]), file)
-}
-
-
-move_to_end <- function(class, file = "R/classes.R") {
-  R <- readLines(file)
-  i <- grep(sprintf("^setClass\\('%s'.*", class), R)
-  write(c(R[-i], R[i]), file)
-}
-
-
-drop_method <- function(class, file) {
-  M <- readLines(file)
-  #  M <- gsub(sprintf("^setAs\\('%s',.*", class), "", M)
-  #  M <- gsub(sprintf("setAs\\('xml_node', '%s',.*", class), "", M)
-  M <- gsub(sprintf("setMethod\\(initialize, '%s'.*", class), "", M)
-  write(M, file)
-}
-
-# make sure that " = 'complex'" etc also gets fixed. not a problem if class isn't used as a type.
-fix_protected <- function(file) {
-  R <- readLines(file)
-  R <-
-    gsub("^setClass\\('language'", "setClass('eml:language'", R)  ## not used? no "= 'language'"
-  R <-
[…]

### 2018-11-27 inst/examples/hf-205.R

@@ -2,12 +2,12 @@
 ## Works only with dev version dplyr due to dplyr bug
 attributes <-
   dplyr::frame_data(
-    ~attributeName,  ~formatString, ~definition,         ~attributeDefinition,
-    "run.num",        NA,           "which run number", "which run number (=block). Range: 1 - 6. (integer)",
-    "year",          "YYYY",        NA,                 "year, 2012",
-    "day",           "DDD",         NA,                 "Julian day. Range: 170 - 209.",
-    "hour.min",      "hhmm",        NA,                 "hour and minute of observation. Range 1 - 2400 (integer)",
-    "i.flag",        NA,            NA,                 "is variable Real, Interpolated or Bad (character/factor)",
-    "variable",      NA,            NA,                 "what variable being measured in what treatment (character/factor).",
-    "value.i",       NA,            NA,                 "value of measured variable for run.num on year/day/hour.min."
+    ~attributeName, ~formatString, ~definition, ~attributeDefinition,
+    "run.num", NA, "which run number", "which run number (=block). Range: 1 - 6. (integer)",
+    "year", "YYYY", NA, "year, 2012",
+    "day", "DDD", NA, "Julian day. Range: 170 - 209.",
+    "hour.min", "hhmm", NA, "hour and minute of observation. Range 1 - 2400 (integer)",
+    "i.flag", NA, NA, "is variable Real, Interpolated or Bad (character/factor)",
+    "variable", NA, NA, "what variable being measured in what treatment (character/factor).",
+    "value.i", NA, NA, "value of measured variable for run.num on year/day/hour.min."
   )

### 2018-11-27 tests/testthat/test-additional-tests.R

@@ -1,0 +1,49 @@
+testthat::context("Additional tests")
+
+test_that("template constructor", {
+  creator <- EML:::template_constructor("creator")
+  expect_is(creator, "function")
+})
+
+test_that("we can validate an xml_document", {
+  f <- system.file("extdata", "example.xml", package = "emld")
+  x <- xml2::read_xml(f)
+  expect_true(eml_validate(x))
+})
+
+
+test_that("set_physical can compute file size automatically", {
+  #file.copy(system.file("examples","hf205-01-TPexp1.csv", package="EML"),
+  #          "hf205-01-TPexp1.csv")
+  skip_if_not(file.exists("hf205-01-TPexp1.csv"))
+  expect_message(
+    physical <-
+      set_physical("hf205-01-TPexp1.csv"),
+    "calculated file size"
+  )
+})
+
+test_that("set_method with additional input files", {
+  
+  skip_on_appveyor()
+  
+  
+  coverage <-
+    set_coverage(
+      date = c("2012-06-01", "2013-12-31"),
+      sci_names = "Sarracenia purpurea",
+      geographicDescription = "text",
+      west = -122.44, east = -117.15,
+      north = 37.38, south = 30.00,
+      altitudeMin = 160, altitudeMaximum = 330,
+      altitudeUnits = "meter"
+    )
+  expect_is(coverage$geographicCoverage, "list")
+  file <- system.file("examples", "hf205-methods.docx", package = "EML")
+  methods <- set_methods(file,
+    qualityControl_file = file,
+    sampling_file = file,
+    sampling_coverage = coverage
+  )
+  expect_is(methods$sampling, "list")
+})

### 2018-11-27 tests/testthat/test-testfiles.R

@@ -1,52 +1,0 @@
-testthat::context("We can parse & serialize test files")
-
-library(xml2)
-xml_tests <- list.files("inst/xsd/test/", "^eml-.*\\.xml")
-## eml-unitDictionary is not EML but STMML; won't validate against EML schema by itself.
-xml_tests <- xml_tests[!xml_tests %in% c("eml-unitDictionary.xml")]
-
-
-
-all_test_examples<- function(xml) {
-  testthat::test_that(xml, {
-
-    message(paste("testing", xml))
-
-    ## Read EML
-    f <- system.file(paste0("xsd/test/", xml), package = "EML")
-    eml <- read_eml(f)
-    testthat::expect_true(isS4(eml))
-
-    ## Because root element in many test files is not "eml" but some sub-class like "dataset" or "access", we need to be explicit about namespace
-    original <- xml2::read_xml(f)
-    namespaces <- xml_ns(original)
-    schemaLocation <- strsplit(xml_attr(original, "schemaLocation"), "\\s+")[[1]]
-    i <- grep(schemaLocation[1], namespaces)
-    ns <- names(namespaces[i])
-
-    ## Write and Validate EML (handling explicit namespacing)
-    write_eml(eml, "unittest.xml", namespaces = namespaces, ns = ns)
-
-    v <- eml_validate("unittest.xml")
-    testthat::expect_true(v)
-
-    ## Check no elements were lost:
-    original <- xml_name( xml_find_all(read_xml(f), "//*") )
-    test <- xml_name(xml_find_all(xml2::read_xml("unittest.xml"), "//*") )
-
-    ## identical number of elements
-    testthat::expect_identical(length(original), length(test))
-    ## elements are identical, modulo ordering
-    testthat::expect_identical(sort(original), sort(test))
-    ## strictly identical:
-    if (!(xml == 'eml-physical.xml'))      # Skip known issue, ordering need not be identical(?)
-      testthat::expect_identical(original, test)
-
-    ## Clean up
-    unlink("unittest.xml")
-    })
-}
-
-out <- lapply(xml_tests, all_test_examples)
-
-

### 2018-11-27 README.Rmd
rebuild docs

--- 2018-09-14 README.Rmd
@@
 ---
-output:
-  md_document:
-    variant: gfm
+output: github_document
 ---
 
-[![Build Status](https://travis-ci.org/cboettig/eml2.svg?branch=master)](https://travis-ci.org/cboettig/eml2)
-[![Coverage Status](https://img.shields.io/codecov/c/github/cboettig/eml2/master.svg)](https://codecov.io/github/cboettig/eml2?branch=master)
-[![stability-experimental](https://img.shields.io/badge/stability-experimental-orange.svg)](https://github.com/joethorley/stability-badges#experimental)
+[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
+[![Build Status](https://travis-ci.org/ropensci/EML.svg?branch=master)](https://travis-ci.org/ropensci/EML)
+[![Windows build status](https://ci.appveyor.com/api/projects/status/u2gw24yfkvxgny96?svg=true)](https://ci.appveyor.com/project/cboettig/eml)
+[![codecov](https://codecov.io/gh/ropensci/EML/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/EML)
+[![CRAN status](https://www.r-pkg.org/badges/version/EML)](https://cran.r-project.org/package=EML)
 
@@
 
-# eml2
+# EML
 
-The goal of `eml2` is to provide both a drop-in replacement for the higher-level functions of the existing EML package while also providing additional functionality.
+The goal of `EML` is to provide both a drop-in replacement for the higher-level functions of the existing EML package while also providing additional functionality.
 
-`eml2` uses only simple and familiar list structures (S3 classes) instead of the more cumbersome use of S4 found in the original EML.
+`EML` uses only simple and familiar list structures (S3 classes) instead of the more cumbersome use of S4 found in the original EML.
 While the higher-level functions are identical, this makes it easier to for most users and developers to work with `eml` objects and also to write their own functions for creating and manipulating EML objects.
-Under the hood, `eml2` relies on the [emld](https://github.com/cboettig/emld) package, which uses a Linked Data representation for EML.
+Under the hood, `EML` relies on the [emld](https://github.com/cboettig/emld) package, which uses a Linked Data representation for EML.
 It is this approach which lets us combine the simplicity of lists with the specificy required by the XML schema.
@@
 ```{r message=FALSE, warning=FALSE}
-library(eml2)
+library(EML)
 ```
@@
 
-Here we show a the creation of a relatively complete EML document using `eml2`.
+Here we show a the creation of a relatively complete EML document using `EML`.
 This closely parallels the function calls shown in the original EML [R-package vignette](https://ropensci.github.io/EML/articles/creating-EML.html).
@@
 The original EML R package defines a set of higher-level `set_*` methods to facilitate the creation of complex metadata structures.
-`eml2` provides these same methods, taking the same arguments for `set_coverage`, `set_attributes`, `set_physical`, `set_methods` and `set_textType`, as illustrated here:
+`EML` provides these same methods, taking the same arguments for `set_coverage`, `set_attributes`, `set_physical`, `set_methods` and `set_textType`, as illustrated here:
 
@@
 ```{r}
-attributes <- read.table(system.file("extdata/hf205_attributes.csv", package = "eml2"))
-factors <- read.table(system.file("extdata/hf205_factors.csv", package = "eml2"))
+attributes <- read.table(system.file("extdata/hf205_attributes.csv", package = "EML"))
+factors <- read.table(system.file("extdata/hf205_factors.csv", package = "EML"))
 attributeList <- 

### 2018-11-27 README.Rmd
update readme

@@
 [![CRAN status](https://www.r-pkg.org/badges/version/EML)](https://cran.r-project.org/package=EML)
+![downloads](http://cranlogs.r-pkg.org/badges/grand-total/RNeXML)
+[![DOI](https://zenodo.org/badge/10894022.svg)](https://zenodo.org/badge/latestdoi/10894022)
 

### 2018-11-27 README.Rmd
rebuild docs

@@
 
-The goal of `EML` is to provide both a drop-in replacement for the higher-level functions of the existing EML package while also providing additional functionality.
+EML is a widely used metadata standard in the ecological and environmental sciences.
+We strongly recommend that interested users visit the [EML Homepage](https://knb.ecoinformatics.org/#external//emlparser/docs/index.html) for an introduction and thorough documentation of the standard.
+Additionally, the scientific article *[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](http://doi.org/10.1146/annurev.ecolsys.37.091305.110031)* provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly hetergenous data that cannot be easily reduced to a traditional vertically integrated database.
+At this time, the `EML` R package provides support for the serializing and parsing of all low-level EML concepts, but still assumes some familiarity with the EML standard, particularly for users seeking to create their own EML files.
+We hope to add more higher-level functions which will make such familiarity less essential in future development.
 
-`EML` uses only simple and familiar list structures (S3 classes) instead of the more cumbersome use of S4 found in the original EML.
+## Notes on the EML v2.0 Release
+
+`EML` v2.0 is a complete re-write which aims to provide both a drop-in replacement for the higher-level functions of the existing EML package while also providing additional functionality.
+This new `EML` version uses only simple and familiar list structures (S3 classes) instead of the more cumbersome use of S4 found in the original `EML`.
 While the higher-level functions are identical, this makes it easier to for most users and developers to work with `eml` objects and also to write their own functions for creating and manipulating EML objects.
@@
 It is this approach which lets us combine the simplicity of lists with the specificy required by the XML schema.
+
+This revision also supports the upcoming release of the EML 2.2 specification.
 
@@
 
+## Setting the version
+
+EML will use the latest EML specification by default.  To switch to a different version, set the option `emld_db`, like so:
+
+```{r}
+options("emld_db" = "eml-2.1.1")
+```
+
+Switch back to the 2.2.0 release:
+
+```{r}
+options("emld_db" = "eml-2.2.0")
+```
+
 ```{r include = FALSE}

### 2019-01-30 README.Rmd
update pkgdown with icon
valid minimal example in README

@@
 ---
+
+# EML <img src="man/figures/logo.svg" align="right" alt="" width="120" />
 
@@
 ```
-
-# EML
 
@@
 ```{r}
-me <- list(individualName = list(givenName = "Carl", surName = "Boettiger"), id="cboettig")
+me <- list(individualName = list(givenName = "Carl", surName = "Boettiger"))
 my_eml <- list(dataset = list(

### 2019-03-12 R/shiny_attributes.R
replace all options("emld_db") calls with emld::eml_version()

@@ -25,7 +25,7 @@
 #' out <- shiny_attributes(data, NULL)
 #' 
 #' # from exisiting attributes
-#' file <- system.file("tests", options("emld_db"),
+#' file <- system.file("tests", emld::eml_version(),
 #'   "eml-datasetWithAttributelevelMethods.xml",
 #'   package = "emld"
 #' )

### 2019-04-22 README.Rmd
replace all options("emld_db") calls with emld::eml_version()
update spelling checks

@@
 We strongly recommend that interested users visit the [EML Homepage](https://knb.ecoinformatics.org/#external//emlparser/docs/index.html) for an introduction and thorough documentation of the standard.
-Additionally, the scientific article *[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](http://doi.org/10.1146/annurev.ecolsys.37.091305.110031)* provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly hetergenous data that cannot be easily reduced to a traditional vertically integrated database.
+Additionally, the scientific article *[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](http://doi.org/10.1146/annurev.ecolsys.37.091305.110031)* provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly heterogeneous data that cannot be easily reduced to a traditional vertically integrated database.
 At this time, the `EML` R package provides support for the serializing and parsing of all low-level EML concepts, but still assumes some familiarity with the EML standard, particularly for users seeking to create their own EML files.
@@
 Under the hood, `EML` relies on the [emld](https://github.com/cboettig/emld) package, which uses a Linked Data representation for EML.
-It is this approach which lets us combine the simplicity of lists with the specificy required by the XML schema.
+It is this approach which lets us combine the simplicity of lists with the specificity required by the XML schema.
 
@@
 my_eml <- list(dataset = list(
-              title = "A Mimimal Valid EML Dataset",
+              title = "A Minimal Valid EML Dataset",
               creator = me,
@@
 
-EML will use the latest EML specification by default.  To switch to a different version, set the option `emld_db`, like so:
+EML will use the latest EML specification by default.  To switch to a different version, use `emld::eml_version()`
 
 ```{r}
-options("emld_db" = "eml-2.1.1")
+emld::eml_version("eml-2.1.1")
 ```
@@
 ```{r}
-options("emld_db" = "eml-2.2.0")
+emld::eml_version("eml-2.2.0")
 ```

### 2019-04-22 R/set_physical.R

@@ -9,7 +9,7 @@
 #' @param numFooterLines Number of footer lines following data. Lines are determined by the physicalLineDelimiter, or if it is absent, by the recordDelimiter. This value indicated the number of footer lines that should be skipped after parsing the data. If this value is omitted, parsers should assume the data continues to the end of the data stream.
 #' @param recordDelimiter This element specifies the record delimiter character when the format is text. The record delimiter is usually a linefeed (\\n) on UNIX, a carriage return (\\r) on MacOS, or both (\\r\\n) on Windows/DOS. Multiline records are usually delimited with two line ending characters, for example on UNIX it would be two linefeed characters (\\n\\n). As record delimiters are often non-printing characters, one can use either the special value "\\n" to represent a linefeed (ASCII 0x0a) and "\\r" to represent a carriage return (ASCII 0x0d). Alternatively, one can use the hex value to represent character values (e.g., 0x0a).
 #' @param fieldDelimiter "," character by default (for csv files). This element specifies a character to be used in the object for indicating the ending column for an attribute. The delimiter character itself is not part of the attribute value, but rather is present in the column following the last character of the value. Typical delimiter characters include commas, tabs, spaces, and semicolons. The only time the fieldDelimiter character is not interpreted as a delimiter is if it is contained in a quoted string (see quoteCharacter) or is immediately preceded by a literalCharacter. Non-printable quote characters can be provided as their hex values, and for tab characters by its ASCII string "\\t". Processors should assume that the field starts in the column following the previous field if the previous field was fixed, or in the column following the delimiter from the previous field if the previous field was delimited.
-#' @param collapseDelimiters The collapseDelimiters element specifies whether sequential delimiters should be treated as a single delimiter or multiple delimiters. An example is when a space delimiter is used; often there may be several repeated spaces that should be treated as a single delimiter, but not always. The valid values are yes or no. If it is set to yes, then consecutive delimiters will be collapsed to one. If set to no or absent, then consecutive delimiters will be treated as separate delimiters. Default behaviour is no; hence, consecutive delimiters will be treated as separate delimiters, by default.
[…]

### 2019-05-19 README.Rmd
Hyperlink DOIs to preferred resolver

@@
 We strongly recommend that interested users visit the [EML Homepage](https://knb.ecoinformatics.org/#external//emlparser/docs/index.html) for an introduction and thorough documentation of the standard.
-Additionally, the scientific article *[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](http://doi.org/10.1146/annurev.ecolsys.37.091305.110031)* provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly heterogeneous data that cannot be easily reduced to a traditional vertically integrated database.
+Additionally, the scientific article *[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](https://doi.org/10.1146/annurev.ecolsys.37.091305.110031)* provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly heterogeneous data that cannot be easily reduced to a traditional vertically integrated database.
 At this time, the `EML` R package provides support for the serializing and parsing of all low-level EML concepts, but still assumes some familiarity with the EML standard, particularly for users seeking to create their own EML files.

### 2019-12-09 README.Rmd
Minor fix (proposal) on string 70 (#281)

--- 2019-04-22 README.Rmd
@@
 
-Here we show a the creation of a relatively complete EML document using `EML`.
-This closely parallels the function calls shown in the original EML [R-package vignette](https://ropensci.github.io/EML/articles/creating-EML.html).
+Here we show the creation of a relatively complete EML document using `EML`.
+This closely parallels the function calls shown in the original EML [R-package vignette](https://docs.ropensci.org/EML/articles/creating-EML.html).
 

### 2019-12-09 README.Rmd
Prep for CRAN's requested patch, #290

@@
 EML is a widely used metadata standard in the ecological and environmental sciences.
-We strongly recommend that interested users visit the [EML Homepage](https://knb.ecoinformatics.org/#external//emlparser/docs/index.html) for an introduction and thorough documentation of the standard.
+We strongly recommend that interested users visit the [EML Homepage](https://eml.ecoinformatics.org/) for an introduction and thorough documentation of the standard.
 Additionally, the scientific article *[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](http://doi.org/10.1146/annurev.ecolsys.37.091305.110031)* provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly heterogeneous data that cannot be easily reduced to a traditional vertically integrated database.
@@
 
-This revision also supports the upcoming release of the EML 2.2 specification.
+This revision also supports the **[recently released EML 2.2.0 specification](https://eml.ecoinformatics.org/whats-new-in-eml-2-2-0.html)**.
 

### 2019-12-09 R/set_TextType.R

@@ -1,9 +1,11 @@
-## Developer note: Some functions not fully compatible with pandoc < 2.0 
+## Developer note: Some functions not fully compatible with pandoc < 2.0
 
 #' set_TextType
 #'
-#' For any EML element of class TextType, this function can be used to generate the appropriate EML from a markdown-formatted file.
-#' @param text a plain text character string which will be used directly as the content of the node if no file is given
+#' For any EML element of class TextType, this function can be used to generate
+#'  the appropriate EML from a markdown-formatted file.
+#' @param text a plain text character string which will be used directly as the content
+#'  of the node if no file is given
 #' @param file path to a file providing formatted input text, see details.
 #' @return a TextType object that can be coerced into any element inheriting from TextType, see examples
 #' @importFrom tools file_ext
@@ -20,23 +22,23 @@
 #' ## using a simple character string
 #' a <- set_TextType(text = "This is the abstract")
 #' as(a, "abstract")
-#' 
+#'
 #' ## Using an external markdown file
 #' f <- system.file("examples/hf205-abstract.md", package = "EML")
 #' a <- set_TextType(f)
 #' as(a, "abstract")
-#' 
+#'
 #' ## Can also import from methods written in a .docx MS Word file.
 #' f <- system.file("examples/hf205-abstract.docx", package = "EML")
 #' a <- set_TextType(f)
 #' as(a, "abstract")
-#' 
+#'
 #' ## Documents with title headings use `section` instead of `para` notation
 #' f <- system.file("examples/hf205-methods.docx", package = "EML")
 #' d <- set_TextType(f)
 #' as(d, "description")
 #' }
-#' 
+#'
 set_TextType <- function(file = NULL, text = NULL) {
   if (!is.null(text)) {
     TextType <- text
@@ -92,6 +94,13 @@ to_docbook <- function(file = NULL) {
         call. = FALSE
       )
     }
+    if (!rmarkdown::pandoc_available()) {
+      stop(paste("Pandoc is required to convert to Docbook format.",
+              "Please supply input text directly"),
+              call. = FALSE
+             )
+    }
+
     pandoc_convert <-
       getExportedValue("rmarkdown", "pandoc_convert")
 

### 2020-01-20 README.Rmd
Add a set_packageId() method
Package ID will now be set explicitly be a dedicated function, and *not set automagically* by write_eml().  write_* functions should not have unexpected side-effects that modify the output data!

@@
             )
+my_eml <- set_packageId(my_eml)
 

### 2020-11-03 README.Rmd
nonsense about redirects

--- 2019-12-09 README.Rmd
@@
 We strongly recommend that interested users visit the [EML Homepage](https://eml.ecoinformatics.org/) for an introduction and thorough documentation of the standard.
-Additionally, the scientific article *[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](http://doi.org/10.1146/annurev.ecolsys.37.091305.110031)* provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly heterogeneous data that cannot be easily reduced to a traditional vertically integrated database.
+Additionally, the scientific article *[The New Bioinformatics: Integrating Ecological Data from the Gene to the Biosphere (Jones et al 2006)](https://doi.org/10.1146/annurev.ecolsys.37.091305.110031)* provides an excellent introduction into the role EML plays in building metadata-driven data repositories to address the needs of highly heterogeneous data that cannot be easily reduced to a traditional vertically integrated database.
 At this time, the `EML` R package provides support for the serializing and parsing of all low-level EML concepts, but still assumes some familiarity with the EML standard, particularly for users seeking to create their own EML files.
@@
 While the higher-level functions are identical, this makes it easier to for most users and developers to work with `eml` objects and also to write their own functions for creating and manipulating EML objects.
-Under the hood, `EML` relies on the [emld](https://github.com/cboettig/emld) package, which uses a Linked Data representation for EML.
+Under the hood, `EML` relies on the [emld](https://github.com/cboettig/emld/) package, which uses a Linked Data representation for EML.
 It is this approach which lets us combine the simplicity of lists with the specificity required by the XML schema.

### 2021-11-03 README.Rmd
more redirects

@@
 
-[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
+[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
 [![Build Status](https://travis-ci.org/ropensci/EML.svg?branch=master)](https://travis-ci.org/ropensci/EML)
@@
 While the higher-level functions are identical, this makes it easier to for most users and developers to work with `eml` objects and also to write their own functions for creating and manipulating EML objects.
-Under the hood, `EML` relies on the [emld](https://github.com/cboettig/emld/) package, which uses a Linked Data representation for EML.
+Under the hood, `EML` relies on the [emld](https://github.com/ropensci/emld/) package, which uses a Linked Data representation for EML.
 It is this approach which lets us combine the simplicity of lists with the specificity required by the XML schema.

### 2022-04-28 README.Rmd
Fix three URLs to satisfy CRAN checks
Note that this breaks the DOI in the creating-EML vignette because the PNAS DOI 503s without either JS or cookies enabled (not sure which) due to that URL being fronted with Cloudflare and probably some sort of anti-abuse config PNAS has with them. This is regrettable but being on CRAN is better than making DOIs URLs.

@@
 [![Windows build status](https://ci.appveyor.com/api/projects/status/u2gw24yfkvxgny96?svg=true)](https://ci.appveyor.com/project/cboettig/eml)
-[![codecov](https://codecov.io/gh/ropensci/EML/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/EML)
+[![codecov](https://codecov.io/gh/ropensci/EML/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ropensci/EML)
 [![CRAN status](https://www.r-pkg.org/badges/version/EML)](https://cran.r-project.org/package=EML)

### 2025-07-28 README.Rmd
url check

--- 2021-11-03 README.Rmd
@@
 [![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
-[![Build Status](https://travis-ci.org/ropensci/EML.svg?branch=master)](https://travis-ci.org/ropensci/EML)
+[![Build Status](https://app.travis-ci.com/ropensci/EML.svg?branch=master)](https://app.travis-ci.com/ropensci/EML)
 [![Windows build status](https://ci.appveyor.com/api/projects/status/u2gw24yfkvxgny96?svg=true)](https://ci.appveyor.com/project/cboettig/eml)
-[![codecov](https://codecov.io/gh/ropensci/EML/branch/master/graph/badge.svg)](https://codecov.io/gh/ropensci/EML)
+[![codecov](https://codecov.io/gh/ropensci/EML/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ropensci/EML)
 [![CRAN status](https://www.r-pkg.org/badges/version/EML)](https://cran.r-project.org/package=EML)
